### 2013-04-09 papers/Paper1/proplyd-bowshocks.tex
Very early draft of first paper
* papers/Paper1/proplyd-bowshocks.tex

\documentclass[useAMS,usenatbib]{mn2e}

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
\usepackage[varg]{txfonts}
% \usepackage[varg]{newtxmath}
% \usepackage{newtxtext}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{color}
\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

%% Bundled custom packages
\usepackage{aastex-compat}
\usepackage{astrojournals}

\title{Photoionized bow shock shells.  I. Wind-wind interactions in the Orion proplyds}

\newcommand\AddressCRyA{Centro de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:intro}

\section{Conclusions}
\label{sec:conc}

\end{document}