### 2014-09-21 Proplyd distribution.tex
Find the proplyd size distribution and fit it

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass{article}

    \usepackage{graphicx} % Used to insert images
    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{color} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2

    \definecolor{orange}{cmyk}{0,0.4,0.8,0.2}
    \definecolor{darkorange}{rgb}{.71,0.21,0.01}
    \definecolor{darkgreen}{rgb}{.12,.54,.11}
    \definecolor{myteal}{rgb}{.26, .44, .56}
    \definecolor{gray}{gray}{0.45}
    \definecolor{lightgray}{gray}{.95}
    \definecolor{mediumgray}{gray}{.8}
    \definecolor{inputbackground}{rgb}{.95, .95, .85}
    \definecolor{outputbackground}{rgb}{.95, .95, .95}
    \definecolor{traceback}{rgb}{1, .95, .95}
    % ansi colors
    \definecolor{red}{rgb}{.6,0,0}
    \definecolor{green}{rgb}{0,.65,0}
    \definecolor{brown}{rgb}{0.6,0.6,0}
    \definecolor{blue}{rgb}{0,.145,.698}
    \definecolor{purple}{rgb}{.698,.145,.698}
    \definecolor{cyan}{rgb}{0,.698,.698}
    \definecolor{lightgray}{gray}{0.5}

    % bright ansi colors
    \definecolor{darkgray}{gray}{0.25}
    \definecolor{lightred}{rgb}{1.0,0.39,0.28}
    \definecolor{lightgreen}{rgb}{0.48,0.99,0.0}
    \definecolor{lightblue}{rgb}{0.53,0.81,0.92}
    \definecolor{lightpurple}{rgb}{0.87,0.63,0.87}
    \definecolor{lightcyan}{rgb}{0.5,1.0,0.83}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{Proplyd distribution}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=blue,
      linkcolor=darkorange,
      citecolor=darkgreen,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}1}]:} \PY{o}{\PYZpc{}}\PY{k}{matplotlib} \PY{n}{inline}
\end{Verbatim}

    Calculations of distributions of proplyd distances and sizes and angles.
These translate into distributions of $\beta$.

    \subsection{Data sources}\label{data-sources}

We have the list of proplyds from Henney \& Arthur (1998) that we used
in Fig 12 of Tsamis et al (2013). This has distances and sizes.

We also have the Ricci et al (2008) catalog, which has distances but not
sizes.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}308}]:} \PY{k+kn}{from} \PY{n+nn}{astropy.table} \PY{k+kn}{import} \PY{n}{Table}
          \PY{k+kn}{import} \PY{n+nn}{numpy} \PY{k+kn}{as} \PY{n+nn}{np}
          \PY{k+kn}{import} \PY{n+nn}{seaborn} \PY{k+kn}{as} \PY{n+nn}{sns}
          \PY{k+kn}{from} \PY{n+nn}{scipy} \PY{k+kn}{import} \PY{n}{stats}
\end{Verbatim}

    Now use seaborn to get nicer graphs.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}359}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{set\PYZus{}palette}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{deep}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{desat}\PY{o}{=}\PY{l+m+mf}{0.6}\PY{p}{)}
          \PY{n}{sns}\PY{o}{.}\PY{n}{set\PYZus{}context}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{talk}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{rc}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{figure.figsize}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{p}{(}\PY{l+m+mi}{12}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}\PY{p}{\PYZcb{}}\PY{p}{)}
\end{Verbatim}

    \subsubsection{Henney \& Arthur (1998)
sample}\label{henney-arthur-1998-sample}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}3}]:} \PY{n}{datafile} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd\PYZhy{}Datasets/vicente\PYZhy{}vs\PYZhy{}HA98.dat}\PY{l+s}{\PYZdq{}}
        \PY{n}{data} \PY{o}{=} \PY{n}{Table}\PY{o}{.}\PY{n}{read}\PY{p}{(}\PY{n}{datafile}\PY{p}{,} \PY{n}{format}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{ascii}\PY{l+s}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}4}]:} \PY{n}{data}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}4}]:} <Table rows=28 names=('ID','d','r14','VicR14')>
        array([('163-323', 2.14, 2.2, 19.4), ('161-324', 6.05, 3.5, 25.8),
               ('168-328', 6.64, 2.8, 21.0), ('163-317', 6.91, 5.0, 27.1),
               ('166-316', 7.01, 2.5, 14.8), ('161-328', 7.74, 9.1, 20.0),
               ('167-317', 7.83, 7.9, 35.8), ('158-323', 9.42, 6.3, 30.6),
               ('158-326', 9.6, 11.3, 31.6), ('161-314', 10.24, 5.3, 17.1),
               ('158-327', 10.6, 16.6, 35.5), ('157-323', 10.97, 2.5, 18.4),
               ('171-334', 14.29, 4.7, 22.9), ('170-337', 16.2, 12.2, 36.8),
               ('176-325', 16.38, 6.9, 23.5), ('169-338', 16.47, 2.8, 16.1),
               ('154-324', 16.63, 3.2, 10.3), ('153-321', 16.97, 1.2, 8.4),
               ('159-338', 17.2, 5.0, 13.5), ('171-340', 19.11, 23.3, 23.2),
               ('152-319', 19.16, 18.2, 14.8), ('155-338', 20.48, 17.0, 29.7),
               ('173-341', 22.48, 4.1, 18.1), ('180-331', 25.12, 12.2, 29.3),
               ('177-341', 25.84, 20.4, 51.6), ('159-350', 28.35, 20.1, 44.5),
               ('182-413', 56.7, 36.9, 48.7), ('244-440', 142.3, 104.0, 180.6)], 
              dtype=[('ID', '<U7'), ('d', '<f8'), ('r14', '<f8'), ('VicR14', '<f8')])
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}5}]:} \PY{k+kn}{from} \PY{n+nn}{matplotlib} \PY{k+kn}{import} \PY{n}{pyplot} \PY{k}{as} \PY{n}{plt}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}360}]:} \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{30}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of proplyds with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Cumulative distribution of proplyds vs distance: HA98}\PY{l+s}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_10_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}361}]:} \PY{n}{m1} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{l+m+mf}{7.5}
          \PY{n}{m2} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{\PYZhy{}} \PY{l+m+mf}{13.0}\PY{p}{)} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{l+m+mf}{5.5}
          \PY{n}{m3} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{abs}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{\PYZhy{}} \PY{l+m+mf}{25.0}\PY{p}{)} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{l+m+mf}{6.5}
          \PY{n}{h1}\PY{p}{,} \PY{n}{edges} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{histogram}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m1}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{10}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{density}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
          \PY{n}{h2}\PY{p}{,} \PY{n}{edges} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{histogram}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m2}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{10}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{density}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
          \PY{n}{h3}\PY{p}{,} \PY{n}{edges} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{histogram}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m3}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{10}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{density}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{bar}\PY{p}{(}\PY{n}{edges}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{h1}\PY{p}{,} \PY{l+m+mf}{2.5}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.5}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{close: D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ \PYZlt{} 7.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{bar}\PY{p}{(}\PY{n}{edges}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{h2}\PY{p}{,} \PY{l+m+mf}{2.5}\PY{p}{,} \PY{n}{h1}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{g}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.5}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{medium: 7.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{ \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ \PYZlt{} 18.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{bar}\PY{p}{(}\PY{n}{edges}\PY{p}{[}\PY{p}{:}\PY{o}{\PYZhy{}}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{,} \PY{n}{h3}\PY{p}{,} \PY{l+m+mf}{2.5}\PY{p}{,} \PY{n}{h1} \PY{o}{+} \PY{n}{h2}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.5}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{far: D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ \PYZgt{} 18.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Proplyd i\PYZhy{}front radius, 1e14 cm}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Fraction of proplyds in each size bin}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Histogram of proplyd sizes: HA98}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_11_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}362}]:} \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{o}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{30.0}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{30.0}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Distance, arcsec}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Radius, 1e14 cm}\PY{l+s}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}362}]:} <matplotlib.text.Text at 0x119f77710>
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_12_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    We need to take account of binaries

    Now we will try out some of the seaborn tools for visualizing data
distributions. See
\href{http://web.stanford.edu/~mwaskom/software/seaborn/tutorial/plotting_distributions.html}{this
tutorial}.

    In the section on the Ricci data it occured to me that we need to mirror
the data about zero because the radius and separation are
positive-definite.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}926}]:} \PY{k}{def} \PY{n+nf}{mirrored}\PY{p}{(}\PY{n}{dataset}\PY{p}{)}\PY{p}{:}
              \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Expand a positive\PYZhy{}definite dataset (1d sequence) to include its negative mirror set\PYZdq{}\PYZdq{}\PYZdq{}}
              \PY{k}{assert} \PY{n}{np}\PY{o}{.}\PY{n}{all}\PY{p}{(}\PY{n}{dataset} \PY{o}{\PYZgt{}}\PY{o}{=} \PY{l+m+mf}{0.0}\PY{p}{)}
              \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{r\PYZus{}}\PY{p}{[}\PY{n}{dataset}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{n}{dataset}\PY{p}{]}

          \PY{k}{def} \PY{n+nf}{doubled}\PY{p}{(}\PY{n}{dataset}\PY{p}{)}\PY{p}{:}
              \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Double up a dataset (1d sequence)\PYZdq{}\PYZdq{}\PYZdq{}}
              \PY{k}{return} \PY{n}{np}\PY{o}{.}\PY{n}{r\PYZus{}}\PY{p}{[}\PY{n}{dataset}\PY{p}{,} \PY{n}{dataset}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}425}]:} \PY{n}{m30} \PY{o}{=} \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{l+m+mf}{30.0}
          \PY{k}{with} \PY{n}{sns}\PY{o}{.}\PY{n}{axes\PYZus{}style}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{white}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{:}
              \PY{n}{sns}\PY{o}{.}\PY{n}{jointplot}\PY{p}{(}\PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{)}\PY{p}{,} 
                            \PY{n}{kind}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{hex}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{9}\PY{p}{,} \PY{n}{xlim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{ylim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_17_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}446}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{kdeplot}\PY{p}{(}\PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{p}{,} 
                      \PY{n}{clip}\PY{o}{=}\PY{p}{(}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{40}\PY{p}{)}\PY{p}{,} \PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{)}\PY{p}{,}
                      \PY{n}{bw}\PY{o}{=}\PY{l+m+mf}{3.0}\PY{p}{,} \PY{n}{shade}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{cmap}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Reds}\PY{l+s}{\PYZdq{}}\PY{p}{,}
                      \PY{p}{)}
          \PY{n+nb+bp}{None}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_18_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}503}]:} \PY{n}{color} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsh{}774499}\PY{l+s}{\PYZdq{}}
          \PY{n}{g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{JointGrid}\PY{p}{(}\PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{9}\PY{p}{,} \PY{n}{xlim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{ylim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{)}
          \PY{c}{\PYZsh{} g.plot(sns.regplot, sns.distplot, stats.pearsonr);}
          \PY{n}{g}\PY{o}{.}\PY{n}{plot\PYZus{}marginals}\PY{p}{(}\PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{10}\PY{p}{,} \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{n+nb}{dict}\PY{p}{(}\PY{n}{bw}\PY{o}{=}\PY{l+m+mf}{5.0}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{KDE}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{plot\PYZus{}joint}\PY{p}{(}\PY{n}{sns}\PY{o}{.}\PY{n}{regplot}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{,} \PY{n}{ci}\PY{o}{=}\PY{l+m+mi}{98}\PY{p}{,} \PY{n}{order}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Cubic regression fit with 98}\PY{l+s+si}{\PYZpc{} c}\PY{l+s}{onfidence bounds}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Projected separation, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd radius \PYZdl{}r\PYZus{}0\PYZdl{}, \PYZdl{}10\PYZca{}\PYZob{}14\PYZcb{}\PYZdl{} cm}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}marg\PYZus{}x}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}marg\PYZus{}y}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{best}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{fig}\PY{o}{.}\PY{n}{tight\PYZus{}layout}\PY{p}{(}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{fig}\PY{o}{.}\PY{n}{savefig}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{proplyd\PYZhy{}joint\PYZhy{}separation\PYZhy{}size.pdf}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
          \PY{c}{\PYZsh{}g.annotate(stats.pearsonr);}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_19_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}494}]:} \PY{n}{stats}\PY{o}{.}\PY{n}{scoreatpercentile}\PY{err}{?}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}401}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{rugplot}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{height}\PY{o}{=}\PY{l+m+mf}{0.002}\PY{p}{)}
          \PY{n}{sns}\PY{o}{.}\PY{n}{kdeplot}\PY{p}{(}\PY{n}{mirrored}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{shade}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{n+nb+bp}{None}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}401}]:} (0.0, 200.0)
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_21_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    Now we will compare the sizes for different intervals of $D'$

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}407}]:} \PY{n}{datalist} \PY{o}{=} \PY{p}{[}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m1}\PY{p}{]}\PY{p}{,} \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m2}\PY{p}{]}\PY{p}{,} \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m3}\PY{p}{]}\PY{p}{]}
          \PY{n}{mirrored\PYZus{}datalist} \PY{o}{=} \PY{p}{[}\PY{n}{mirrored}\PY{p}{(}\PY{n}{x}\PY{p}{)} \PY{k}{for} \PY{n}{x} \PY{o+ow}{in} \PY{n}{datalist}\PY{p}{]}
          \PY{n}{namelist} \PY{o}{=} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{close: D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ \PYZlt{} 7.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{mid: 7.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{ \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ \PYZlt{} 18.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{far: D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ \PYZgt{} 18.5}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsq{}}\PY{p}{]}
\end{Verbatim}

    The simplest option is \texttt{boxplot}, which shows the median and
quartiles. By setting \texttt{whis=np.inf} we make the whiskers extend
out to the full range of the data.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}408}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{boxplot}\PY{p}{(}\PY{n}{datalist}\PY{p}{,} \PY{n}{names}\PY{o}{=}\PY{n}{namelist}\PY{p}{,} \PY{n}{whis}\PY{o}{=}\PY{n}{np}\PY{o}{.}\PY{n}{inf}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{pastel}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd size, 1e14 cm}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Projected distance band}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n+nb+bp}{None}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_25_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    Alternatively we can use \texttt{violinplot}, which shows the Kernel
Density Estimate (KDE) of the values in each category. By setting
\texttt{inner="points"} it shows the actual data points instead of the
quartiles.

I am a little uncomfortable of this way of presenting the data, since we
are forcing the distance to be categorical, when really it is
continuous.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}421}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{violinplot}\PY{p}{(}\PY{n}{mirrored\PYZus{}datalist}\PY{p}{,} 
                         \PY{n}{names}\PY{o}{=}\PY{n}{namelist}\PY{p}{,}
                         \PY{n}{inner}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{points}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{bw}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{PuBuGn}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{ylim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{n+nb+bp}{None}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd size, 1e14 cm}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Projected distance band}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n+nb+bp}{None}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_27_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}422}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{jointplot}\PY{err}{?}
\end{Verbatim}

    \subsubsection{Ricci et al (2008) sample}\label{ricci-et-al-2008-sample}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}9}]:} \PY{n}{rdata} \PY{o}{=} \PY{n}{Table}\PY{o}{.}\PY{n}{read}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd\PYZhy{}Datasets/ricci\PYZhy{}2008.dat}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{format}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{ascii}\PY{l+s}{\PYZsq{}}\PY{p}{)}
        \PY{n}{rdata}\PY{o}{.}\PY{n}{remove\PYZus{}row}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{)} \PY{c}{\PYZsh{} First row is rubbish}
        \PY{n}{rdata}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}9}]:} <Table rows=219 names=('Name','RAJ2000','DEJ2000','[OW94]','Type','Note')>
        masked\_array(data = [('4364-146', '05 34 36.44', '-05 21 45.95', '0', 'j', 'J')
         ('4466-324', '05 34 46.59', '-05 23 24.19', '0', 'j', 'J,B')
         ('4468-605', '05 34 46.76', '-05 26 04.79', '0', 'i', 'J')
         ('4538-311', '05 34 53.79', '-05 23 10.73', '0', 'rn', '0')
         ('4596-400', '05 34 59.56', '-05 24 00.19', '4596-400', 'i', '0')
         ('4582-635', '05 34 58.16', '-05 26 35.13', '0', 'i', '0')
         ('005-514', '05 35 00.47', '-05 25 14.34', '005-514', 'i', '0')
         ('006-439', '05 35 00.58', '-05 24 38.79', '0', 'j', 'J')
         ('016-149', '05 35 01.60', '-05 21 49.35', '0', 'rn', '0')
         ('038-627', '05 35 04.19', '-05 26 27.89', '038-627', 'i', '0')
         ('044-527', '05 35 04.42', '-05 25 27.40', '044-527', 'i', '0')
         ('046-3838', '05 35 04.61', '-05 38 38.00', '0', 'rn', '0')
         ('046-245', '05 35 04.63', '-05 22 44.85', '0', 'i', '0')
         ('049-143', '05 35 04.94', '-05 21 42.99', '0', 'i', '0')
         ('051-3541', '05 35 05.05', '-05 35 40.84', '0', 'rn', 'EO')
         ('053-717', '05 35 05.40', '-05 27 16.99', '0', 'd', '0')
         ('057-419', '05 35 05.73', '-05 24 18.55', '057-419', 'i', '0')
         ('061-401', '05 35 06.09', '-05 24 00.60', '061-401', 'i', '0')
         ('064-3335', '05 35 06.44', '-05 33 35.25', '0', 'i', '0')
         ('066-3251', '05 35 06.57', '-05 32 51.49', '0', 'i', '0')
         ('066-652', '05 35 06.59', '-05 26 51.99', '066-652', 'i', 'B')
         ('069-601', '05 35 06.91', '-05 26 00.60', '069-601', 'i', '0')
         ('072-135', '05 35 07.21', '-05 21 34.43', '072-135', 'i', 'EO')
         ('073-227', '05 35 07.27', '-05 22 26.56', '073-227', 'i', '0')
         ('078-3658', '05 35 07.84', '-05 36 58.15', '0', 'j', 'J,B')
         ('090-326', '05 35 09.02', '-05 23 26.20', '0', 'd', '0')
         ('093-822', '05 35 09.59', '-05 28 22.92', '093-822', 'i', '0')
         ('099-339', '05 35 09.89', '-05 23 38.50', '0', 'i', 'J')
         ('102-233', '05 35 10.13', '-05 22 32.74', '102-233', 'i', '0')
         ('102-021', '05 35 10.19', '-05 20 20.99', '102-021', 'i', '0')
         ('102-322', '05 35 10.20', '-05 23 21.56', '102-322', 'i', '0')
         ('106-417', '05 35 10.54', '-05 24 16.70', '106-417', 'i', '0')
         ('106-156', '05 35 10.58', '-05 21 56.24', '106-156', 'i', '0')
         ('109-246', '05 35 10.90', '-05 22 46.36', '109-246', 'i', '0')
         ('109-327', '05 35 10.95', '-05 23 26.45', '109-327', 'i', 'J')
         ('110-3035', '05 35 10.98', '-05 30 35.23', '0', 'rn', 'J')
         ('114-426', '05 35 11.30', '-05 24 26.50', '114-426', 'd', 'EO,RN')
         ('117-421', '05 35 11.65', '-05 24 21.50', '117-421', 'i', '0')
         ('117-352', '05 35 11.73', '-05 23 51.70', '0', 'i', '0')
         ('119-340', '05 35 11.90', '-05 23 39.70', '119-340', 'i', '0')
         ('121-1925', '05 35 12.09', '-05 19 24.80', '121-1925', 'd', '0')
         ('121-434', '05 35 12.12', '-05 24 33.80', '121-434', 'i', '0')
         ('124-132', '05 35 12.38', '-05 21 31.39', '124-132', 'i', 'J,EO,B')
         ('131-046', '05 35 13.05', '-05 20 45.79', '131-046', 'i', '0')
         ('131-247', '05 35 13.11', '-05 22 47.11', '131-247', 'i', 'J')
         ('132-1832', '05 35 13.24', '-05 18 32.95', '0', 'd', 'EO')
         ('132-042', '05 35 13.24', '-05 20 41.94', '132-042', 'i', 'J,EO')
         ('133-353', '05 35 13.30', '-05 23 53.05', '133-353', 'i', 'B')
         ('135-220', '05 35 13.51', '-05 22 19.49', '135-220', 'i', '0')
         ('138-207', '05 35 13.78', '-05 22 07.39', '138-207', 'i', '0')
         ('139-320', '05 35 13.92', '-05 23 20.16', '0', 'i', '0')
         ('141-520', '05 35 14.05', '-05 25 20.50', '141-520', 'i', 'FO')
         ('140-1952', '05 35 14.05', '-05 19 51.90', '140-1952', 'd', 'FO')
         ('142-301', '05 35 14.15', '-05 23 00.91', '142-301', 'i', '0')
         ('143-425', '05 35 14.27', '-05 24 24.55', '143-425', 'i', '0')
         ('144-522', '05 35 14.34', '-05 25 22.30', '144-522', 'i', '0')
         ('146-201', '05 35 14.61', '-05 22 00.94', '0', 'i', '0')
         ('147-323', '05 35 14.72', '-05 23 23.01', '0', 'i', '0')
         ('148-305', '05 35 14.80', '-05 23 04.76', '148-305', 'i', 'B')
         ('149-329', '05 35 14.92', '-05 23 29.05', '149-329', 'i', '0')
         ('150-147', '05 35 15.00', '-05 21 47.34', '0', 'i', '0')
         ('150-231', '05 35 15.02', '-05 22 31.11', '150-231', 'i', 'B')
         ('152-319', '05 35 15.20', '-05 23 18.81', '152-319', 'i', '0')
         ('152-738', '05 35 15.21', '-05 27 37.85', '152-738', 'i', '0')
         ('154-324', '05 35 15.35', '-05 23 24.11', '154-324', 'i', 'B')
         ('154-225', '05 35 15.37', '-05 22 25.35', '154-225', 'i', 'B')
         ('154-240', '05 35 15.38', '-05 22 39.85', '0', 'i', '0')
         ('155-338', '05 35 15.51', '-05 23 37.45', '155-338', 'i', '0')
         ('157-323', '05 35 15.72', '-05 23 22.59', '157-323', 'i', 'B')
         ('158-323', '05 35 15.83', '-05 23 22.59', '158-323', 'i', 'B')
         ('158-327', '05 35 15.79', '-05 23 26.51', '158-327', 'i', 'B')
         ('158-326', '05 35 15.81', '-05 23 25.51', '158-326', 'i', 'B')
         ('159-338', '05 35 15.90', '-05 23 38.00', '159-338', 'i', 'B')
         ('159-418', '05 35 15.90', '-05 24 17.85', '159-418', 'i', '0')
         ('159-221', '05 35 15.93', '-05 22 21.05', '159-221', 'd', '0')
         ('159-350', '05 35 15.96', '-05 23 50.30', '159-350', 'i', 'B')
         ('160-353', '05 35 16.01', '-05 23 53.00', '160-353', 'i', 'J')
         ('161-324', '05 35 16.05', '-05 23 24.35', '161-324', 'i', '0')
         ('161-328', '05 35 16.07', '-05 23 27.81', '161-328', 'i', '0')
         ('161-314', '05 35 16.10', '-05 23 14.05', '161-314', 'i', '0')
         ('162-133', '05 35 16.19', '-05 21 32.39', '162-133', 'i', '0')
         ('163-026', '05 35 16.31', '-05 20 25.24', '0', 'd', 'B')
         ('163-210', '05 35 16.27', '-05 22 10.45', '163-210', 'i', 'B')
         ('163-317', '05 35 16.27', '-05 23 16.51', '163-317', 'i', '0')
         ('163-222', '05 35 16.30', '-05 22 21.50', '163-222', 'i', 'B,FO')
         ('163-249', '05 35 16.33', '-05 22 49.01', '163-249', 'i', 'B')
         ('164-511', '05 35 16.35', '-05 25 09.60', '164-511', 'i', '0')
         ('165-235', '05 35 16.48', '-05 22 35.16', '165-235', 'i', '0')
         ('165-254', '05 35 16.54', '-05 22 53.70', '0', 'd', 'RN')
         ('166-316', '05 35 16.61', '-05 23 16.19', '166-315', 'i', '0')
         ('166-519', '05 35 16.57', '-05 25 17.74', '166-519', 'd', '0')
         ('166-406', '05 35 16.57', '-05 24 06.00', '166-406', 'i', '0')
         ('166-250', '05 35 16.59', '-05 22 50.36', '166-250', 'i', '0')
         ('167-231', '05 35 16.73', '-05 22 31.30', '167-231', 'd', 'FO')
         ('167-317', '05 35 16.74', '-05 23 16.51', '167-317', 'i', '0')
         ('168-235', '05 35 16.81', '-05 22 34.71', '168-235', 'i', '0')
         ('168-328', '05 35 16.77', '-05 23 28.06', '168-328', 'i', 'B')
         ('168-326', '05 35 16.83', '-05 23 25.91', '168-326', 'i', 'B')
         ('169-338', '05 35 16.88', '-05 23 38.10', '169-338', 'i', 'B')
         ('170-301', '05 35 16.95', '-05 23 00.91', '170-301', 'i', '0')
         ('170-249', '05 35 16.96', '-05 22 48.51', '170-249', 'i', 'B')
         ('170-337', '05 35 16.97', '-05 23 37.15', '170-337', 'i', 'B')
         ('171-340', '05 35 17.04', '-05 23 39.75', '171-340', 'i', 'B')
         ('171-434', '05 35 17.11', '-05 24 34.40', '171-434', 'i', 'B')
         ('172-028', '05 35 17.22', '-05 20 27.84', '172-028', 'd', '0')
         ('173-341', '05 35 17.32', '-05 23 41.40', '173-341', 'i', 'B')
         ('173-236', '05 35 17.34', '-05 22 35.81', '173-236', 'i', '0')
         ('174-305', '05 35 17.37', '-05 23 04.86', '174-305', 'i', '0')
         ('174-414', '05 35 17.38', '-05 24 13.90', '174-414', 'i', '0')
         ('175-251', '05 35 17.47', '-05 22 51.26', '175-251', 'i', 'B')
         ('175-355', '05 35 17.54', '-05 23 55.05', '175-355', 'i', '0')
         ('175-543', '05 35 17.54', '-05 25 42.89', '175-543', 'i', 'J,EO,B')
         ('176-325', '05 35 17.55', '-05 23 24.96', '176-325', 'i', '0')
         ('176-252', '05 35 17.64', '-05 22 51.66', '176-252', 'i', '0')
         ('177-341W', '05 35 17.66', '-05 23 41.00', '177-341', 'i', 'B')
         ('177-454', '05 35 17.69', '-05 24 54.10', '177-454', 'i', '0')
         ('177-541', '05 35 17.71', '-05 25 40.76', '0', 'i', 'EO,B')
         ('177-444', '05 35 17.73', '-05 24 43.75', '177-444', 'i', '0')
         ('177-341E', '05 35 17.73', '-05 23 41.10', '177-341', 'i', 'B')
         ('178-441', '05 35 17.81', '-05 24 41.05', '0', 'i', '0')
         ('178-258', '05 35 17.84', '-05 22 58.15', '0', 'i', '0')
         ('179-056', '05 35 17.92', '-05 20 55.44', '179-056', 'i', 'B')
         ('179-354', '05 35 17.96', '-05 23 53.50', '179-354', 'i', '0')
         ('180-331', '05 35 18.03', '-05 23 30.80', '180-331', 'i', 'B')
         ('181-247', '05 35 18.08', '-05 22 47.10', '181-247', 'i', 'B')
         ('181-825', '05 35 18.10', '-05 28 25.04', '0', 'i', 'RN')
         ('182-316', '05 35 18.19', '-05 23 31.55', '182-316', 'd', 'B')
         ('182-413', '05 35 18.22', '-05 24 13.45', '182-413', 'i', '0')
         ('183-439', '05 35 18.28', '-05 24 38.85', '183-439', 'i', 'B')
         ('183-419', '05 35 18.31', '-05 24 18.85', '183-419', 'i', '0')
         ('183-405', '05 35 18.33', '-05 24 04.85', '183-405', 'd', 'FO,B')
         ('184-427', '05 35 18.35', '-05 24 26.85', '184-427', 'i', 'B')
         ('184-520', '05 35 18.44', '-05 25 19.29', '184-520', 'i', '0')
         ('187-314', '05 35 18.68', '-05 23 14.01', '187-314', 'i', 'B')
         ('189-329', '05 35 18.87', '-05 23 28.85', '189-329', 'i', '0')
         ('190-251', '05 35 19.03', '-05 22 50.65', '0', 'i', 'B')
         ('191-232', '05 35 19.13', '-05 22 31.20', '0', 'd', '0')
         ('191-350', '05 35 19.06', '-05 23 49.50', '191-350', 'i', 'J')
         ('193-1659', '05 35 19.25', '-05 16 58.69', '0', 'rn', '0')
         ('197-427', '05 35 19.65', '-05 24 26.70', '197-427', 'i', 'RN')
         ('198-222', '05 35 19.82', '-05 22 21.55', '198-222', 'i', '0')
         ('198-448', '05 35 19.83', '-05 24 47.95', '198-448', 'i', '0')
         ('199-1508', '05 35 19.89', '-05 15 08.25', '0', 'i', 'B')
         ('200-106', '05 35 20.04', '-05 21 05.99', '200-106', 'i', '0')
         ('201-534', '05 35 20.14', '-05 25 33.84', '201-534', 'i', '0')
         ('202-228', '05 35 20.15', '-05 22 28.30', '202-228', 'i', 'B')
         ('203-504', '05 35 20.26', '-05 25 04.05', '203-504', 'i', 'B')
         ('203-506', '05 35 20.32', '-05 25 05.55', '0', 'd', 'RN')
         ('205-330', '05 35 20.45', '-05 23 29.96', '205-330', 'i', 'B')
         ('205-052', '05 35 20.52', '-05 20 52.05', '205-052', 'i', '0')
         ('205-421', '05 35 20.53', '-05 24 21.00', '205-421', 'i', 'FO')
         ('206-446', '05 35 20.62', '-05 24 46.45', '206-446', 'i', '0')
         ('208-122', '05 35 20.83', '-05 21 21.45', '208-122', 'rn', '0')
         ('209-151', '05 35 21.00', '-05 21 52.30', '209-151', 'i', 'B')
         ('210-225', '05 35 21.03', '-05 22 25.20', '0', 'i', '0')
         ('212-557', '05 35 21.15', '-05 25 57.04', '212-557', 'i', '0')
         ('212-400', '05 35 21.19', '-05 24 00.20', '0', 'i', '0')
         ('212-260', '05 35 21.24', '-05 22 59.51', '212-260', 'i', '0')
         ('213-533', '05 35 21.28', '-05 25 33.11', '0', 'i', 'B')
         ('213-346', '05 35 21.30', '-05 23 46.10', '0', 'i', 'B')
         ('215-652', '05 35 21.45', '-05 26 52.40', '0', 'i', '0')
         ('215-317', '05 35 21.49', '-05 23 16.71', '215-317', 'i', '0')
         ('215-106', '05 35 21.55', '-05 21 05.60', '215-106', 'i', 'J')
         ('216-541', '05 35 21.60', '-05 25 40.70', '0', 'i', '0')
         ('216-715', '05 35 21.62', '-05 27 14.65', '216-715', 'i', '0')
         ('218-339', '05 35 21.77', '-05 23 39.30', '218-339', 'i', '0')
         ('218-354', '05 35 21.79', '-05 23 53.90', '218-354', 'd', 'EO,B')
         ('218-529', '05 35 21.82', '-05 25 28.46', '0', 'i', 'B')
         ('218-306', '05 35 21.84', '-05 23 06.46', '0', 'i', 'B')
         ('221-433', '05 35 22.08', '-05 24 32.95', '221-433', 'i', '0')
         ('223-414', '05 35 22.31', '-05 24 14.25', '223-414', 'i', 'B')
         ('224-728', '05 35 22.37', '-05 27 28.40', '224-728', 'i', '0')
         ('228-548', '05 35 22.83', '-05 25 47.69', '228-548', 'i', '0')
         ('230-536', '05 35 23.02', '-05 25 36.29', '0', 'd', '0')
         ('231-460', '05 35 23.05', '-05 24 59.86', '231-460', 'i', '0')
         ('231-502', '05 35 23.16', '-05 25 02.19', '231-502', 'i', 'B')
         ('231-838', '05 35 23.10', '-05 28 37.34', '231-838', 'i', 'J')
         ('232-453', '05 35 23.22', '-05 24 52.79', '232-453', 'i', '0')
         ('234-853', '05 35 23.40', '-05 28 53.19', '0', 'i', '0')
         ('236-527', '05 35 23.59', '-05 25 26.54', '236-527', 'i', '0')
         ('237-627', '05 35 23.66', '-05 26 27.15', '237-627', 'i', '0')
         ('238-334', '05 35 23.80', '-05 23 34.30', '238-334', 'i', 'B')
         ('239-334', '05 35 23.86', '-05 23 34.05', '0', 'd', 'EO,B')
         ('239-510', '05 35 23.98', '-05 25 09.94', '239-510', 'i', '0')
         ('240-314', '05 35 24.02', '-05 23 13.85', '240-314', 'i', 'B')
         ('242-519', '05 35 24.22', '-05 25 18.79', '242-519', 'i', '0')
         ('244-440', '05 35 24.38', '-05 24 39.74', '244-440', 'i', '0')
         ('245-632', '05 35 24.45', '-05 26 31.55', '245-632', 'i', 'B')
         ('245-1910', '05 35 24.48', '-05 19 09.84', '0', 'i', 'RN,B')
         ('245-502', '05 35 24.51', '-05 25 01.59', '245-502', 'i', '0')
         ('247-436', '05 35 24.69', '-05 24 35.74', '247-436', 'i', 'J')
         ('250-439', '05 35 25.02', '-05 24 38.49', '250-439', 'i', '0')
         ('252-457', '05 35 25.21', '-05 24 57.34', '252-457', 'i', '0')
         ('253-1536', '05 35 25.30', '-05 15 35.54', '0', 'd', 'J,RN,B')
         ('254-412', '05 35 25.37', '-05 24 11.50', '254-412', 'i', '0')
         ('255-512', '05 35 25.52', '-05 25 11.84', '0', 'i', '0')
         ('262-521', '05 35 26.18', '-05 25 20.49', '262-521', 'i', '0')
         ('264-532', '05 35 26.42', '-05 25 31.69', '264-532', 'i', '0')
         ('266-558', '05 35 26.62', '-05 25 57.84', '0', 'i', '0')
         ('280-931', '05 35 27.96', '-05 29 31.15', '0', 'i', '0')
         ('280-1720', '05 35 28.05', '-05 17 20.33', '0', 'd', 'FO')
         ('281-306', '05 35 28.13', '-05 23 06.45', '0', 'd', 'FO')
         ('282-458', '05 35 28.20', '-05 24 58.19', '282-458', 'i', 'J')
         ('282-614', '05 35 28.20', '-05 26 14.20', '0', 'i', '0')
         ('284-439', '05 35 28.40', '-05 24 38.69', '0', 'i', '0')
         ('294-757', '05 35 29.43', '-05 37 56.60', '0', 'rn', '0')
         ('294-606', '05 35 29.48', '-05 26 06.63', '0', 'd', 'RN')
         ('297-025', '05 35 29.67', '-05 30 24.75', '0', 'i', '0')
         ('304-539', '05 35 30.41', '-05 25 38.63', '304-539', 'i', '0')
         ('307-1807', '05 35 30.70', '-05 18 07.24', '307-1807', 'i', '0')
         ('314-816', '05 35 31.40', '-05 28 16.48', '0', 'i', '0')
         ('321-602', '05 35 32.10', '-05 26 01.94', '321-602', 'd', 'EO')
         ('332-405', '05 35 33.19', '-05 24 04.74', '0', 'd', '0')
         ('332-1605', '05 35 33.20', '-05 16 05.38', '332-1605', 'i', '0')
         ('346-1553', '05 35 34.62', '-05 15 52.92', '0', 'd', 'FO')
         ('347-1535', '05 35 34.67', '-05 15 34.88', '347-1535', 'd', 'J')
         ('351-3349', '05 35 35.13', '-05 33 49.18', '0', 'i', 'J')
         ('353-130', '05 35 35.32', '-05 21 29.59', '0', 'j', 'J,B')
         ('473-245', '05 35 47.34', '-05 22 44.82', '0', 'd', 'EO')],
                     mask = [(False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)
         (False, False, False, False, False, False)],
               fill\_value = ('N/A', 'N/A', 'N/A', 'N/A', 'N/', 'N/A'),
                    dtype = [('Name', '<U8'), ('RAJ2000', '<U11'), ('DEJ2000', '<U12'), ('[OW94]', '<U8'), ('Type', '<U2'), ('Note', '<U6')])
\end{Verbatim}

    Bright proplyds have \texttt{i} in the \texttt{Type} column.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}10}]:} \PY{n}{m} \PY{o}{=} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+s}{\PYZsq{}}\PY{l+s}{i}\PY{l+s}{\PYZsq{}}
         \PY{n}{md} \PY{o}{=} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+s}{\PYZsq{}}\PY{l+s}{d}\PY{l+s}{\PYZsq{}}
         \PY{n}{mj} \PY{o}{=} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+s}{\PYZsq{}}\PY{l+s}{j}\PY{l+s}{\PYZsq{}}
         \PY{n}{mr} \PY{o}{=} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+s}{\PYZsq{}}\PY{l+s}{rn}\PY{l+s}{\PYZsq{}}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}11}]:} \PY{n}{rdata}\PY{p}{[}\PY{n}{m}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}11}]:} <Table rows=178 names=('Name','RAJ2000','DEJ2000','[OW94]','Type','Note')>
         masked\_array(data = [('4468-605', '05 34 46.76', '-05 26 04.79', '0', 'i', 'J')
          ('4596-400', '05 34 59.56', '-05 24 00.19', '4596-400', 'i', '0')
          ('4582-635', '05 34 58.16', '-05 26 35.13', '0', 'i', '0')
          ('005-514', '05 35 00.47', '-05 25 14.34', '005-514', 'i', '0')
          ('038-627', '05 35 04.19', '-05 26 27.89', '038-627', 'i', '0')
          ('044-527', '05 35 04.42', '-05 25 27.40', '044-527', 'i', '0')
          ('046-245', '05 35 04.63', '-05 22 44.85', '0', 'i', '0')
          ('049-143', '05 35 04.94', '-05 21 42.99', '0', 'i', '0')
          ('057-419', '05 35 05.73', '-05 24 18.55', '057-419', 'i', '0')
          ('061-401', '05 35 06.09', '-05 24 00.60', '061-401', 'i', '0')
          ('064-3335', '05 35 06.44', '-05 33 35.25', '0', 'i', '0')
          ('066-3251', '05 35 06.57', '-05 32 51.49', '0', 'i', '0')
          ('066-652', '05 35 06.59', '-05 26 51.99', '066-652', 'i', 'B')
          ('069-601', '05 35 06.91', '-05 26 00.60', '069-601', 'i', '0')
          ('072-135', '05 35 07.21', '-05 21 34.43', '072-135', 'i', 'EO')
          ('073-227', '05 35 07.27', '-05 22 26.56', '073-227', 'i', '0')
          ('093-822', '05 35 09.59', '-05 28 22.92', '093-822', 'i', '0')
          ('099-339', '05 35 09.89', '-05 23 38.50', '0', 'i', 'J')
          ('102-233', '05 35 10.13', '-05 22 32.74', '102-233', 'i', '0')
          ('102-021', '05 35 10.19', '-05 20 20.99', '102-021', 'i', '0')
          ('102-322', '05 35 10.20', '-05 23 21.56', '102-322', 'i', '0')
          ('106-417', '05 35 10.54', '-05 24 16.70', '106-417', 'i', '0')
          ('106-156', '05 35 10.58', '-05 21 56.24', '106-156', 'i', '0')
          ('109-246', '05 35 10.90', '-05 22 46.36', '109-246', 'i', '0')
          ('109-327', '05 35 10.95', '-05 23 26.45', '109-327', 'i', 'J')
          ('117-421', '05 35 11.65', '-05 24 21.50', '117-421', 'i', '0')
          ('117-352', '05 35 11.73', '-05 23 51.70', '0', 'i', '0')
          ('119-340', '05 35 11.90', '-05 23 39.70', '119-340', 'i', '0')
          ('121-434', '05 35 12.12', '-05 24 33.80', '121-434', 'i', '0')
          ('124-132', '05 35 12.38', '-05 21 31.39', '124-132', 'i', 'J,EO,B')
          ('131-046', '05 35 13.05', '-05 20 45.79', '131-046', 'i', '0')
          ('131-247', '05 35 13.11', '-05 22 47.11', '131-247', 'i', 'J')
          ('132-042', '05 35 13.24', '-05 20 41.94', '132-042', 'i', 'J,EO')
          ('133-353', '05 35 13.30', '-05 23 53.05', '133-353', 'i', 'B')
          ('135-220', '05 35 13.51', '-05 22 19.49', '135-220', 'i', '0')
          ('138-207', '05 35 13.78', '-05 22 07.39', '138-207', 'i', '0')
          ('139-320', '05 35 13.92', '-05 23 20.16', '0', 'i', '0')
          ('141-520', '05 35 14.05', '-05 25 20.50', '141-520', 'i', 'FO')
          ('142-301', '05 35 14.15', '-05 23 00.91', '142-301', 'i', '0')
          ('143-425', '05 35 14.27', '-05 24 24.55', '143-425', 'i', '0')
          ('144-522', '05 35 14.34', '-05 25 22.30', '144-522', 'i', '0')
          ('146-201', '05 35 14.61', '-05 22 00.94', '0', 'i', '0')
          ('147-323', '05 35 14.72', '-05 23 23.01', '0', 'i', '0')
          ('148-305', '05 35 14.80', '-05 23 04.76', '148-305', 'i', 'B')
          ('149-329', '05 35 14.92', '-05 23 29.05', '149-329', 'i', '0')
          ('150-147', '05 35 15.00', '-05 21 47.34', '0', 'i', '0')
          ('150-231', '05 35 15.02', '-05 22 31.11', '150-231', 'i', 'B')
          ('152-319', '05 35 15.20', '-05 23 18.81', '152-319', 'i', '0')
          ('152-738', '05 35 15.21', '-05 27 37.85', '152-738', 'i', '0')
          ('154-324', '05 35 15.35', '-05 23 24.11', '154-324', 'i', 'B')
          ('154-225', '05 35 15.37', '-05 22 25.35', '154-225', 'i', 'B')
          ('154-240', '05 35 15.38', '-05 22 39.85', '0', 'i', '0')
          ('155-338', '05 35 15.51', '-05 23 37.45', '155-338', 'i', '0')
          ('157-323', '05 35 15.72', '-05 23 22.59', '157-323', 'i', 'B')
          ('158-323', '05 35 15.83', '-05 23 22.59', '158-323', 'i', 'B')
          ('158-327', '05 35 15.79', '-05 23 26.51', '158-327', 'i', 'B')
          ('158-326', '05 35 15.81', '-05 23 25.51', '158-326', 'i', 'B')
          ('159-338', '05 35 15.90', '-05 23 38.00', '159-338', 'i', 'B')
          ('159-418', '05 35 15.90', '-05 24 17.85', '159-418', 'i', '0')
          ('159-350', '05 35 15.96', '-05 23 50.30', '159-350', 'i', 'B')
          ('160-353', '05 35 16.01', '-05 23 53.00', '160-353', 'i', 'J')
          ('161-324', '05 35 16.05', '-05 23 24.35', '161-324', 'i', '0')
          ('161-328', '05 35 16.07', '-05 23 27.81', '161-328', 'i', '0')
          ('161-314', '05 35 16.10', '-05 23 14.05', '161-314', 'i', '0')
          ('162-133', '05 35 16.19', '-05 21 32.39', '162-133', 'i', '0')
          ('163-210', '05 35 16.27', '-05 22 10.45', '163-210', 'i', 'B')
          ('163-317', '05 35 16.27', '-05 23 16.51', '163-317', 'i', '0')
          ('163-222', '05 35 16.30', '-05 22 21.50', '163-222', 'i', 'B,FO')
          ('163-249', '05 35 16.33', '-05 22 49.01', '163-249', 'i', 'B')
          ('164-511', '05 35 16.35', '-05 25 09.60', '164-511', 'i', '0')
          ('165-235', '05 35 16.48', '-05 22 35.16', '165-235', 'i', '0')
          ('166-316', '05 35 16.61', '-05 23 16.19', '166-315', 'i', '0')
          ('166-406', '05 35 16.57', '-05 24 06.00', '166-406', 'i', '0')
          ('166-250', '05 35 16.59', '-05 22 50.36', '166-250', 'i', '0')
          ('167-317', '05 35 16.74', '-05 23 16.51', '167-317', 'i', '0')
          ('168-235', '05 35 16.81', '-05 22 34.71', '168-235', 'i', '0')
          ('168-328', '05 35 16.77', '-05 23 28.06', '168-328', 'i', 'B')
          ('168-326', '05 35 16.83', '-05 23 25.91', '168-326', 'i', 'B')
          ('169-338', '05 35 16.88', '-05 23 38.10', '169-338', 'i', 'B')
          ('170-301', '05 35 16.95', '-05 23 00.91', '170-301', 'i', '0')
          ('170-249', '05 35 16.96', '-05 22 48.51', '170-249', 'i', 'B')
          ('170-337', '05 35 16.97', '-05 23 37.15', '170-337', 'i', 'B')
          ('171-340', '05 35 17.04', '-05 23 39.75', '171-340', 'i', 'B')
          ('171-434', '05 35 17.11', '-05 24 34.40', '171-434', 'i', 'B')
          ('173-341', '05 35 17.32', '-05 23 41.40', '173-341', 'i', 'B')
          ('173-236', '05 35 17.34', '-05 22 35.81', '173-236', 'i', '0')
          ('174-305', '05 35 17.37', '-05 23 04.86', '174-305', 'i', '0')
          ('174-414', '05 35 17.38', '-05 24 13.90', '174-414', 'i', '0')
          ('175-251', '05 35 17.47', '-05 22 51.26', '175-251', 'i', 'B')
          ('175-355', '05 35 17.54', '-05 23 55.05', '175-355', 'i', '0')
          ('175-543', '05 35 17.54', '-05 25 42.89', '175-543', 'i', 'J,EO,B')
          ('176-325', '05 35 17.55', '-05 23 24.96', '176-325', 'i', '0')
          ('176-252', '05 35 17.64', '-05 22 51.66', '176-252', 'i', '0')
          ('177-341W', '05 35 17.66', '-05 23 41.00', '177-341', 'i', 'B')
          ('177-454', '05 35 17.69', '-05 24 54.10', '177-454', 'i', '0')
          ('177-541', '05 35 17.71', '-05 25 40.76', '0', 'i', 'EO,B')
          ('177-444', '05 35 17.73', '-05 24 43.75', '177-444', 'i', '0')
          ('177-341E', '05 35 17.73', '-05 23 41.10', '177-341', 'i', 'B')
          ('178-441', '05 35 17.81', '-05 24 41.05', '0', 'i', '0')
          ('178-258', '05 35 17.84', '-05 22 58.15', '0', 'i', '0')
          ('179-056', '05 35 17.92', '-05 20 55.44', '179-056', 'i', 'B')
          ('179-354', '05 35 17.96', '-05 23 53.50', '179-354', 'i', '0')
          ('180-331', '05 35 18.03', '-05 23 30.80', '180-331', 'i', 'B')
          ('181-247', '05 35 18.08', '-05 22 47.10', '181-247', 'i', 'B')
          ('181-825', '05 35 18.10', '-05 28 25.04', '0', 'i', 'RN')
          ('182-413', '05 35 18.22', '-05 24 13.45', '182-413', 'i', '0')
          ('183-439', '05 35 18.28', '-05 24 38.85', '183-439', 'i', 'B')
          ('183-419', '05 35 18.31', '-05 24 18.85', '183-419', 'i', '0')
          ('184-427', '05 35 18.35', '-05 24 26.85', '184-427', 'i', 'B')
          ('184-520', '05 35 18.44', '-05 25 19.29', '184-520', 'i', '0')
          ('187-314', '05 35 18.68', '-05 23 14.01', '187-314', 'i', 'B')
          ('189-329', '05 35 18.87', '-05 23 28.85', '189-329', 'i', '0')
          ('190-251', '05 35 19.03', '-05 22 50.65', '0', 'i', 'B')
          ('191-350', '05 35 19.06', '-05 23 49.50', '191-350', 'i', 'J')
          ('197-427', '05 35 19.65', '-05 24 26.70', '197-427', 'i', 'RN')
          ('198-222', '05 35 19.82', '-05 22 21.55', '198-222', 'i', '0')
          ('198-448', '05 35 19.83', '-05 24 47.95', '198-448', 'i', '0')
          ('199-1508', '05 35 19.89', '-05 15 08.25', '0', 'i', 'B')
          ('200-106', '05 35 20.04', '-05 21 05.99', '200-106', 'i', '0')
          ('201-534', '05 35 20.14', '-05 25 33.84', '201-534', 'i', '0')
          ('202-228', '05 35 20.15', '-05 22 28.30', '202-228', 'i', 'B')
          ('203-504', '05 35 20.26', '-05 25 04.05', '203-504', 'i', 'B')
          ('205-330', '05 35 20.45', '-05 23 29.96', '205-330', 'i', 'B')
          ('205-052', '05 35 20.52', '-05 20 52.05', '205-052', 'i', '0')
          ('205-421', '05 35 20.53', '-05 24 21.00', '205-421', 'i', 'FO')
          ('206-446', '05 35 20.62', '-05 24 46.45', '206-446', 'i', '0')
          ('209-151', '05 35 21.00', '-05 21 52.30', '209-151', 'i', 'B')
          ('210-225', '05 35 21.03', '-05 22 25.20', '0', 'i', '0')
          ('212-557', '05 35 21.15', '-05 25 57.04', '212-557', 'i', '0')
          ('212-400', '05 35 21.19', '-05 24 00.20', '0', 'i', '0')
          ('212-260', '05 35 21.24', '-05 22 59.51', '212-260', 'i', '0')
          ('213-533', '05 35 21.28', '-05 25 33.11', '0', 'i', 'B')
          ('213-346', '05 35 21.30', '-05 23 46.10', '0', 'i', 'B')
          ('215-652', '05 35 21.45', '-05 26 52.40', '0', 'i', '0')
          ('215-317', '05 35 21.49', '-05 23 16.71', '215-317', 'i', '0')
          ('215-106', '05 35 21.55', '-05 21 05.60', '215-106', 'i', 'J')
          ('216-541', '05 35 21.60', '-05 25 40.70', '0', 'i', '0')
          ('216-715', '05 35 21.62', '-05 27 14.65', '216-715', 'i', '0')
          ('218-339', '05 35 21.77', '-05 23 39.30', '218-339', 'i', '0')
          ('218-529', '05 35 21.82', '-05 25 28.46', '0', 'i', 'B')
          ('218-306', '05 35 21.84', '-05 23 06.46', '0', 'i', 'B')
          ('221-433', '05 35 22.08', '-05 24 32.95', '221-433', 'i', '0')
          ('223-414', '05 35 22.31', '-05 24 14.25', '223-414', 'i', 'B')
          ('224-728', '05 35 22.37', '-05 27 28.40', '224-728', 'i', '0')
          ('228-548', '05 35 22.83', '-05 25 47.69', '228-548', 'i', '0')
          ('231-460', '05 35 23.05', '-05 24 59.86', '231-460', 'i', '0')
          ('231-502', '05 35 23.16', '-05 25 02.19', '231-502', 'i', 'B')
          ('231-838', '05 35 23.10', '-05 28 37.34', '231-838', 'i', 'J')
          ('232-453', '05 35 23.22', '-05 24 52.79', '232-453', 'i', '0')
          ('234-853', '05 35 23.40', '-05 28 53.19', '0', 'i', '0')
          ('236-527', '05 35 23.59', '-05 25 26.54', '236-527', 'i', '0')
          ('237-627', '05 35 23.66', '-05 26 27.15', '237-627', 'i', '0')
          ('238-334', '05 35 23.80', '-05 23 34.30', '238-334', 'i', 'B')
          ('239-510', '05 35 23.98', '-05 25 09.94', '239-510', 'i', '0')
          ('240-314', '05 35 24.02', '-05 23 13.85', '240-314', 'i', 'B')
          ('242-519', '05 35 24.22', '-05 25 18.79', '242-519', 'i', '0')
          ('244-440', '05 35 24.38', '-05 24 39.74', '244-440', 'i', '0')
          ('245-632', '05 35 24.45', '-05 26 31.55', '245-632', 'i', 'B')
          ('245-1910', '05 35 24.48', '-05 19 09.84', '0', 'i', 'RN,B')
          ('245-502', '05 35 24.51', '-05 25 01.59', '245-502', 'i', '0')
          ('247-436', '05 35 24.69', '-05 24 35.74', '247-436', 'i', 'J')
          ('250-439', '05 35 25.02', '-05 24 38.49', '250-439', 'i', '0')
          ('252-457', '05 35 25.21', '-05 24 57.34', '252-457', 'i', '0')
          ('254-412', '05 35 25.37', '-05 24 11.50', '254-412', 'i', '0')
          ('255-512', '05 35 25.52', '-05 25 11.84', '0', 'i', '0')
          ('262-521', '05 35 26.18', '-05 25 20.49', '262-521', 'i', '0')
          ('264-532', '05 35 26.42', '-05 25 31.69', '264-532', 'i', '0')
          ('266-558', '05 35 26.62', '-05 25 57.84', '0', 'i', '0')
          ('280-931', '05 35 27.96', '-05 29 31.15', '0', 'i', '0')
          ('282-458', '05 35 28.20', '-05 24 58.19', '282-458', 'i', 'J')
          ('282-614', '05 35 28.20', '-05 26 14.20', '0', 'i', '0')
          ('284-439', '05 35 28.40', '-05 24 38.69', '0', 'i', '0')
          ('297-025', '05 35 29.67', '-05 30 24.75', '0', 'i', '0')
          ('304-539', '05 35 30.41', '-05 25 38.63', '304-539', 'i', '0')
          ('307-1807', '05 35 30.70', '-05 18 07.24', '307-1807', 'i', '0')
          ('314-816', '05 35 31.40', '-05 28 16.48', '0', 'i', '0')
          ('332-1605', '05 35 33.20', '-05 16 05.38', '332-1605', 'i', '0')
          ('351-3349', '05 35 35.13', '-05 33 49.18', '0', 'i', 'J')],
                      mask = [(False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)
          (False, False, False, False, False, False)],
                fill\_value = ('N/A', 'N/A', 'N/A', 'N/A', 'N/', 'N/A'),
                     dtype = [('Name', '<U8'), ('RAJ2000', '<U11'), ('DEJ2000', '<U12'), ('[OW94]', '<U8'), ('Type', '<U2'), ('Note', '<U6')])
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}12}]:} \PY{k}{print}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Number of bright proplyds:}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n+nb}{len}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
Number of bright proplyds: 178
    \end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}13}]:} \PY{k+kn}{from} \PY{n+nn}{astropy} \PY{k+kn}{import} \PY{n}{coordinates} \PY{k}{as} \PY{n}{coord}
         \PY{k+kn}{import} \PY{n+nn}{astropy.units} \PY{k+kn}{as} \PY{n+nn}{u}
\end{Verbatim}

    Try out the new coordinates interface.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}14}]:} \PY{n}{c} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{p}{(}\PY{n}{ra}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{05 35 33.20}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{dec}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZhy{}05 16 05.38}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{hourangle}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{deg}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
         \PY{n}{c}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}14}]:} <SkyCoord (ICRS): ra=83.88833333333332 deg, dec=-5.268161111111111 deg>
\end{Verbatim}

    Add a column with an astropy \texttt{SkyCoord} instance for each object

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}15}]:} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coord}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{p}{(}\PY{n}{ra}\PY{o}{=}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{RAJ2000}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{dec}\PY{o}{=}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{DEJ2000}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{hourangle}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{deg}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    Find the coordinates of $\theta^1$ Ori C:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}16}]:} \PY{n}{c0} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{get\PYZus{}icrs\PYZus{}coordinates}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{tet01 ori c}\PY{l+s}{\PYZdq{}}\PY{p}{)}
         \PY{n}{c0}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}16}]:} <SkyCoord (ICRS): ra=83.818598977 deg, dec=-5.389680154 deg>
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}17}]:} \PY{n}{c0}\PY{o}{.}\PY{n}{separation}\PY{p}{(}\PY{n}{c}\PY{p}{)}\PY{o}{.}\PY{n}{arcsec}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}17}]:} 503.84327222379954
\end{Verbatim}

    Add another column with the projected separation in arcsec:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}18}]:} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{c0}\PY{o}{.}\PY{n}{separation}\PY{p}{(}\PY{n}{c}\PY{p}{)}\PY{o}{.}\PY{n}{arcsec} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coord}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

    Make a cumulative histogram to compare with the HA98 sample:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}19}]:} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}\PY{p}{,} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{o}{.}\PY{n}{max}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}19}]:} (685.54604284372397, 932.10782801757205)
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}257}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{800.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{4}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Henney \PYZam{} Arthur 1998}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{md}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dark disks}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{mr}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{g}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Reflec nebs}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{mj}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{m}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Bare jets}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of proplyds with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Cumulative distribution of proplyds vs distance: Both samples}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{savefig}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{proplyd\PYZhy{}distance\PYZhy{}distribution.pdf}\PY{l+s}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_47_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    Try it again, but non-cumulative. If the cumulative distribution is
linear in $D'$ then the non-cumulative one should be constant.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}256}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{800.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{20}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Henney \PYZam{} Arthur 1998}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{md}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dark disks}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{mr}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{g}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Reflec nebs}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{mj}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{m}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Bare jets}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of proplyds with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Distribution of proplyds vs distance: Both samples}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{12}\PY{p}{,} \PY{l+m+mi}{6}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_49_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    The ultimate tool for plotting a distribution is \texttt{distplot}. This
can show the following if we turn on all the options: 1. A rugplot of
the individual data points 2. A histogram of the distribution 3. The KDE
of the distribution 4. A fit to the distribution of any function from
\texttt{scipy.stats}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}387}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} 
                       \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} 
                       \PY{n}{fit}\PY{o}{=}\PY{n}{stats}\PY{o}{.}\PY{n}{expon}\PY{p}{,} 
                       \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{clip}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{800.0}\PY{p}{)}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{kernel}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{gau}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Gaussian KDE}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{rug\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{alpha}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+m+mf}{0.3}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{fit\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Exponential fit}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{800}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{25}\PY{p}{,}
                       \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Histogram (25 arcsec bins)}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{axlabel}\PY{o}{=}\PY{l+s}{r\PYZdq{}}\PY{l+s}{Projected distance, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{800.0}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Distribution of projected distances for all bright proplyds from Ricci et al. (2008) sample}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n+nb+bp}{None}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_51_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    What I don't like about this is that the KDE is affected by the fact
that there are no points with $D' < 0$. That is why it drops towards
zero. To fix this, we could reflect the dataset about zero and add a
negative mirrored dataset.

Note that \texttt{stats.laplace} is actually a 2-sided explonential. And
it doesn't fit very well. On the other hand, the KDE now looks much
nicer.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}546}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{r\PYZus{}}\PY{p}{[}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{o}{\PYZhy{}}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{]}\PY{p}{,} 
                       \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} 
                       \PY{n}{fit}\PY{o}{=}\PY{n}{stats}\PY{o}{.}\PY{n}{logistic}\PY{p}{,} 
                       \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{kernel}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{gau}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{KDE}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{rug\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{alpha}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+m+mf}{0.3}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{fit\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Sech\PYZhy{}squared fit}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{2}\PY{o}{*}\PY{l+m+mi}{800}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{25}\PY{p}{,}
                       \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Histogram (25 arcsec bins)}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{axlabel}\PY{o}{=}\PY{l+s}{r\PYZdq{}}\PY{l+s}{Projected distance, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{800.0}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Distribution of projected distances for all bright proplyds from Ricci et al. (2008) sample}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n+nb+bp}{None}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_53_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    So now, we will try the same trick on the joint distribution of $D'$ and
$r_0$. This is in a previous section.

    Zooming in on the central part, which is relevant to the bowshocks.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}531}]:} \PY{n}{Dmax} \PY{o}{=} \PY{l+m+mf}{110.0}
          \PY{n}{Dmaxx} \PY{o}{=} \PY{l+m+mf}{800.0}
          \PY{n}{mzoom} \PY{o}{=} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{\PYZlt{}} \PY{n}{Dmax}
          \PY{n}{dataset} \PY{o}{=} \PY{n}{mirrored}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m} \PY{o}{\PYZam{}} \PY{n}{mzoom}\PY{p}{]}\PY{p}{)}\PY{p}{)}
          \PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{(}\PY{n}{dataset}\PY{p}{,}
                       \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} 
                       \PY{n}{fit}\PY{o}{=}\PY{n}{stats}\PY{o}{.}\PY{n}{uniform}\PY{p}{,} 
                       \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{kernel}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{gau}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Gaussian KDE}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{rug\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{alpha}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+m+mf}{0.3}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{fit\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Constant fit}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{hist\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{range}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{p}{(}\PY{o}{\PYZhy{}}\PY{n}{Dmaxx}\PY{p}{,} \PY{n}{Dmaxx}\PY{p}{)}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{Dmaxx}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{10}\PY{p}{,}
                       \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Histogram (10 arcsec bins)}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{axlabel}\PY{o}{=}\PY{l+s}{r\PYZdq{}}\PY{l+s}{Projected distance, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{100.0}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Distribution of projected distances for all bright proplyds from Ricci et al. (2008) sample}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_56_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    So this suggests that at scales $< 100''$ the proplyd distribution is
uniform (implying density $\sim D^{-2}$.

Whereas at larger scales it does fall off.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}532}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{kdeplot}\PY{err}{?}
\end{Verbatim}

    \subsubsection{What about the stars?}\label{what-about-the-stars}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}21}]:} \PY{n}{sdata} \PY{o}{=} \PY{n}{Table}\PY{o}{.}\PY{n}{read}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd\PYZhy{}Datasets/Robberto2013/table5.dat}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{format}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{ascii}\PY{l+s}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}22}]:} \PY{n}{sdata}\PY{o}{.}\PY{n}{colnames}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}22}]:} ['Seq',
          'oncacs',
          'RAh',
          'RAm',
          'RAs',
          'DE-',
          'DEd',
          'DEm',
          'DEs',
          'visit',
          'CCD',
          'x435',
          'x555',
          'x658',
          'x775',
          'x850',
          'y435',
          'y555',
          'y658',
          'y775',
          'y850',
          'F435mag',
          'e\_F435mag',
          'f\_F435mag',
          'F555mag',
          'e\_F555mag',
          'f\_F555mag',
          'F658mag',
          'e\_F658mag',
          'f\_F658mag',
          'F775mag',
          'e\_F775mag',
          'f\_F775mag',
          'F850mag',
          'e\_F850mag',
          'f\_F850mag',
          'rad435',
          'rad555',
          'rad658',
          'rad775',
          'rad850',
          'csky435',
          'csky555',
          'csky658',
          'csky775',
          'csky850',
          'e\_csky435',
          'e\_csky555',
          'e\_csky658',
          'e\_csky775',
          'e\_csky850',
          'isky435',
          'isky555',
          'isky658',
          'isky775',
          'isky850',
          'osky435',
          'osky555',
          'osky658',
          'osky775',
          'osky850',
          'max435',
          'max555',
          'max658',
          'max775',
          'max850',
          'pixsat435',
          'pixsat555',
          'pixsat658',
          'pixsat775',
          'pixsat850',
          'radsat435',
          'radsat555',
          'radsat658',
          'radsat775',
          'radsat850',
          'type',
          'strip',
          'xstrip',
          'ystrip',
          'F435smag',
          'F555smag',
          'F658smag',
          'F775smag',
          'F850smag',
          'e\_F435smag',
          'e\_F555smag',
          'e\_F658smag',
          'e\_F775smag',
          'e\_F850smag',
          'rad435s',
          'rad555s',
          'rad658s',
          'rad775s',
          'rad850s',
          'sky435s',
          'sky555s',
          'sky658s',
          'sky775s',
          'sky850s',
          'dsky435s',
          'dksy555s',
          'dsky658s',
          'dksy775s',
          'dsky850s',
          'Date',
          'time']
\end{Verbatim}

    We first need to restrict the table to unique sources, since it contains
multiple observations of the same source. We can do this by comparing
the first two columns:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}23}]:} \PY{n}{m\PYZus{}uniq} \PY{o}{=} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Seq}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{oncacs}\PY{l+s}{\PYZsq{}}\PY{p}{]}
\end{Verbatim}

    Then winnowing down the table:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}24}]:} \PY{n}{sdata} \PY{o}{=} \PY{n}{sdata}\PY{p}{[}\PY{n}{m\PYZus{}uniq}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}24}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}24}]:} 
\end{Verbatim}

    Description of the \texttt{type} column from notes to Table 5:

\begin{verbatim}
Source type, from visual inspection: (0) not measured; (1) detected in at least one filter and unresolved; (2) double (companion closer than ≈3 pixels, one entry for both sources); (3) wide double (companion further than ≈3 pixel, one entry for each source); (5) silhouette disk; (6) photoionized (proplyd or with other evidence of photoionization); (7) galaxy; (8) Herbig-Haro.
\end{verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}25}]:} \PY{n}{source\PYZus{}types} \PY{o}{=} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{not measured}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{single}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{close double}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{wide double}\PY{l+s}{\PYZsq{}}\PY{p}{,} 
                         \PY{l+s}{\PYZsq{}}\PY{l+s}{???}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{disk}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{ionized}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{galaxy}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{herbig haro}\PY{l+s}{\PYZsq{}}\PY{p}{]}
         \PY{n}{count\PYZus{}table} \PY{o}{=} \PY{n}{Table}\PY{p}{(}\PY{n}{names}\PY{o}{=}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Code}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{Type}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{Number}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{,} 
                             \PY{n}{dtype}\PY{o}{=}\PY{p}{[}\PY{n+nb}{int}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{S12}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n+nb}{int}\PY{p}{]}\PY{p}{)}
         \PY{k}{for} \PY{n}{itype}\PY{p}{,} \PY{n}{label} \PY{o+ow}{in} \PY{n+nb}{enumerate}\PY{p}{(}\PY{n}{source\PYZus{}types}\PY{p}{)}\PY{p}{:}
             \PY{n}{count\PYZus{}table}\PY{o}{.}\PY{n}{add\PYZus{}row}\PY{p}{(}\PY{p}{[}\PY{n}{itype}\PY{p}{,} \PY{n}{label}\PY{p}{,} \PY{n}{np}\PY{o}{.}\PY{n}{sum}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{n}{itype}\PY{p}{)}\PY{p}{]}\PY{p}{)}
         \PY{n}{count\PYZus{}table}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}25}]:} <Table rows=9 names=('Code','Type','Number')>
         array([(0, b'not measured', 2), (1, b'single', 2912),
                (2, b'close double', 47), (3, b'wide double', 154), (4, b'???', 0),
                (5, b'disk', 36), (6, b'ionized', 184), (7, b'galaxy', 59),
                (8, b'herbig haro', 5)], 
               dtype=[('Code', '<i8'), ('Type', 'S12'), ('Number', '<i8')])
\end{Verbatim}

    The \texttt{ionized} type is ``\emph{proplyd or with other evidence of
photoionization}'' and there are 184 of them, compared with 178 in the
Ricci catalog. This is close enough for me.

    Next job is to winnow out types 0, 7, and 8, which we don't want.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}26}]:} \PY{n}{sdata} \PY{o}{=} \PY{n}{sdata}\PY{p}{[}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{\PYZgt{}} \PY{l+m+mi}{0}\PY{p}{)} \PY{o}{\PYZam{}} \PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{\PYZlt{}} \PY{l+m+mi}{7}\PY{p}{)}\PY{p}{]}
\end{Verbatim}

    And now create a mask to select out the \texttt{ionized} sources (mainly
proplyds).

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}27}]:} \PY{n}{m\PYZus{}rob\PYZus{}ionized} \PY{o}{=} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{type}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+m+mi}{6}
\end{Verbatim}

    Now we combine the 6 columns used for coordinates into a single RA-Dec
string:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}28}]:} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coordstring}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZob{}\PYZcb{} \PYZob{}\PYZcb{} \PYZob{}\PYZcb{} \PYZhy{}\PYZob{}\PYZcb{} \PYZob{}\PYZcb{} \PYZob{}\PYZcb{}}\PY{l+s}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{o}{*}\PY{n}{args}\PY{p}{)}
                           \PY{k}{for} \PY{n}{args} \PY{o+ow}{in} \PY{n+nb}{zip}\PY{p}{(}\PY{o}{*}\PY{p}{[}\PY{n}{sdata}\PY{p}{[}\PY{n}{col}\PY{p}{]} \PY{k}{for} \PY{n}{col} \PY{o+ow}{in} \PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{RAh}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{RAm}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{RAs}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{DEd}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{DEm}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{DEs}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{]}\PY{p}{)}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}29}]:} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coordstring}\PY{l+s}{\PYZdq{}}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}29}]:} <MaskedColumn name='coordstring' unit=None format=None description=None>
         masked\_BaseColumn(data = ['5 34 9.889 -5 22 26.26' '5 34 10.25 -5 21 52.29'
          '5 34 11.111 -5 22 54.46' \ldots, '5 35 52.472 -5 38 7.85'
          '5 35 53.093 -5 30 10.38' '5 35 53.504 -5 39 32.88'],
                           mask = [False False False \ldots, False False False],
                     fill\_value = N/A)
\end{Verbatim}

    Then that string can be used to initialize an
\texttt{astropy.coordinates.SkyCoord} instance for each source.

\textbf{Warning:} The following cell takes a couple of minutes to run.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}30}]:} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coord}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coordstring}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{hourangle}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{deg}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}31}]:} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coord}\PY{l+s}{\PYZdq{}}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}31}]:} <MaskedColumn name='coord' unit=None format=None description=None>
         masked\_BaseColumn(data = [<SkyCoord (ICRS): ra=83.54120416666666 deg, dec=-5.373961111111111 deg>
          <SkyCoord (ICRS): ra=83.54270833333332 deg, dec=-5.3645249999999995 deg>
          <SkyCoord (ICRS): ra=83.54629583333332 deg, dec=-5.381794444444444 deg>
          \ldots,
          <SkyCoord (ICRS): ra=83.96863333333332 deg, dec=-5.635513888888888 deg>
          <SkyCoord (ICRS): ra=83.97122083333332 deg, dec=-5.502883333333333 deg>
          <SkyCoord (ICRS): ra=83.97293333333332 deg, dec=-5.659133333333334 deg>],
                           mask = [False False False \ldots, False False False],
                     fill\_value = ?)
\end{Verbatim}

    And that in turn can be used to find the separation of each source from
$\theta^1$ C.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}32}]:} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{c0}\PY{o}{.}\PY{n}{separation}\PY{p}{(}\PY{n}{c}\PY{p}{)}\PY{o}{.}\PY{n}{arcsec} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{coord}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}33}]:} \PY{n+nb}{len}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}33}]:} 3333
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}258}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{1200.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{4}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ionized, Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Proplyds, Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of stars with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Cumulative distribution of stars vs distance: Robberto ACS sample}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_84_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}259}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{1200.0}
          \PY{n}{binsize} \PY{o}{=} \PY{l+m+mi}{25}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{n}{binsize}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ionized, Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Proplyds, Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of stars with projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ (25 arcsec bins)}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Distribution of stars vs distance: Robberto ACS sample}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_85_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}554}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{1200.0}
          \PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{(}\PY{n}{mirrored}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{)}\PY{p}{,} 
                       \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} 
                       \PY{n}{fit}\PY{o}{=}\PY{n}{stats}\PY{o}{.}\PY{n}{semicircular}\PY{p}{,} 
                       \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{kernel}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{gau}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{KDE}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{rug\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{alpha}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+m+mf}{0.03}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{fit\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Semi\PYZhy{}circular fit}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{25}\PY{p}{,}
                       \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Histogram (25 arcsec bins)}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{axlabel}\PY{o}{=}\PY{l+s}{r\PYZdq{}}\PY{l+s}{Projected distance, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{1200.0}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Distribution of projected distances for all stars from Robberto et al. (2013) sample}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n+nb+bp}{None}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_86_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    Zoom in on inner 200''

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}260}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{200.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ionized, Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Proplyds, Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of stars with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Cumulative distribution of stars vs distance: Robberto ACS sample}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_88_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    Zoom in on inner 30''

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}261}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{30.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ionized, Robberto 2013}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.7}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Proplyds, Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of stars with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Cumulative distribution of stars vs distance: Robberto ACS sample}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_90_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsection{Maps of different source
samples}\label{maps-of-different-source-samples}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}51}]:} \PY{n}{rob\PYZus{}ra} \PY{o}{=} \PY{p}{[}\PY{n}{c}\PY{o}{.}\PY{n}{ra}\PY{o}{.}\PY{n}{deg} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{coord}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}
         \PY{n}{rob\PYZus{}dec} \PY{o}{=} \PY{p}{[}\PY{n}{c}\PY{o}{.}\PY{n}{dec}\PY{o}{.}\PY{n}{deg} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{sdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{coord}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}56}]:} \PY{n}{rob\PYZus{}ra} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{rob\PYZus{}ra}\PY{p}{)}
         \PY{n}{rob\PYZus{}dec} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{n}{rob\PYZus{}dec}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}74}]:} \PY{n}{ricci\PYZus{}ra} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{p}{[}\PY{n}{c}\PY{o}{.}\PY{n}{ra}\PY{o}{.}\PY{n}{deg} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{coord}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}\PY{p}{)}
         \PY{n}{ricci\PYZus{}dec} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{array}\PY{p}{(}\PY{p}{[}\PY{n}{c}\PY{o}{.}\PY{n}{dec}\PY{o}{.}\PY{n}{deg} \PY{k}{for} \PY{n}{c} \PY{o+ow}{in} \PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{coord}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}38}]:} \PY{k+kn}{import} \PY{n+nn}{aplpy}
         \PY{k+kn}{import} \PY{n+nn}{os}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}41}]:} \PY{k+kn}{from} \PY{n+nn}{imp} \PY{k+kn}{import} \PY{n+nb}{reload}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}39}]:} \PY{n}{aplpy}\PY{o}{.}\PY{n}{\PYZus{}\PYZus{}version\PYZus{}\PYZus{}}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}39}]:} '0.9.13.dev310'
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}40}]:} \PY{n}{hstdir} \PY{o}{=} \PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{expanduser}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZti{}/Dropbox/JorgeBowshocks/HST}\PY{l+s}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}105}]:} \PY{n}{fig} \PY{o}{=} \PY{n}{aplpy}\PY{o}{.}\PY{n}{FITSFigure}\PY{p}{(}\PY{n}{os}\PY{o}{.}\PY{n}{path}\PY{o}{.}\PY{n}{join}\PY{p}{(}\PY{n}{hstdir}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{mosaicf656\PYZhy{}fixw\PYZhy{}align.fits}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
          \PY{n}{fig}\PY{o}{.}\PY{n}{show\PYZus{}grayscale}\PY{p}{(}\PY{n}{vmin}\PY{o}{=}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{n}{vmax}\PY{o}{=}\PY{l+m+mf}{5.e3}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
WARNING: Cannot determine equinox. Assuming J2000. [aplpy.wcs\_util]
WARNING:astropy:Cannot determine equinox. Assuming J2000.
WARNING: Cannot determine equinox. Assuming J2000. [aplpy.wcs\_util]
WARNING:astropy:Cannot determine equinox. Assuming J2000.
    \end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_99_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}108}]:} \PY{n}{fig}\PY{o}{.}\PY{n}{\PYZus{}figure}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{18}\PY{p}{)}
          \PY{n}{fig}\PY{o}{.}\PY{n}{show\PYZus{}markers}\PY{p}{(}\PY{n}{ricci\PYZus{}ra}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{ricci\PYZus{}dec}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{layer}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ricci proplyds}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{edgecolor}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{c}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{lw}\PY{o}{=}\PY{l+m+mi}{3}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.8}\PY{p}{)}
          \PY{n}{fig}\PY{o}{.}\PY{n}{show\PYZus{}markers}\PY{p}{(}\PY{n}{rob\PYZus{}ra}\PY{p}{[}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{rob\PYZus{}dec}\PY{p}{[}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{layer}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Robb proplyds}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{edgecolor}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{orange}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{fig}\PY{o}{.}\PY{n}{show\PYZus{}markers}\PY{p}{(}\PY{n}{rob\PYZus{}ra}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{rob\PYZus{}dec}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{m\PYZus{}rob\PYZus{}ionized}\PY{p}{]}\PY{p}{,} \PY{n}{layer}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Robb non\PYZhy{}proplyds}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{edgecolor}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{*}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{fig}\PY{o}{.}\PY{n}{recenter}\PY{p}{(}\PY{n}{c0}\PY{o}{.}\PY{n}{ra}\PY{o}{.}\PY{n}{deg}\PY{p}{,} \PY{n}{c0}\PY{o}{.}\PY{n}{dec}\PY{o}{.}\PY{n}{deg}\PY{p}{,} \PY{n}{radius}\PY{o}{=}\PY{l+m+mi}{60}\PY{o}{/}\PY{l+m+mi}{3600}\PY{p}{)}
          \PY{n}{ax} \PY{o}{=} \PY{n}{fig}\PY{o}{.}\PY{n}{\PYZus{}figure}\PY{o}{.}\PY{n}{axes}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}
          \PY{n}{ax}\PY{o}{.}\PY{n}{set\PYZus{}title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{A title}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{fig}\PY{o}{.}\PY{n}{\PYZus{}figure}
\end{Verbatim}
\texttt{\color{outcolor}Out[{\color{outcolor}108}]:}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_100_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    So the conclusion of this section is that the Ricci proplyd catalog
(shown by blue circles on the image above) is more reliable than the
Robberto \texttt{ionized} catalog (shown by orange circles) for the
close-in proplyds. Robberto is missing a lot of obvious proplyds, and
furthermore has a few sources that don't seem to correspond to anything,
although some might be bowshocks or jet knots.

The Ricci catalog is only missing a very few, such as the one very close
to th1c.

    \subsection{Monte Carlo simulations}\label{monte-carlo-simulations}

Assume power law distribution in radius.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}562}]:} \PY{k}{class} \PY{n+nc}{SourcePopulation}\PY{p}{(}\PY{n+nb}{object}\PY{p}{)}\PY{p}{:}

              \PY{k}{def} \PY{n+nf}{\PYZus{}\PYZus{}init\PYZus{}\PYZus{}}\PY{p}{(}\PY{n+nb+bp}{self}\PY{p}{,} \PY{n}{N}\PY{p}{,} \PY{n}{Dmin}\PY{o}{=}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{n}{Dmax}\PY{o}{=}\PY{l+m+mf}{800.0}\PY{p}{,} \PY{n}{m}\PY{o}{=}\PY{l+m+mf}{2.6}\PY{p}{)}\PY{p}{:}
                  \PY{l+s+sd}{\PYZdq{}\PYZdq{}\PYZdq{}Return `N` samples of sources drawn from a radial distribution with density \PYZti{} D\PYZca{}\PYZhy{}m}
          \PY{l+s+sd}{        Maximum and minimum radii can be specified with `Dmin` and `Dmax`. }
          \PY{l+s+sd}{        If `Dmin = 0` then `m \PYZlt{} 3` is required.\PYZdq{}\PYZdq{}\PYZdq{}}
                  \PY{k}{assert}\PY{p}{(}\PY{n}{m} \PY{o}{\PYZlt{}} \PY{l+m+mi}{3}\PY{p}{)}

                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{N} \PY{o}{=} \PY{n}{N}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{m} \PY{o}{=} \PY{n}{m}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{Dmin} \PY{o}{=} \PY{n}{Dmin}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{Dmax} \PY{o}{=} \PY{n}{Dmax}

                  \PY{c}{\PYZsh{} Radial distances from center}
                  \PY{n}{xi} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{random\PYZus{}sample}\PY{p}{(}\PY{p}{(}\PY{n}{N}\PY{p}{,}\PY{p}{)}\PY{p}{)}
                  \PY{n}{x0} \PY{o}{=} \PY{n}{Dmin}\PY{o}{/}\PY{n}{Dmax}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{D} \PY{o}{=} \PY{n}{Dmax}\PY{o}{*}\PY{p}{(}\PY{p}{(}\PY{l+m+mi}{1} \PY{o}{\PYZhy{}} \PY{n}{x0}\PY{o}{*}\PY{o}{*}\PY{p}{(}\PY{l+m+mf}{3.0} \PY{o}{\PYZhy{}} \PY{n}{m}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{n}{xi} \PY{o}{+} \PY{n}{x0}\PY{o}{*}\PY{o}{*}\PY{p}{(}\PY{l+m+mf}{3.0} \PY{o}{\PYZhy{}} \PY{n}{m}\PY{p}{)}\PY{p}{)}\PY{o}{*}\PY{o}{*}\PY{p}{(}\PY{l+m+mf}{1.0}\PY{o}{/}\PY{p}{(}\PY{l+m+mf}{3.0} \PY{o}{\PYZhy{}} \PY{n}{m}\PY{p}{)}\PY{p}{)}
                  \PY{c}{\PYZsh{} Angle cosine with LOS}
                  \PY{n}{xi} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{random\PYZus{}sample}\PY{p}{(}\PY{p}{(}\PY{n}{N}\PY{p}{,}\PY{p}{)}\PY{p}{)}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mu} \PY{o}{=} \PY{l+m+mi}{2}\PY{o}{*}\PY{n}{xi} \PY{o}{\PYZhy{}} \PY{l+m+mf}{1.0}
                  \PY{c}{\PYZsh{} Position angle}
                  \PY{n}{xi} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{random\PYZus{}sample}\PY{p}{(}\PY{p}{(}\PY{n}{N}\PY{p}{,}\PY{p}{)}\PY{p}{)}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{PA} \PY{o}{=} \PY{l+m+mf}{360.0}\PY{o}{*}\PY{n}{xi}

                  \PY{c}{\PYZsh{} Inclination to plane of sky inc = [\PYZhy{}90:90]}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{inc} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{degrees}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{arcsin}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mu}\PY{p}{)}\PY{p}{)}

                  \PY{c}{\PYZsh{} Projected distance}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{Dprime} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{D}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{sqrt}\PY{p}{(}\PY{l+m+mf}{1.0} \PY{o}{\PYZhy{}} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{mu}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{)}

                  \PY{c}{\PYZsh{} Plane\PYZhy{}of\PYZhy{}sky coords}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{Dec} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{Dprime}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{radians}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{PA}\PY{p}{)}\PY{p}{)}
                  \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{RA} \PY{o}{=} \PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{Dprime}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{sin}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{radians}\PY{p}{(}\PY{n+nb+bp}{self}\PY{o}{.}\PY{n}{PA}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}662}]:} \PY{n}{p} \PY{o}{=} \PY{n}{SourcePopulation}\PY{p}{(}\PY{l+m+mi}{150}\PY{p}{,} \PY{n}{m}\PY{o}{=}\PY{l+m+mf}{2.1}\PY{p}{,} \PY{n}{Dmin}\PY{o}{=}\PY{l+m+mf}{5.0}\PY{p}{,} \PY{n}{Dmax}\PY{o}{=}\PY{l+m+mf}{250.0}\PY{p}{)}
          \PY{n}{p2} \PY{o}{=} \PY{n}{SourcePopulation}\PY{p}{(}\PY{l+m+mi}{30}\PY{p}{,} \PY{n}{m}\PY{o}{=}\PY{l+m+mf}{2.5}\PY{p}{,} \PY{n}{Dmin}\PY{o}{=}\PY{l+m+mf}{150.0}\PY{p}{,} \PY{n}{Dmax}\PY{o}{=}\PY{l+m+mf}{800.0}\PY{p}{)}
          \PY{n}{Dprimes} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{r\PYZus{}}\PY{p}{[}\PY{n}{p}\PY{o}{.}\PY{n}{Dprime}\PY{p}{,} \PY{n}{p2}\PY{o}{.}\PY{n}{Dprime}\PY{p}{]}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}663}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{800.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{4}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{Dprimes}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{Fake sample}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{c}{\PYZsh{} \PYZus{} = plt.hist(p.Dprime, bins=nbins, range=(0, rmax), cumulative=True, color=\PYZsq{}g\PYZsq{}, alpha=0.3, label=\PYZsq{}p\PYZsq{})}
          \PY{c}{\PYZsh{} \PYZus{} = plt.hist(p2.Dprime, bins=nbins, range=(0, rmax), cumulative=True, color=\PYZsq{}y\PYZsq{}, alpha=0.8, label=\PYZsq{}p2\PYZsq{})}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of proplyds with projected distance \PYZlt{} D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Cumulative distribution of proplyds vs distance: Both samples}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_105_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}250}]:} \PY{k+kn}{import} \PY{n+nn}{seaborn}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}251}]:} \PY{n}{cmap} \PY{o}{=} \PY{n}{seaborn}\PY{o}{.}\PY{n}{palettes}\PY{o}{.}\PY{n}{dark\PYZus{}palette}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{red}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{as\PYZus{}cmap}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}664}]:} \PY{n}{rmax} \PY{o}{=} \PY{l+m+mf}{800.0}
          \PY{n}{nbins} \PY{o}{=} \PY{n}{rmax}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{25}
          \PY{n}{datasets} \PY{o}{=} \PY{p}{[}\PY{n}{rdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{[}\PY{n}{m}\PY{p}{]}\PY{p}{,} \PY{n}{Dprimes}\PY{p}{,} \PY{n}{p}\PY{o}{.}\PY{n}{Dprime}\PY{p}{,} \PY{n}{p2}\PY{o}{.}\PY{n}{Dprime}\PY{p}{]}
          \PY{n}{colors} \PY{o}{=} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{b}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{g}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{y}\PY{l+s}{\PYZsq{}}\PY{p}{]}
          \PY{n}{labels} \PY{o}{=} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Ricci 2008}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{Fake sample}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{p}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{p2}\PY{l+s}{\PYZsq{}}\PY{p}{]}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{hist}\PY{p}{(}\PY{n}{datasets}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{n}{nbins}\PY{p}{,} \PY{n+nb}{range}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{n}{rmax}\PY{p}{)}\PY{p}{,} \PY{n}{cumulative}\PY{o}{=}\PY{n+nb+bp}{False}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n+nb+bp}{None}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{n}{labels}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{3}\PY{p}{]}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{xlabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Projected distance D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{, arcsec}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{ylabel}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZsh{} of proplyds with projected distance = D}\PY{l+s+se}{\PYZbs{}\PYZsq{}}\PY{l+s}{ (bin width 25 arcsec)}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{\PYZus{}} \PY{o}{=} \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Distribution of proplyds vs distance: Both samples}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{gcf}\PY{p}{(}\PY{p}{)}\PY{o}{.}\PY{n}{set\PYZus{}size\PYZus{}inches}\PY{p}{(}\PY{l+m+mi}{18}\PY{p}{,} \PY{l+m+mi}{8}\PY{p}{)}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_108_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}665}]:} \PY{n}{Dmax} \PY{o}{=} \PY{l+m+mf}{110.0}
          \PY{n}{Dmaxx} \PY{o}{=} \PY{l+m+mf}{800.0}
          \PY{n}{mzoom} \PY{o}{=} \PY{n}{Dprimes} \PY{o}{\PYZlt{}} \PY{n}{Dmax}
          \PY{n}{dataset} \PY{o}{=} \PY{n}{mirrored}\PY{p}{(}\PY{n}{Dprimes}\PY{p}{[}\PY{n}{mzoom}\PY{p}{]}\PY{p}{)}
          \PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{(}\PY{n}{dataset}\PY{p}{,}
                       \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} 
                       \PY{n}{fit}\PY{o}{=}\PY{n}{stats}\PY{o}{.}\PY{n}{uniform}\PY{p}{,} 
                       \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{kernel}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{gau}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Gaussian KDE}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,} 
                       \PY{n}{rug\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{alpha}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+m+mf}{0.3}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{fit\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{label}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{l+s}{\PYZdq{}}\PY{l+s}{Constant fit}\PY{l+s}{\PYZdq{}}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{hist\PYZus{}kws}\PY{o}{=}\PY{p}{\PYZob{}}\PY{l+s}{\PYZdq{}}\PY{l+s}{range}\PY{l+s}{\PYZdq{}}\PY{p}{:} \PY{p}{(}\PY{o}{\PYZhy{}}\PY{n}{Dmaxx}\PY{p}{,} \PY{n}{Dmaxx}\PY{p}{)}\PY{p}{\PYZcb{}}\PY{p}{,}
                       \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{2}\PY{o}{*}\PY{n}{Dmaxx}\PY{o}{/}\PY{o}{/}\PY{l+m+mi}{10}\PY{p}{,}
                       \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Histogram (10 arcsec bins)}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{axlabel}\PY{o}{=}\PY{l+s}{r\PYZdq{}}\PY{l+s}{Projected distance, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{100.0}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Distribution of projected distances from Monta Carlo fake sample}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_109_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Compare the projected separation and the inclination for
the inner 30 arcsec of the MC
sample}\label{compare-the-projected-separation-and-the-inclination-for-the-inner-30-arcsec-of-the-mc-sample}

Assume that the inner wind shock lies at about $D = 26''$. Only proplyds
that are physically clser than this will have Type I bowshocks. These
are shown by orange points in the figure. Physically more distant
proplyds are shown by gray points. We also show contours of physical
separation every $10''$ from $D = 10''$ to $D = 200''$.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}674}]:} \PY{n}{color} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{.5}\PY{l+s}{\PYZdq{}}
          \PY{n}{m30} \PY{o}{=} \PY{n}{p}\PY{o}{.}\PY{n}{Dprime} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{l+m+mf}{30.0}
          \PY{n}{Dmax} \PY{o}{=} \PY{l+m+mf}{26.0}
          \PY{n}{mwind} \PY{o}{=} \PY{n}{p}\PY{o}{.}\PY{n}{D} \PY{o}{\PYZlt{}}\PY{o}{=} \PY{n}{Dmax}
          \PY{n}{g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{JointGrid}\PY{p}{(}\PY{n}{mirrored}\PY{p}{(}\PY{n}{p}\PY{o}{.}\PY{n}{Dprime}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{)}\PY{p}{,} \PY{n}{np}\PY{o}{.}\PY{n}{r\PYZus{}}\PY{p}{[}\PY{n}{p}\PY{o}{.}\PY{n}{inc}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{,} \PY{n}{p}\PY{o}{.}\PY{n}{inc}\PY{p}{[}\PY{n}{m30}\PY{p}{]}\PY{p}{]}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{9}\PY{p}{,} \PY{n}{xlim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{30}\PY{p}{)}\PY{p}{,} \PY{n}{ylim}\PY{o}{=}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mi}{90}\PY{p}{,} \PY{l+m+mi}{90}\PY{p}{)}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{plot\PYZus{}marginals}\PY{p}{(}\PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{,} \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} 
                           \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{n+nb}{dict}\PY{p}{(}\PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{KDE}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{bw}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{silverman}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{cut}\PY{o}{=}\PY{l+m+mi}{0}\PY{p}{)}\PY{p}{)}
          \PY{c}{\PYZsh{}g.plot\PYZus{}joint(sns.regplot, color=color, fit\PYZus{}reg=False, scatter\PYZus{}kws=dict(s=50))}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{p}\PY{o}{.}\PY{n}{Dprime}\PY{p}{[}\PY{n}{mwind}\PY{p}{]}\PY{p}{,} \PY{n}{p}\PY{o}{.}\PY{n}{inc}\PY{p}{[}\PY{n}{mwind}\PY{p}{]}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{o}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{c}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{orange}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{s}\PY{o}{=}\PY{l+m+mi}{50}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{D \PYZlt{} \PYZob{}\PYZcb{}}\PY{l+s}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{Dmax}\PY{p}{)}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{scatter}\PY{p}{(}\PY{n}{p}\PY{o}{.}\PY{n}{Dprime}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{mwind}\PY{p}{]}\PY{p}{,} \PY{n}{p}\PY{o}{.}\PY{n}{inc}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{mwind}\PY{p}{]}\PY{p}{,} \PY{n}{marker}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{o}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{c}\PY{o}{=}\PY{n}{color}\PY{p}{,} \PY{n}{s}\PY{o}{=}\PY{l+m+mi}{50}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{D \PYZgt{} \PYZob{}\PYZcb{}}\PY{l+s}{\PYZdq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{Dmax}\PY{p}{)}\PY{p}{)}
          \PY{n}{inc\PYZus{}pts} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{linspace}\PY{p}{(}\PY{o}{\PYZhy{}}\PY{l+m+mf}{90.0}\PY{p}{,} \PY{l+m+mf}{90.0}\PY{p}{)}
          \PY{n}{Dprime\PYZus{}pts} \PY{o}{=} \PY{n}{Dmax}\PY{o}{*}\PY{n}{np}\PY{o}{.}\PY{n}{cos}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{radians}\PY{p}{(}\PY{n}{inc\PYZus{}pts}\PY{p}{)}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{Dprime\PYZus{}pts}\PY{p}{,} \PY{n}{inc\PYZus{}pts}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZhy{}}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{k}{for} \PY{n}{D} \PY{o+ow}{in} \PY{n}{np}\PY{o}{.}\PY{n}{arange}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{,} \PY{l+m+mi}{200}\PY{p}{,} \PY{l+m+mi}{10}\PY{p}{)}\PY{p}{:}
              \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{n}{Dprime\PYZus{}pts}\PY{o}{*}\PY{n}{D}\PY{o}{/}\PY{n}{Dmax}\PY{p}{,} \PY{n}{inc\PYZus{}pts}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZhy{}}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{k}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{lw}\PY{o}{=}\PY{l+m+mf}{0.5}\PY{p}{,} \PY{n}{alpha}\PY{o}{=}\PY{l+m+mf}{0.3}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper center}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{frameon}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{shadow}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Projected separation, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Inclination, degrees}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}marg\PYZus{}x}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}marg\PYZus{}y}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{best}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{fig}\PY{o}{.}\PY{n}{tight\PYZus{}layout}\PY{p}{(}\PY{p}{)}
          \PY{c}{\PYZsh{}g.fig.savefig(\PYZdq{}proplyd\PYZhy{}joint\PYZhy{}separation\PYZhy{}size.pdf\PYZdq{});}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_111_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Distribution of sizes}\label{distribution-of-sizes}

    Read in the size table that I converted from the DS9 regions.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}874}]:} \PY{n}{sizedata} \PY{o}{=} \PY{n}{Table}\PY{o}{.}\PY{n}{read}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd\PYZhy{}Datasets/proplyd\PYZhy{}chords\PYZhy{}D60.dat}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{n}{format}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{ascii.commented\PYZus{}header}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{sizedata}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}874}]:} <Table rows=4 names=('ID','RA1','DEC1','RA2','DEC2')>
          array([('154-240', '5:35:15.394', '-5:22:39.61', '5:35:15.354', '-5:22:39.71'),
                 ('163-249', '5:35:16.340', '-5:22:48.79', '5:35:16.324', '-5:22:48.80'),
                 ('165-235S', '5:35:16.487', '-5:22:35.29', '5:35:16.458', '-5:22:35.22'),
                 ('165-235N', '5:35:16.488', '-5:22:34.97', '5:35:16.479', '-5:22:34.97')], 
                dtype=[('ID', '<U8'), ('RA1', '<U11'), ('DEC1', '<U11'), ('RA2', '<U11'), ('DEC2', '<U11')])
\end{Verbatim}

    Create \texttt{coord.SkyCoord} objects for the two ends of the chord
that I measured at the back of each proplyd's head. Then use those to
find the chord diameter in arcsec (\texttt{Chord} column), and then
convert to a radius in units of $10^{14}$ cm (\texttt{r14} column).

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}875}]:} \PY{n}{coords1} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{p}{(}\PY{n}{ra}\PY{o}{=}\PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{RA1}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{dec}\PY{o}{=}\PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{DEC1}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{hourangle}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{deg}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
          \PY{n}{coords2} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{p}{(}\PY{n}{ra}\PY{o}{=}\PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{RA2}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{dec}\PY{o}{=}\PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{DEC2}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{unit}\PY{o}{=}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{hourangle}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{deg}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
          \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Chord}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{p}{[}\PY{n}{c1}\PY{o}{.}\PY{n}{separation}\PY{p}{(}\PY{n}{c2}\PY{p}{)}\PY{o}{.}\PY{n}{arcsec} \PY{k}{for} \PY{n}{c1}\PY{p}{,} \PY{n}{c2} \PY{o+ow}{in} \PY{n+nb}{zip}\PY{p}{(}\PY{n}{coords1}\PY{p}{,} \PY{n}{coords2}\PY{p}{)}\PY{p}{]}
          \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{=} \PY{l+m+mf}{0.5} \PY{o}{*} \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Chord}\PY{l+s}{\PYZdq{}}\PY{p}{]} \PY{o}{*} \PY{l+m+mi}{430}\PY{o}{*}\PY{n}{u}\PY{o}{.}\PY{n}{AU}\PY{o}{.}\PY{n}{in\PYZus{}units}\PY{p}{(}\PY{n}{u}\PY{o}{.}\PY{n}{cm}\PY{p}{)}\PY{o}{/}\PY{l+m+mf}{1.e14}
          \PY{n}{sizedata}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}875}]:} <Table rows=4 names=('ID','RA1','DEC1','RA2','DEC2','Chord','r14')>
          array([ ('154-240', '5:35:15.394', '-5:22:39.61', '5:35:15.354', '-5:22:39.71', 0.6056714671056789, 19.48053979189225),
                 ('163-249', '5:35:16.340', '-5:22:48.79', '5:35:16.324', '-5:22:48.80', 0.23915181790427623, 7.691969587490479),
                 ('165-235S', '5:35:16.487', '-5:22:35.29', '5:35:16.458', '-5:22:35.22', 0.4387068370843108, 14.11036586820911),
                 ('165-235N', '5:35:16.488', '-5:22:34.97', '5:35:16.479', '-5:22:34.97', 0.13440609108383775, 4.322975982579263)], 
                dtype=[('ID', '<U8'), ('RA1', '<U11'), ('DEC1', '<U11'), ('RA2', '<U11'), ('DEC2', '<U11'), ('Chord', '<f8'), ('r14', '<f8')])
\end{Verbatim}

    Take the average of the RA and DEC components of the two ends of the
chord to get an average coordinate (\texttt{coords0}) for each source.
\textbf{Note: This will not quite be the center of the proplyd, but it
is close enough.}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}876}]:} \PY{n}{coords0} \PY{o}{=} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{p}{(}\PY{n}{ra}\PY{o}{=}\PY{l+m+mf}{0.5}\PY{o}{*}\PY{p}{(}\PY{n}{coords1}\PY{o}{.}\PY{n}{ra} \PY{o}{+} \PY{n}{coords2}\PY{o}{.}\PY{n}{ra}\PY{p}{)}\PY{p}{,} \PY{n}{dec}\PY{o}{=}\PY{l+m+mf}{0.5}\PY{o}{*}\PY{p}{(}\PY{n}{coords1}\PY{o}{.}\PY{n}{dec} \PY{o}{+} \PY{n}{coords2}\PY{o}{.}\PY{n}{dec}\PY{p}{)}\PY{p}{)}
          \PY{n}{coords0}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}876}]:} <SkyCoord (ICRS): (ra, dec) in deg
              [(83.81405833333332, -5.377683333333333),
               (83.81804999999999, -5.380220833333333),
               (83.81863541666667, -5.376459722222222),
               (83.81868124999998, -5.3763805555555555)]>
\end{Verbatim}

    Remove the columns with the chord end coordinates because we don't need
them any more.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}877}]:} \PY{n}{sizedata}\PY{o}{.}\PY{n}{remove\PYZus{}columns}\PY{p}{(}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{RA1}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{RA2}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{DEC1}\PY{l+s}{\PYZdq{}}\PY{p}{,} \PY{l+s}{\PYZdq{}}\PY{l+s}{DEC2}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{)}
          \PY{n}{sizedata}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}877}]:} <Table rows=4 names=('ID','Chord','r14')>
          array([('154-240', 0.6056714671056789, 19.48053979189225),
                 ('163-249', 0.23915181790427623, 7.691969587490479),
                 ('165-235S', 0.4387068370843108, 14.11036586820911),
                 ('165-235N', 0.13440609108383775, 4.322975982579263)], 
                dtype=[('ID', '<U8'), ('Chord', '<f8'), ('r14', '<f8')])
\end{Verbatim}

    Add some more columns:

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \texttt{RA} and \texttt{Dec} corresponding to \texttt{coords0}
\item
  Projected distance \texttt{Dprime} from th1C (in arcsec)
\item
  Position angle \texttt{PA} from th1C (in degrees).
\end{itemize}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}878}]:} \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{RA}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{coords0}\PY{o}{.}\PY{n}{ra}\PY{o}{.}\PY{n}{to\PYZus{}string}\PY{p}{(}\PY{n}{unit}\PY{o}{=}\PY{n}{u}\PY{o}{.}\PY{n}{hourangle}\PY{p}{,} \PY{n}{sep}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{:}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dec}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{coords0}\PY{o}{.}\PY{n}{dec}\PY{o}{.}\PY{n}{to\PYZus{}string}\PY{p}{(}\PY{n}{sep}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{:}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{c0}\PY{o}{.}\PY{n}{separation}\PY{p}{(}\PY{n}{coords0}\PY{p}{)}\PY{o}{.}\PY{n}{arcsec}
          \PY{n}{sizedata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{PA}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{c0}\PY{o}{.}\PY{n}{position\PYZus{}angle}\PY{p}{(}\PY{n}{coords0}\PY{p}{)}\PY{o}{.}\PY{n}{deg}
          \PY{n}{sizedata}\PY{p}{[}\PY{o}{\PYZhy{}}\PY{l+m+mi}{4}\PY{p}{:}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}878}]:} <Table rows=4 names=('ID','Chord','r14','RA','Dec','Dprime','PA')>
          array([ ('163-323', 0.1011089261716131, 3.2520212137679168, '5:35:16.3125', '-5:23:22.6', 2.272420836013117, 276.2794710623649),
                 ('153-321', 0.08000000000590013, 2.573083376229769, '5:35:15.342', '-5:23:21.31', 16.82243690794256, 275.2472980493684),
                 ('167-231', 0.32854879080509697, 10.567292897982984, '5:35:16.728', '-5:22:31.18', 51.8190317547333, 4.367559662217565),
                 ('171-334', 0.20948633602138664, 6.737822609051992, '5:35:17.0625', '-5:23:34.115', 14.383406373150356, 141.56334754726322)], 
                dtype=[('ID', '<U8'), ('Chord', '<f8'), ('r14', '<f8'), ('RA', 'O'), ('Dec', 'O'), ('Dprime', '<f8'), ('PA', '<f8')])
\end{Verbatim}

    \begin{center}\rule{3in}{0.4pt}\end{center}

Now we try and merge in the HA98 table.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}747}]:} \PY{k+kn}{from} \PY{n+nn}{astropy.table} \PY{k+kn}{import} \PY{n}{join}\PY{p}{,} \PY{n}{Column}
\end{Verbatim}

    First naive attempt was a straight outer join on the \texttt{ID} column.
This gave some objects that are in the HA98 table but not in the new
data. Of these, 244-440 is OK, since it is outside our distance cut-off,
but the others need fixing. 177-341 (not in Ricci catalog) is just a
matter of the name, which should be 177-341W. There were also two really
missing: 171-334 and 154-324 (my typo).

So we need to fix things up. First make the \texttt{ID} field in the
HA98 table wider and fix the name of 177-341W.

\emph{This took much longer than it should have done. It turns out that
you can't change the dtype of a column without making a new table}.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}798}]:} \PY{n}{data} \PY{o}{=} \PY{n}{Table}\PY{p}{(}\PY{n}{data}\PY{p}{,} \PY{n}{names}\PY{o}{=}\PY{n}{data}\PY{o}{.}\PY{n}{colnames}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZlt{}U12}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZlt{}f8}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZlt{}f8}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZlt{}f8}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{)}
          \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{=} \PY{n}{Column}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{,} \PY{n}{dtype}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZlt{}U12}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{irow} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{argmax}\PY{p}{(}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+s}{\PYZsq{}}\PY{l+s}{177\PYZhy{}341}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{n}{irow}\PY{p}{]} \PY{o}{=} \PY{l+s}{\PYZsq{}}\PY{l+s}{177\PYZhy{}341W}\PY{l+s}{\PYZsq{}}
\end{Verbatim}

    So now we can do the join properly:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}917}]:} \PY{n}{jdata} \PY{o}{=} \PY{n}{join}\PY{p}{(}\PY{n}{sizedata}\PY{p}{,} \PY{n}{data}\PY{p}{,} \PY{n}{join\PYZus{}type}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{outer}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{keys}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

    And the only one missing is 244-440:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}918}]:} \PY{n}{jdata}\PY{p}{[}\PY{o}{\PYZti{}}\PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14\PYZus{}2}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{mask} \PY{o}{\PYZam{}} \PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14\PYZus{}1}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{mask}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}918}]:} <Table rows=1 names=('ID','Chord','r14\_1','RA','Dec','Dprime','PA','d','r14\_2','VicR14')>
          masked\_array(data = [('244-440', --, --, --, --, --, --, 142.3, 104.0, 180.6)],
                       mask = [(False, True, True, True, True, True, True, False, False, False)],
                 fill\_value = ('N/A', 1e+20, 1e+20, '?', '?', 1e+20, 1e+20, 1e+20, 1e+20, 1e+20),
                      dtype = [('ID', '<U12'), ('Chord', '<f8'), ('r14\_1', '<f8'), ('RA', 'O'), ('Dec', 'O'), ('Dprime', '<f8'), ('PA', '<f8'), ('d', '<f8'), ('r14\_2', '<f8'), ('VicR14', '<f8')])
\end{Verbatim}

    Now we tidy up some of the column names:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}919}]:} \PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14\PYZus{}1}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{name} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}
          \PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14\PYZus{}2}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{name} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZdq{}}
          \PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{d}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{name} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime\PYZus{}HA}\PY{l+s}{\PYZdq{}}
\end{Verbatim}

    Look at the rows that are in both datasets:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}920}]:} \PY{n}{overlap} \PY{o}{=} \PY{o}{\PYZti{}}\PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{mask} \PY{o}{\PYZam{}} \PY{o}{\PYZti{}}\PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{o}{.}\PY{n}{mask}
          \PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}920}]:} <Table rows=27 names=('ID','Chord','r14','RA','Dec','Dprime','PA','Dprime\_HA','r14\_HA','VicR14')>
          masked\_array(data = [ ('152-319', 0.4021371270602539, 12.934154456590305, '5:35:15.197', '-5:23:18.66', 19.37548214448735, 282.4844035826249, 19.16, 18.2, 14.8)
           ('153-321', 0.08000000000590013, 2.573083376229769, '5:35:15.342', '-5:23:21.31', 16.82243690794256, 275.2472980493684, 16.97, 1.2, 8.4)
           ('154-324', 0.09999999999592658, 3.216354219918985, '5:35:15.342', '-5:23:24.06', 16.79566852896244, 265.8635327222988, 16.63, 3.2, 10.3)
           ('155-338', 0.4026223646938783, 12.949761417696655, '5:35:15.505', '-5:23:37.455', 20.453470258425444, 224.42783197907104, 20.48, 17.0, 29.7)
           ('157-323', 0.1000000000016516, 3.2163542201031214, '5:35:15.705', '-5:23:22.45', 11.338007448530387, 272.01433163356654, 10.97, 2.5, 18.4)
           ('158-323', 0.1706546658700654, 5.488858547424075, '5:35:15.8045', '-5:23:22.455', 9.852961584895814, 272.2890310719733, 9.42, 6.3, 30.6)
           ('158-326', 0.23232597535592958, 7.472426312632764, '5:35:15.8285', '-5:23:25.54', 9.86108732117705, 254.1608255382134, 9.6, 11.3, 31.6)
           ('158-327', 0.4002366670401905, 12.873028930534636, '5:35:15.7755', '-5:23:26.585', 10.936257012268827, 250.02202328035557, 10.6, 16.6, 35.5)
           ('159-338', 0.19157551869142098, 6.161747280014195, '5:35:15.886', '-5:23:38.08', 17.50539471231517, 209.52961366072773, 17.2, 5.0, 13.5)
           ('159-350', 0.625922271425744, 20.13187739123473, '5:35:15.9375', '-5:23:50.125', 28.386023025109218, 196.07260600961948, 28.35, 20.1, 44.5)
           ('161-314', 0.18604946713139633, 5.9840098874612035, '5:35:16.095', '-5:23:14.055', 10.375560591839491, 327.9434914814732, 10.24, 5.3, 17.1)
           ('161-324', 0.10110892444353344, 3.252021158186754, '5:35:16.0435', '-5:23:24.31', 6.443858510726929, 256.8913668901656, 6.05, 3.5, 25.8)
           ('161-328', 0.19956873670361883, 6.418837484867318, '5:35:16.0625', '-5:23:27.805', 7.7764272710109426, 230.4040876150403, 7.74, 9.1, 20.0)
           ('163-317', 0.18169847973921033, 5.844066720858787, '5:35:16.27', '-5:23:16.325', 7.136451814769168, 336.08065007739856, 6.91, 5.0, 27.1)
           ('163-323', 0.1011089261716131, 3.2520212137679168, '5:35:16.3125', '-5:23:22.6', 2.272420836013117, 276.2794710623649, 2.14, 2.2, 19.4)
           ('166-316', 0.11587838074598988, 3.7270591892492524, '5:35:16.6065', '-5:23:15.955', 7.215630357963379, 17.183402450357246, 7.01, 2.5, 14.8)
           ('167-317', 0.24955691955806011, 8.026634513632452, '5:35:16.7455', '-5:23:16.245', 7.830071974020209, 32.50361387175205, 7.83, 7.9, 35.8)
           ('168-328', 0.1269980216773827, 4.084706229600512, '5:35:16.76', '-5:23:28.155', 6.908720522453734, 140.1817864462777, 6.64, 2.8, 21.0)
           ('169-338', 0.09449034060536078, 3.03914405760014, '5:35:16.883', '-5:23:38.095', 16.48188238894839, 157.67488234452193, 16.47, 2.8, 16.1)
           ('170-337', 0.4162404272647277, 13.38776654788322, '5:35:16.98', '-5:23:37.205', 16.295481457204364, 151.7642697781941, 16.2, 12.2, 36.8)
           ('171-334', 0.20948633602138664, 6.737822609051992, '5:35:17.0625', '-5:23:34.115', 14.383406373150356, 141.56334754726322, 14.29, 4.7, 22.9)
           ('171-340', 0.5462728212701063, 17.57006893990726, '5:35:17.0555', '-5:23:39.805', 19.120982071000434, 152.47369841097154, 19.11, 23.3, 23.2)
           ('173-341', 0.16941402292211163, 5.448955075611812, '5:35:17.327', '-5:23:41.42', 22.607217547181378, 145.23369658054506, 22.48, 4.1, 18.1)
           ('176-325', 0.2699999999970833, 8.684156394041189, '5:35:17.569', '-5:23:24.785', 16.618585154414777, 96.69168720243154, 16.38, 6.9, 23.5)
           ('177-341W', 0.6072973409169007, 19.532833652832174, '5:35:17.684', '-5:23:41.095', 25.78757452887539, 135.03756966648623, 25.84, 20.4, 51.6)
           ('180-331', 0.3261921659047774, 10.491495493550804, '5:35:18.057', '-5:23:30.745', 25.069100775623742, 108.3603522626762, 25.12, 12.2, 29.3)
           ('182-413', 0.9989495547792666, 32.12975616131363, '5:35:18.2155', '-5:24:13.155', 56.70157842265729, 152.52558854406146, 56.7, 36.9, 48.7)],
                       mask = [(False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)
           (False, False, False, False, False, False, False, False, False, False)],
                 fill\_value = ('N/A', 1e+20, 1e+20, '?', '?', 1e+20, 1e+20, 1e+20, 1e+20, 1e+20),
                      dtype = [('ID', '<U12'), ('Chord', '<f8'), ('r14', '<f8'), ('RA', 'O'), ('Dec', 'O'), ('Dprime', '<f8'), ('PA', '<f8'), ('Dprime\_HA', '<f8'), ('r14\_HA', '<f8'), ('VicR14', '<f8')])
\end{Verbatim}

    \begin{center}\rule{3in}{0.4pt}\end{center}

Check that the two estimates of the proplyd sizes are not too different.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}895}]:} \PY{n}{x}\PY{p}{,} \PY{n}{y} \PY{o}{=} \PY{p}{[}\PY{n}{np}\PY{o}{.}\PY{n}{log10}\PY{p}{(}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{n}{colname}\PY{p}{]}\PY{p}{)} \PY{k}{for} \PY{n}{colname} \PY{o+ow}{in} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}
          \PY{n}{g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{jointplot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{8}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{,} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZhy{}w}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{zorder}\PY{o}{=}\PY{o}{\PYZhy{}}\PY{l+m+mi}{100}\PY{p}{,} \PY{n}{lw}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{2.0}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{2.0}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_136_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    So it doesn't look too bad. Here are the cases where HA98 got
\textgreater{}30\% larger sizes. These are mainly large sized.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}884}]:} \PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{\PYZgt{}} \PY{l+m+mf}{1.3}\PY{o}{*}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}884}]:} <Table rows=5 names=('ID','r14','r14\_HA')>
          masked\_array(data = [('152-319', 12.934154456590305, 18.2)
           ('155-338', 12.949761417696655, 17.0) ('158-326', 7.472426312632764, 11.3)
           ('161-328', 6.418837484867318, 9.1) ('171-340', 17.57006893990726, 23.3)],
                       mask = [(False, False, False) (False, False, False) (False, False, False)
           (False, False, False) (False, False, False)],
                 fill\_value = ('N/A', 1e+20, 1e+20),
                      dtype = [('ID', '<U12'), ('r14', '<f8'), ('r14\_HA', '<f8')])
\end{Verbatim}

    And here are the cases where HA98 sizes were more than 30\% smaller.
These are mainly small sized. In this case, I think that the current
results are more reliable because in HA98 we did not remove the central
stars and so overestimated the compactness of some of the small
proplyds.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}887}]:} \PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{\PYZgt{}} \PY{l+m+mf}{1.3}\PY{o}{*}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}887}]:} <Table rows=6 names=('ID','r14','r14\_HA')>
          masked\_array(data = [('153-321', 2.573083376229769, 1.2) ('163-323', 3.2520212137679168, 2.2)
           ('166-316', 3.7270591892492524, 2.5) ('168-328', 4.084706229600512, 2.8)
           ('171-334', 6.737822609051992, 4.7) ('173-341', 5.448955075611812, 4.1)],
                       mask = [(False, False, False) (False, False, False) (False, False, False)
           (False, False, False) (False, False, False) (False, False, False)],
                 fill\_value = ('N/A', 1e+20, 1e+20),
                      dtype = [('ID', '<U12'), ('r14', '<f8'), ('r14\_HA', '<f8')])
\end{Verbatim}

    If we look at the distribution of ratios, then it seems we have
agreement to a level of about 30\%. This is not bad, given that the
sizes range over an order of magnitude.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}898}]:} \PY{n}{ratio} \PY{o}{=} \PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{o}{/}\PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]}
          \PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{(}\PY{n}{ratio}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{,} \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{fit}\PY{o}{=}\PY{n}{stats}\PY{o}{.}\PY{n}{norm}\PY{p}{,} \PY{n}{axlabel}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{r14 / r14\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{)}
          \PY{n}{plt}\PY{o}{.}\PY{n}{title}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{Current vs HA98 estimates of proplyd size:}\PY{l+s}{\PYZsq{}}
                    \PY{l+s}{\PYZsq{}}\PY{l+s}{ ratio = \PYZob{}:.3f\PYZcb{} +/\PYZhy{} \PYZob{}:.3f\PYZcb{}}\PY{l+s}{\PYZsq{}}\PY{o}{.}\PY{n}{format}\PY{p}{(}\PY{n}{ratio}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{o}{.}\PY{n}{mean}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
                                                           \PY{n}{ratio}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{o}{.}\PY{n}{std}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}898}]:} <matplotlib.text.Text at 0x1575874d0>
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_142_1.png}
    \end{center}
    { \hspace*{\fill} \\}

    We can do the same thing for the separations. This shows a much tighter
correlation, as one would expect

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}900}]:} \PY{n}{x}\PY{p}{,} \PY{n}{y} \PY{o}{=} \PY{p}{[}\PY{n}{np}\PY{o}{.}\PY{n}{log10}\PY{p}{(}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{n}{colname}\PY{p}{]}\PY{p}{)} \PY{k}{for} \PY{n}{colname} \PY{o+ow}{in} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{Dprime\PYZus{}HA}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}
          \PY{n}{g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{jointplot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{8}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{,} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZhy{}w}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{zorder}\PY{o}{=}\PY{o}{\PYZhy{}}\PY{l+m+mi}{100}\PY{p}{,} \PY{n}{lw}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{2.0}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{2.0}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_144_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    And just for fun, here is the same plot for the Vicente \& Alves (2005)
sizes.

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}896}]:} \PY{n}{x}\PY{p}{,} \PY{n}{y} \PY{o}{=} \PY{p}{[}\PY{n}{np}\PY{o}{.}\PY{n}{log10}\PY{p}{(}\PY{n}{jdata}\PY{p}{[}\PY{n}{overlap}\PY{p}{]}\PY{p}{[}\PY{n}{colname}\PY{p}{]}\PY{p}{)} \PY{k}{for} \PY{n}{colname} \PY{o+ow}{in} \PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{VicR14}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{r14}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{]}
          \PY{n}{g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{jointplot}\PY{p}{(}\PY{n}{x}\PY{p}{,} \PY{n}{y}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{l+s}{\PYZsq{}}\PY{l+s}{brown}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{8}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{plot}\PY{p}{(}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{,} \PY{p}{[}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{,} \PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZhy{}w}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{zorder}\PY{o}{=}\PY{o}{\PYZhy{}}\PY{l+m+mi}{100}\PY{p}{,} \PY{n}{lw}\PY{o}{=}\PY{l+m+mi}{1}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}xlim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{2.0}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}ylim}\PY{p}{(}\PY{l+m+mf}{0.0}\PY{p}{,} \PY{l+m+mf}{2.0}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_146_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{center}\rule{3in}{0.4pt}\end{center}

Finally, we can go back to the size-distance relationship:

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}923}]:} \PY{n}{jdata}\PY{o}{.}\PY{n}{filled}\PY{p}{(}\PY{n}{np}\PY{o}{.}\PY{n}{nan}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}923}]:} <Table rows=70 names=('ID','Chord','r14','RA','Dec','Dprime','PA','Dprime\_HA','r14\_HA','VicR14')>
          array([ ('133-353', 0.19956810372609624, 6.418817126068057, '5:35:13.2955', '-5:23:53.055', 56.13363987705124, 237.44385762445836, nan, nan, nan),
                 ('139-320', 0.15074155191820304, 4.848382266489986, '5:35:13.9075', '-5:23:20.045', 38.277129043789, 274.1998097040322, nan, nan, nan),
                 ('142-301', 0.9679778688332578, 31.13359703336854, '5:35:14.1165', '-5:23:00.71', 41.459036843045375, 302.27472634362283, nan, nan, nan),
                 ('147-323', 0.3103594930344309, 9.982260651538704, '5:35:14.7055', '-5:23:22.775', 26.25731935056034, 270.16015831371493, nan, nan, nan),
                 ('148-305', 0.14187554655648482, 4.563220128888503, '5:35:14.79', '-5:23:04.565', 30.96868722101388, 306.18424926889605, nan, nan, nan),
                 ('149-329', 0.0948264445804458, 3.049954351986543, '5:35:14.914', '-5:23:28.995', 23.945790120512257, 255.12644132749037, nan, nan, nan),
                 ('150-231', 0.19822035630358525, 6.37546879496851, '5:35:15.0175', '-5:22:30.83', 56.32415700268409, 337.45143636973563, nan, nan, nan),
                 ('152-319', 0.4021371270602539, 12.934154456590305, '5:35:15.197', '-5:23:18.66', 19.37548214448735, 282.4844035826249, 19.16, 18.2, 14.8),
                 ('153-321', 0.08000000000590013, 2.573083376229769, '5:35:15.342', '-5:23:21.31', 16.82243690794256, 275.2472980493684, 16.97, 1.2, 8.4),
                 ('154-225', 0.1803214866471138, 5.799777745431342, '5:35:15.362', '-5:22:25.15', 59.998665326323156, 344.08361682405405, nan, nan, nan),
                 ('154-240', 0.6056714671056789, 19.48053979189225, '5:35:15.374', '-5:22:39.66', 46.153018460637284, 339.35249280337854, nan, nan, nan),
                 ('154-324', 0.09999999999592658, 3.216354219918985, '5:35:15.342', '-5:23:24.06', 16.79566852896244, 265.8635327222988, 16.63, 3.2, 10.3),
                 ('155-338', 0.4026223646938783, 12.949761417696655, '5:35:15.505', '-5:23:37.455', 20.453470258425444, 224.42783197907104, 20.48, 17.0, 29.7),
                 ('157-323', 0.1000000000016516, 3.2163542201031214, '5:35:15.705', '-5:23:22.45', 11.338007448530387, 272.01433163356654, 10.97, 2.5, 18.4),
                 ('158-323', 0.1706546658700654, 5.488858547424075, '5:35:15.8045', '-5:23:22.455', 9.852961584895814, 272.2890310719733, 9.42, 6.3, 30.6),
                 ('158-326', 0.23232597535592958, 7.472426312632764, '5:35:15.8285', '-5:23:25.54', 9.86108732117705, 254.1608255382134, 9.6, 11.3, 31.6),
                 ('158-327', 0.4002366670401905, 12.873028930534636, '5:35:15.7755', '-5:23:26.585', 10.936257012268827, 250.02202328035557, 10.6, 16.6, 35.5),
                 ('159-338', 0.19157551869142098, 6.161747280014195, '5:35:15.886', '-5:23:38.08', 17.50539471231517, 209.52961366072773, 17.2, 5.0, 13.5),
                 ('159-350', 0.625922271425744, 20.13187739123473, '5:35:15.9375', '-5:23:50.125', 28.386023025109218, 196.07260600961948, 28.35, 20.1, 44.5),
                 ('159-418', 0.3142365584879447, 10.106960809866894, '5:35:15.8965', '-5:24:17.87', 55.669730359443506, 188.7524133614498, nan, nan, nan),
                 ('160-350', 0.18361197284160802, 5.905611437008119, '5:35:15.958', '-5:23:49.7', 27.89343785153164, 195.7100452719107, nan, nan, nan),
                 ('161-314', 0.18604946713139633, 5.9840098874612035, '5:35:16.095', '-5:23:14.055', 10.375560591839491, 327.9434914814732, 10.24, 5.3, 17.1),
                 ('161-324', 0.10110892444353344, 3.252021158186754, '5:35:16.0435', '-5:23:24.31', 6.443858510726929, 256.8913668901656, 6.05, 3.5, 25.8),
                 ('161-328', 0.19956873670361883, 6.418837484867318, '5:35:16.0625', '-5:23:27.805', 7.7764272710109426, 230.4040876150403, 7.74, 9.1, 20.0),
                 ('163-249', 0.23915181790427623, 7.691969587490479, '5:35:16.332', '-5:22:48.795', 34.11035029756079, 356.6931333118622, nan, nan, nan),
                 ('163-317', 0.18169847973921033, 5.844066720858787, '5:35:16.27', '-5:23:16.325', 7.136451814769168, 336.08065007739856, 6.91, 5.0, 27.1),
                 ('163-323', 0.1011089261716131, 3.2520212137679168, '5:35:16.3125', '-5:23:22.6', 2.272420836013117, 276.2794710623649, 2.14, 2.2, 19.4),
                 ('165-235N', 0.13440609108383775, 4.322975982579263, '5:35:16.4835', '-5:22:34.97', 47.87946244086696, 0.352875196377189, nan, nan, nan),
                 ('165-235S', 0.4387068370843108, 14.11036586820911, '5:35:16.4725', '-5:22:35.255', 47.59373359902839, 0.15723000502306675, nan, nan, nan),
                 ('166-250', 0.1493390719216531, 4.803273541935594, '5:35:16.586', '-5:22:50.07', 32.82935280918948, 3.187798038848313, nan, nan, nan),
                 ('166-316', 0.11587838074598988, 3.7270591892492524, '5:35:16.6065', '-5:23:15.955', 7.215630357963379, 17.183402450357246, 7.01, 2.5, 14.8),
                 ('166-406', 0.18361065843219077, 5.905569160945359, '5:35:16.569', '-5:24:06.1', 43.27999247683998, 177.9189001606502, nan, nan, nan),
                 ('167-231', 0.32854879080509697, 10.567292897982984, '5:35:16.728', '-5:22:31.18', 51.8190317547333, 4.367559662217565, nan, nan, nan),
                 ('167-317', 0.24955691955806011, 8.026634513632452, '5:35:16.7455', '-5:23:16.245', 7.830071974020209, 32.50361387175205, 7.83, 7.9, 35.8),
                 ('168-326E', 0.2435727716841236, 7.834163120955061, '5:35:16.8425', '-5:23:26.36', 6.657422789500048, 121.83327500978822, nan, nan, nan),
                 ('168-326W', 0.1005342149006856, 3.2335364635523374, '5:35:16.8225', '-5:23:25.985', 6.207970631425778, 120.34661250624949, nan, nan, nan),
                 ('168-328', 0.1269980216773827, 4.084706229600512, '5:35:16.76', '-5:23:28.155', 6.908720522453734, 140.1817864462777, 6.64, 2.8, 21.0),
                 ('168-404', 0.3001709084357675, 9.654559680836229, '5:35:16.758', '-5:24:04.305', 41.68867041110204, 173.94964873867485, nan, nan, nan),
                 ('169-338', 0.09449034060536078, 3.03914405760014, '5:35:16.883', '-5:23:38.095', 16.48188238894839, 157.67488234452193, 16.47, 2.8, 16.1),
                 ('170-249', 0.4937316418995504, 15.880158499958343, '5:35:16.9675', '-5:22:48.135', 35.519345688540575, 12.227689559813806, nan, nan, nan),
                 ('170-301', 0.16698907149310396, 5.370960047990761, '5:35:16.9735', '-5:23:00.685', 23.434419846272746, 18.955929422072533, nan, nan, nan),
                 ('170-337', 0.4162404272647277, 13.38776654788322, '5:35:16.98', '-5:23:37.205', 16.295481457204364, 151.7642697781941, 16.2, 12.2, 36.8),
                 ('170-400', 0.2947976834595547, 9.48173773256103, '5:35:16.9535', '-5:23:59.55', 37.42306087593395, 168.73016693660202, nan, nan, nan),
                 ('171-334', 0.20948633602138664, 6.737822609051992, '5:35:17.0625', '-5:23:34.115', 14.383406373150356, 141.56334754726322, 14.29, 4.7, 22.9),
                 ('171-340', 0.5462728212701063, 17.57006893990726, '5:35:17.0555', '-5:23:39.805', 19.120982071000434, 152.47369841097154, 19.11, 23.3, 23.2),
                 ('173-236', 0.521743596411042, 16.78112218100719, '5:35:17.349', '-5:22:35.5', 49.15950733797618, 15.600358889696293, nan, nan, nan),
                 ('173-341', 0.16941402292211163, 5.448955075611812, '5:35:17.327', '-5:23:41.42', 22.607217547181378, 145.23369658054506, 22.48, 4.1, 18.1),
                 ('174-305', 0.07789927861626998, 2.505516735162907, '5:35:17.368', '-5:23:04.715', 22.609238452630468, 36.674651861161394, nan, nan, nan),
                 ('174-400', 0.2947975908462822, 9.48173475379013, '5:35:17.3775', '-5:24:00.29', 39.85047911040879, 159.9759979965743, nan, nan, nan),
                 ('174-414', 0.26954302964271354, 8.66945860876404, '5:35:17.385', '-5:24:13.97', 52.94024013769274, 164.9380024540518, nan, nan, nan),
                 ('175-251', 0.19821860324606935, 6.375412410429117, '5:35:17.4775', '-5:22:51.03', 35.236510725236485, 25.445004768609063, nan, nan, nan),
                 ('175-355', 0.22822885465100226, 7.340648397939293, '5:35:17.5345', '-5:23:54.97', 35.88132097887825, 153.53608556029468, nan, nan, nan),
                 ('176-252', 0.11947118063428656, 3.8426163600744343, '5:35:17.641', '-5:22:51.48', 35.95926243531263, 29.26900271056063, nan, nan, nan),
                 ('176-325', 0.2699999999970833, 8.684156394041189, '5:35:17.569', '-5:23:24.785', 16.618585154414777, 96.69168720243154, 16.38, 6.9, 23.5),
                 ('177-341E', 0.042332549361707744, 1.3615647378500364, '5:35:17.736', '-5:23:41.065', 26.321279121121652, 133.7952787598883, nan, nan, nan),
                 ('177-341W', 0.6072973409169007, 19.532833652832174, '5:35:17.684', '-5:23:41.095', 25.78757452887539, 135.03756966648623, 25.84, 20.4, 51.6),
                 ('178-258', 0.6506124739054455, 20.9260017606295, '5:35:17.8295', '-5:22:57.93', 32.20124304201232, 39.3004420700999, nan, nan, nan),
                 ('179-354', 0.12011790445720019, 3.8634172890447855, '5:35:17.965', '-5:23:53.53', 37.9994898856793, 143.8446508426508, nan, nan, nan),
                 ('180-331', 0.3261921659047774, 10.491495493550804, '5:35:18.057', '-5:23:30.745', 25.069100775623742, 108.3603522626762, 25.12, 12.2, 29.3),
                 ('181-247', 0.30308283098129235, 9.748217424513806, '5:35:18.084', '-5:22:46.9', 43.3331964236521, 33.94422218032231, nan, nan, nan),
                 ('181-401', 0.13588072788920955, 4.3704055256992485, '5:35:18.0645', '-5:24:01.23', 45.216983977869724, 148.08474600353034, nan, nan, nan),
                 ('182-316', 0.2245153617232573, 7.221209311446509, '5:35:18.2485', '-5:23:15.48', 27.652686160939773, 74.54615273372958, nan, nan, nan),
                 ('182-316', 0.2283166074842956, 7.343470839896137, '5:35:18.187', '-5:23:31.455', 27.135360773050742, 108.49203085401089, nan, nan, nan),
                 ('182-413', 0.9989495547792666, 32.12975616131363, '5:35:18.2155', '-5:24:13.155', 56.70157842265729, 152.52558854406146, 56.7, 36.9, 48.7),
                 ('189-329', 0.14079422461206662, 4.52844098489688, '5:35:18.8795', '-5:23:28.82', 36.5668000502415, 99.39910341888888, nan, nan, nan),
                 ('190-251', 0.13794221136562448, 4.4367101364885535, '5:35:19.0345', '-5:22:50.465', 50.22511576873447, 49.85220926340271, nan, nan, nan),
                 ('191-350', 0.5153676065060261, 16.576047760627247, '5:35:19.075', '-5:23:49.68', 47.33453201276577, 124.5311720003038, nan, nan, nan),
                 ('204-330', 0.10999999999952767, 3.537989642039808, '5:35:20.42', '-5:23:29.655', 59.47200193070507, 96.57255594388522, nan, nan, nan),
                 ('205-330', 0.31035949075157415, 9.982260578113944, '5:35:20.4725', '-5:23:29.725', 60.25888129615085, 96.55337000935337, nan, nan, nan),
                 ('244-440', nan, nan, nan, nan, nan, nan, 142.3, 104.0, 180.6)], 
                dtype=[('ID', '<U12'), ('Chord', '<f8'), ('r14', '<f8'), ('RA', 'O'), ('Dec', 'O'), ('Dprime', '<f8'), ('PA', '<f8'), ('Dprime\_HA', '<f8'), ('r14\_HA', '<f8'), ('VicR14', '<f8')])
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}931}]:} \PY{n}{color} \PY{o}{=} \PY{l+s}{\PYZdq{}}\PY{l+s}{\PYZsh{}774499}\PY{l+s}{\PYZdq{}}
          \PY{c}{\PYZsh{} g = sns.JointGrid(mirrored(jdata[\PYZdq{}Dprime\PYZdq{}].filled(0.0)), doubled(jdata[\PYZdq{}r14\PYZdq{}].filled(0.0)), }
          \PY{n}{g} \PY{o}{=} \PY{n}{sns}\PY{o}{.}\PY{n}{JointGrid}\PY{p}{(}\PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{Dprime}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} \PY{n}{jdata}\PY{p}{[}\PY{l+s}{\PYZdq{}}\PY{l+s}{r14}\PY{l+s}{\PYZdq{}}\PY{p}{]}\PY{p}{,} 
                            \PY{n}{size}\PY{o}{=}\PY{l+m+mi}{9}\PY{p}{,} \PY{n}{xlim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{60}\PY{p}{)}\PY{p}{,} \PY{n}{ylim}\PY{o}{=}\PY{p}{(}\PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{35}\PY{p}{)}\PY{p}{)}
          \PY{c}{\PYZsh{} g.plot(sns.regplot, sns.distplot, stats.pearsonr);}
          \PY{n}{g}\PY{o}{.}\PY{n}{plot\PYZus{}marginals}\PY{p}{(}\PY{n}{sns}\PY{o}{.}\PY{n}{distplot}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{,} \PY{n}{bins}\PY{o}{=}\PY{l+m+mi}{10}\PY{p}{,} \PY{n}{rug}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{,} \PY{n}{kde\PYZus{}kws}\PY{o}{=}\PY{n+nb}{dict}\PY{p}{(}\PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{KDE}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{plot\PYZus{}joint}\PY{p}{(}\PY{n}{sns}\PY{o}{.}\PY{n}{regplot}\PY{p}{,} \PY{n}{color}\PY{o}{=}\PY{n}{color}\PY{p}{,} \PY{n}{ci}\PY{o}{=}\PY{l+m+mi}{99}\PY{p}{,} \PY{n}{order}\PY{o}{=}\PY{l+m+mi}{2}\PY{p}{,} \PY{n}{label}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{Quadratic regression fit with 99}\PY{l+s+si}{\PYZpc{} c}\PY{l+s}{onfidence bounds}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper left}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}xlabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Projected separation, \PYZdl{}D}\PY{l+s}{\PYZsq{}}\PY{l+s}{\PYZdl{}}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}joint}\PY{o}{.}\PY{n}{set\PYZus{}ylabel}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{Proplyd radius \PYZdl{}r\PYZus{}0\PYZdl{}, \PYZdl{}10\PYZca{}\PYZob{}14\PYZcb{}\PYZdl{} cm}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}marg\PYZus{}x}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{upper right}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{ax\PYZus{}marg\PYZus{}y}\PY{o}{.}\PY{n}{legend}\PY{p}{(}\PY{n}{loc}\PY{o}{=}\PY{l+s}{\PYZdq{}}\PY{l+s}{best}\PY{l+s}{\PYZdq{}}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{fig}\PY{o}{.}\PY{n}{tight\PYZus{}layout}\PY{p}{(}\PY{p}{)}
          \PY{n}{g}\PY{o}{.}\PY{n}{fig}\PY{o}{.}\PY{n}{savefig}\PY{p}{(}\PY{l+s}{\PYZdq{}}\PY{l+s}{proplyd\PYZhy{}joint\PYZhy{}separation\PYZhy{}size\PYZhy{}new.pdf}\PY{l+s}{\PYZdq{}}\PY{p}{)}\PY{p}{;}
\end{Verbatim}

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{Proplyd distribution_files/Proplyd distribution_149_0.png}
    \end{center}
    { \hspace*{\fill} \\}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}743}]:} \PY{n}{data}\PY{p}{[}\PY{n}{data}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{]} \PY{o}{==} \PY{l+s}{\PYZsq{}}\PY{l+s}{177\PYZhy{}341}\PY{l+s}{\PYZsq{}}\PY{p}{]}\PY{p}{[}\PY{l+s}{\PYZsq{}}\PY{l+s}{ID}\PY{l+s}{\PYZsq{}}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}743}]:} <Column name='ID' unit=None format=None description=None>
          array(['177-341'], 
                dtype='<U7')
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}897}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{jointplot}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}729}]:} \PY{n}{sizedata}\PY{p}{[}\PY{p}{:}\PY{p}{:}\PY{l+m+mi}{10}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}729}]:} <Table rows=7 names=('ID','Chord','r14','RA','Dec','Dprime','PA')>
          array([ ('154-240', 0.6056714671056789, 19.48053979189225, '5:35:15.374', '-5:22:39.66', 46.153018460637284, 339.35249280337854),
                 ('174-305', 0.07789927861626998, 2.505516735162907, '5:35:17.368', '-5:23:04.715', 22.609238452630468, 36.674651861161394),
                 ('149-329', 0.0948264445804458, 3.049954351986543, '5:35:14.914', '-5:23:28.995', 23.945790120512257, 255.12644132749037),
                 ('177-341E', 0.042332549361707744, 1.3615647378500364, '5:35:17.736', '-5:23:41.065', 26.321279121121652, 133.7952787598883),
                 ('190-251', 0.13794221136562448, 4.4367101364885535, '5:35:19.0345', '-5:22:50.465', 50.22511576873447, 49.85220926340271),
                 ('158-326', 0.1726037583473396, 5.551548265569561, '5:35:15.824', '-5:23:25.565', 9.932562348684426, 254.12789181523758),
                 ('174-414', 0.26954302964271354, 8.66945860876404, '5:35:17.385', '-5:24:13.97', 52.94024013769274, 164.9380024540518)], 
                dtype=[('ID', '<U8'), ('Chord', '<f8'), ('r14', '<f8'), ('RA', 'O'), ('Dec', 'O'), ('Dprime', '<f8'), ('PA', '<f8')])
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    \begin{center}\rule{3in}{0.4pt}\end{center}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}719}]:} \PY{n}{coord}\PY{o}{.}\PY{n}{SkyCoord}\PY{o}{.}\PY{n}{to\PYZus{}string}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}720}]:} \PY{n}{coords0}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{]}\PY{o}{.}\PY{n}{to\PYZus{}string}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{hmsdms}\PY{l+s}{\PYZsq{}}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}720}]:} ['05h35m15.374s -05d22m39.66s',
           '05h35m16.332s -05d22m48.795s',
           '05h35m16.4725s -05d22m35.255s',
           '05h35m16.4835s -05d22m34.97s']
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}721}]:} \PY{n}{coord}\PY{o}{.}\PY{n}{Angle}\PY{o}{.}\PY{n}{to\PYZus{}string}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}723}]:} \PY{n}{coords0}\PY{p}{[}\PY{p}{:}\PY{l+m+mi}{4}\PY{p}{]}\PY{o}{.}\PY{n}{ra}\PY{o}{.}\PY{n}{to\PYZus{}string}\PY{p}{(}\PY{n}{unit}\PY{o}{=}\PY{n}{u}\PY{o}{.}\PY{n}{hourangle}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}723}]:} array(['5h35m15.374s', '5h35m16.332s', '5h35m16.4725s', '5h35m16.4835s'], dtype=object)
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}588}]:} \PY{n}{plt}\PY{o}{.}\PY{n}{scatter}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}672}]:} \PY{n}{sns}\PY{o}{.}\PY{n}{kdeplot}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}79}]:} \PY{k+kn}{import} \PY{n+nn}{matplotlib}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}583}]:} \PY{n}{matplotlib}\PY{o}{.}\PY{n}{markers}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}600}]:} \PY{n}{plt}\PY{o}{.}\PY{n}{legend}\PY{err}{?}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}93}]:} \PY{n}{f}\PY{o}{.}\PY{n}{axes}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]} \PY{o}{==} \PY{n}{fig}\PY{o}{.}\PY{n}{\PYZus{}ax1}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}93}]:} True
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}242}]:} \PY{n}{r} \PY{o}{=} \PY{n}{np}\PY{o}{.}\PY{n}{random}\PY{o}{.}\PY{n}{RandomState}\PY{p}{(}\PY{l+m+mi}{5}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}243}]:} \PY{n}{r}\PY{o}{.}\PY{n}{random\PYZus{}sample}\PY{p}{(}\PY{l+m+mi}{10}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}243}]:} array([ 0.22199317,  0.87073231,  0.20671916,  0.91861091,  0.48841119,
                  0.61174386,  0.76590786,  0.51841799,  0.2968005 ,  0.18772123])
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}182}]:} \PY{n}{np}\PY{o}{.}\PY{n}{r\PYZus{}}\PY{p}{[}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{,} \PY{l+m+mi}{3}\PY{p}{]}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{]}\PY{p}{]}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}182}]:} array([1, 2, 3, 0, 0, 1, 2])
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}244}]:} \PY{k+kn}{import} \PY{n+nn}{seaborn}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}249}]:} \PY{n}{seaborn}\PY{o}{.}\PY{n}{palettes}\PY{o}{.}\PY{n}{dark\PYZus{}palette}\PY{p}{(}\PY{l+s}{\PYZsq{}}\PY{l+s}{red}\PY{l+s}{\PYZsq{}}\PY{p}{,} \PY{n}{as\PYZus{}cmap}\PY{o}{=}\PY{n+nb+bp}{True}\PY{p}{)}
\end{Verbatim}

            \begin{Verbatim}[commandchars=\\\{\}]
{\color{outcolor}Out[{\color{outcolor}249}]:} <matplotlib.colors.LinearSegmentedColormap at 0x111cb7490>
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}}]:} 
\end{Verbatim}

    % Add a bibliography block to the postdoc

    \end{document}