### 2016-02-01 papers/Paper1/proplyds-table.tex
table with observations. still incomplete

\begin{table}
\begin{tabular}{|c|c|c|c|c|c|c|}
Proplyd & $R_0/D$ (1) & $R_0/D $ (2) & $R_0/D$ &$R_c/D$ (1) & $R_c/D $ (2)  & $R_c/D$  & $R_c/R_0$ \hline \\
LV2 & & & & & & & 
\end{tabular}
\end{table}