### 2016-02-02 papers/Paper1/proplyds-table.tex
Advance in completion

\begin{table}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
Proplyd & $R_0/D$ (1) & $R_0/D $ (2) & $R_0/D$ & $R_c/D$ (1) & $R_c/D $ (2)  & $R_c/D$  & $R_c/R_0$ \\
LV2     &   0.30  & 0.4  & 0.35  $\pm$ 0.05    & & & & \\
LV3     &   0.58  & 0.55 & 0.565 $\pm$ 0.015   & & & & \\
LV4     &   0.35  & 0.37 & 0.36  $\pm$ 0.01    & & & & \\
LV5     &   0.41  & 0.33 & 0.37  $\pm$ 0.04    & & & & \\
177-341 &   0.17  & 0.2  & 0.185 $\pm$ 0.015   & & & & \\
167-328 &   0.15  & 0.14 & 0.145 $\pm$ 0.005   & & & & 
\end{tabular}
\end{table}