### 2016-03-31 papers/Paper1/proplyds-table.tex
Updated methodology section. Table updated

 \begin{table*}
\begin{tabular}{c|ccccccccc}\hline
Proplyd & LV2 & LV2b & LV3 & LV4  & LV5 & 177-341 & 168-328 & 169-338 & 180-331\\\hline
$D (\")$ &7.76 & 7.21 &6.89 & 6.2 & 9.55 & 25.65 & 6.83 & 16.43 & 25.07 \\
$R_c/D$  & 0.23  & 0.19& 0.61  & 0.5  & 0.34  & 0.18  & 0.22 & 0.09 & 0.09\\
$R_0/D$  & 0.32 & 0.1 & 0.34 & 0.19 & 0.21 & 0.14 & 0.15 & 0.06 & 0.07\\
$R_c/R_0$ & 0.73 & 2.00 & 1.83 & 2.68  & 1.57 & 1.22 & 1.47 & 1.46 & 1.30  
\end{tabular}
\caption{Characteristic Radii measurements for a sample of proplyds. %Enhance description when more rows were included
}

\label{tab:proplyds}
\end{table*}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "proplyd-bowshocks.tex"
%%% End: