### 2016-06-16 papers/Paper1/sec-projection.tex

\newenvironment{Vector}{\left(\begin{array}{c}}{\end{array}\right)}
\newcommand\uvec[1]{\bm{\hat{#1}}}
\newcommand\T{_{\mathrm{\scriptscriptstyle T}}}

\section{Projection onto the plane of the sky}
\label{sec:projection}

In this section we calculate the apparent shape on the plane of the
sky of the limb brightened border of a shock or shell that is
idealized as an arbitrary cylindrically symmetric surface.

%Note: I'm aware that some of this material should be moved to an appendix, but I think it will be a future edition.
\subsection{Frames of reference}

Consider body-frame cartesian coordinates $(x,y,z)$, where \(x\) is
the symmetry axis, and spherical polar coordinates
\((R, \theta, \phi)\), where \(\theta\) is the polar angle and
\(\phi\) the azimuthal angle.  Since the surface is cylindrically
symmetric, it is can be specified as $R = R(\theta)$, so that
cartesian coordinates on the surface are:
\begin{equation}
  \begin{Vector}
    x \\ y \\ z
  \end{Vector} 
  = R(\theta)
  \begin{Vector}
    \cos\theta \\
    \sin\theta\cos\phi \\
    \sin\theta\sin\phi
  \end{Vector}.
  \label{eq:body-frame}
\end{equation} 
Suppose that the viewing direction makes an angle \(i\) with the
\(z\)~axis, so that we can define observer-frame coordinates
\((x', y', z')\), which are found by rotating the body-frame
coordinates about the \(y\)~axis:
\begin{equation}
  \begin{Vector}
    x' \\ y' \\ z'
  \end{Vector}
  = 
  \begin{Vector}
    x\cos i - z\sin i\\
    y \\
    z\cos i + x\sin i
  \end{Vector}.
  \label{eq:Trans}
\end{equation} 
The relationship between the two frames is illustrated in
Figure~\ref{fig:projection-pos}.  All quantities in the observer's
frame are denoted by attaching a prime to the equivalent quantity in
the body frame.  The celestial coordinates on the ``plane of the sky''
are given by \((x', y')\) with \(x'\) being the projected symmetry
axis of the surface, while the line of sight lies along \(-z'\).  The
inclination angle \(i\) is defined so that \(i = 0^\circ\) when the
surface is viewed perpendicular to its axis (\textit{side on}) and
\(i = 90^\circ\) when it is viewed along its axis (\textit{end on}).

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/projection-pos}
  \caption{Relationship between body frame (unprimed coordinates) and
    observer frame (primed coordinates).}
  \label{fig:projection-pos}
\end{figure}

\subsection{Unit vectors normal and tangential to the surface}
\label{sec:unit-vectors-normal}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/bowshock-unit-vectors}
  \caption{Normal and tangential unit vectors}
  \label{fig:unitvec}
\end{figure}
We define unit vectors \(\uvec{n}, \uvec{t}\), such that \(\uvec{n}\) is normal to the surface, while \(\uvec{t}\) is tangent to the surface in a plane of constant \(\phi\). For \(\phi = 0\) the surface lies in the \(xy\) plane and it is straightforward to show (Fig.~\ref{fig:unitvec}) that in this case the unit vectors are given by 
\begin{equation}
  \label{eq:uvec-phi0}
  \uvec{t}_0 =
  \begin{Vector}
    -\cos\alpha \\ \sin\alpha \\ 0
  \end{Vector}
  \quad \mathrm{and} \quad
  \uvec{n}_0 =
  \begin{Vector}
    \sin\alpha \\ \cos\alpha \\ 0
  \end{Vector}, 
\end{equation}
where
\begin{equation}
  \label{eq:alpha}
  \tan\alpha = -\left.\frac{dy}{dx}\right\vert_{R(\theta)} 
  = \frac{1 + \omega \tan\theta}{\tan\theta - \omega}
\end{equation}
and 
\begin{equation}
  \label{eq:omega}
  \omega(\theta) = \frac{1}{R} \frac{dR}{d\theta} . 
\end{equation}
For general \(\phi\), \(\uvec{n}\) and \(\uvec{t}\) can be found by rotating equations~(\ref{eq:uvec-phi0}) around the \(x\)-axis, which can then be converted to the observer frame by use of equation~(\ref{eq:Trans}) to give
\begin{gather}
  \label{eq:nprime}
  \begin{split}
    \uvec{n}' = & \frac{1}{(1+\omega^2)^{1/2}} \\
    & \times 
      \begin{Vector}
        (\cos\theta+\omega\sin\theta)\cos i 
        - (\sin\theta-\omega\cos\theta) \sin i \sin\phi\\
        (\sin\theta-\omega\cos\theta)\cos\phi \\
        (\cos\theta+\omega\sin\theta)\sin i 
        + (\sin\theta-\omega\cos\theta)\sin\phi\cos i
      \end{Vector}\\
  \end{split}\\
  \label{eq:tprime}
  \begin{split}
    \uvec{t}' = & \frac{1}{(1+\omega^2)^{1/2}} \\
    & \times 
      \begin{Vector}
        -(\sin\theta-\omega\cos\theta)\cos i
        - (\cos\theta+\omega\sin\theta)\sin\phi\sin i \\
        (\cos\theta+\omega\sin\theta)\cos\phi \\
        -(\cos\theta+\omega\sin\theta)\sin i 
        + (\sin\theta-\omega\cos\theta)\sin\phi\cos i
      \end{Vector}
  \end{split}
  \end{gather}

\subsection{Tangent line}

% For optically thin emission with isotropic emissivity \(j\), the observed
% surface brightness for any line of sight that intersects the shell is
% \(S = \int j \, dz' / 4\pi\).

% If the shell thickness is \(h(\theta)\) and emissivity is
% \(j(\theta)\), then for optically thin emission the observed
% surface brightness for any line of sight that intersects the shell is
% \(S = j s / 4\pi\), where \(s\)

% in the limit \(h \ll R\) and
% \(\mu\) is
% the cosine of the angle between the line of sight and the local normal
% to the shell surface at the point of intersection:
% \begin{equation}
%   \label{eq:2}
%   \mu = \uvec{n} \cdot \uvec{z}'
% \end{equation}

The boundary on the plane of the sky of the projected surface is the
locus of those lines of sight that graze the surface tangentially.
This corresponds to a curved line on the surface itself, which we
denote the \textit{tangent line}, and which is defined by the
condition 
\begin{equation}
  \label{eq:tangent-line-condition}
  \uvec{n}'\cdot \uvec{z}' = 0.
\end{equation}
We denote by \(\phi\T\) that value of $\phi$ that satisfies this
relation for a given inclination, \(i\), and polar angle, \(\theta\):
\begin{equation}
\sin\phi\T = \tan i\tan\alpha = \tan i \frac{1+\omega\tan\theta}{\omega-\tan\theta}
\label{eq:tanphi}
\end{equation}
From equations~(\ref{eq:body-frame}, \ref{eq:Trans}) it follows that
the observer-frame coordinates of the tangent line are given by
\begin{equation}
\left(\begin{array}{c}
x'\T \\ y'\T \\ z'\T
\end{array}\right)= R(\theta)\left(\begin{array}{c}
\cos\theta\cos i - \sin\theta\sin\phi\T \sin i \\
\sin\theta(1-\sin^2\phi\T)^{1/2} \\
\cos\theta\sin i +\sin\theta\sin\phi\T\cos i
\end{array}\right).
\label{eq:tangential}
\end{equation} 
Note that, in general, \(z'\T\) is not a linear function of \(x'\T\)
and \(y'\T\), so that the tangent line is not a plane curve.

The apparent shape \((x'\T, y'\T)\) of the tangent line on the plane
of the sky can also be described in polar form as \(R'(\theta')\),
where
\begin{equation}
  \label{eq:R-prime-theta-prime}
  R' = (x'^2\T+ y'^2\T)^{1/2} 
  \quad \text{and} \quad
  \tan\theta' = y'\T / x'\T.
\end{equation}
It is important to note that equation~(\ref{eq:tanphi}) does not have
a solution for arbitrary values of $\theta$ and $i$, but only when
$|\tan i\tan\alpha|<1$. In particular, if $i\neq 0$, then the tangent
line only exists for \(\theta > \theta_{0}\) where \(\theta_{0}\) is
given implicitly by
\begin{align}
\tan\theta_{0} = \frac{|\tan i| + \omega(\theta_{0})}{1-\omega(\theta_{0}) |\tan i|} . 
\label{eq:thetapar}
\end{align}
In addition, if the surface is sufficiently ``open''
(\(\alpha > \alpha_{\mathrm{min}} > 0\) for all \(\theta\)), then for
those inclinations with
\(\vert i\vert > (90^\circ - \alpha_{\mathrm{min}}) \) the tangent
line does not exist for any value of \(\theta\).  In other words, when
the viewing angle is sufficiently close to face-on, the projected
surface has no ``edge'' and will no longer appear to be a bow shock.

\subsection{Parallel and perpendicular projected shell radii}

Considering further applications to bow shocks, we will consider open shells. In order to compare the shell shape given by $R(\theta)$ with observations,
it is convenient to define the following apparent radii in the observer frame: $R'_{0}$ and $R'_{90}$. These are projected distances of the shell tangent line
from the origin. The first is measured in the direction of the symetry axis, and the second in a perpendicular direction. More concretely $R'_{0} = x'\T(y'\T=0)$
and $R'_{90} = y'\T(x'\T=0)$. From equations (\ref{eq:tanphi}) and (\ref{eq:tangential}) we find that:
\begin{align}
R'_{0} = R(\theta_{0})\cos(\theta + i) \label{eq:Rpar} 
\end{align}
Where $\theta_{0}$ is the solution of equation (\ref{eq:thetapar}), and
\begin{align}
R'_{90} = R(\theta_{90})\sin\theta_{90}\left(1-\sin^2(\phi\T(\theta_{90}))\right)^{1/2}
\end{align}
Where $\theta_{90}$ is the solution of the next implicit equation:
\begin{align}
\cot\theta_{90} = \frac{1-\left(1+\omega(\theta_{90})^2\sin^22i\right)^{1/2}}{2\omega(\theta_{90})\cos^2 i}
\end{align}

  \subsection{Line-of-sight velocities on the tangent line}
  \label{sec:line-sight-veloc}
  Motions in a thin shocked shell will be predominantly tangential to the shell surface. In addition, for the particular case of wind-wind bowshocks, the flow in each azimuthal slice can be shown to be independent \citep{Wilkin:2000a}, which implies that the shell velocity is parallel to \(\uvec{t}\). The projected line-of-sight shell velocity is therefore
  \begin{equation}
    \label{eq:vlos}
    v_{\mathrm{los}} = (\uvec{t}' \cdot -\uvec{z}') \, v_{\parallel}(\theta) = \frac{v_{\parallel}(\theta) (1+\omega^2)^{1/2} \sin i }{\sin\theta - \omega\cos\theta} ,
  \end{equation}
  where \( v_{\parallel}(\theta)\) is the gas velocity along the shell and the standard sign convention has been adopted such that velocities away from the observer are deemed positive.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "proplyd-bowshocks"
%%% End: