### 2016-09-26 papers/Paper1/sec-conclusions-proplyd.tex

\section{Conclusions}
\label{sec:conclusions-proplyd}

Evidence for the leaky sieve model for channeling the stellar wind
momentum.

Alternative would be nebular champagne flow from Orion S.  We need to
work out the numbers for that. But the orientation seems too close to
radial for that to be plausible.

We also need to show that radiation pressure cannot be important.

Make sure we analyse the distribution of inclinations, taking into
account there is a maximum inclination above which a bowshock will not
be seen.