### 2016-10-21 papers/Paper1/proplyd-bowshocks.tex
Incorporate appendix into both papers

\RequirePackage{amsmath}
\documentclass[useAMS, usenatbib]{mnras}

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
%\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage[varg]{newtxmath}
\usepackage{newtxtext}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{siunitx}
\usepackage{color}
\usepackage{appendix}
\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}
%\usepackage{astrojournals}

\title[Proplyd bowshocks]{What is confining the proplyd bow shocks in Orion?}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Tarango Yong \& Henney]{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  We analyze the projected shapes of bowshocks associated with
  photoevaporating disks (proplyds) in the inner Orion Nebula.  We
  show that the photoevaporation flows must be anisotropic in order to
  reproduce the observed shapes.  The stagnation pressures of the
  shocked shells within 0.02~pc of the dominant high-mass star are
  consistent with an interaction with its unshocked stellar wind, but
  the more distant shells require a confining pressure that is an
  order of magnitude higher than can be provided by the unmodified
  wind.  We show that these outer objects may be proplyds that are
  interacting with the conical wake generated by the wind-wind
  interaction in the inner proplyds. 
\end{abstract}

\newcommand\thC{\(\theta^1\)\,Ori~C}
\defcitealias{Canto:1996}{CRW}
\newcommand\CRW{\citetalias{Canto:1996}}

\input{sec-intro-proplyd}
\input{sec-application}
\input{sec-conclusions-proplyd}

\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: