### 2017-05-08 papers/Paper1/quadrics-bowshock.tex
Add a better abstract to the quadrics paper
This was originally written for the Grazyna Puebla conference

%\RequirePackage{amsmath}
\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
%\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage[varg]{newtxmath}
\usepackage{newtxtext}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

% Definitions needed in abstract
\newcommand\hii{\ion{H}{ii}}

\title[Bowshock shapes]{True versus apparent shapes of astrophysical bow shocks}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Tarango Yong \& Henney]{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.  We
  present a new two-dimensional classification scheme for the shape of
  such bow shocks, which accounts for both the apex (pointy versus
  flat) and the wings (open versus closed).  For different theoretical
  bow shock models, we show how variations in the true shape can be
  disentangled from projection effects due to variations in the
  inclination angle.  We apply our classification scheme to three
  different observational datasets: mid-infrared arcs around hot
  main-sequence stars (\(N = 227\)), far-infrared arcs around luminous
  cool stars (\(N = 7\)), and emission-line arcs around proplyds and other
  young stars in the Orion Nebula (\(N = 18\)).  We find significant
  differences between the three datasets: cool star bow shocks have
  markedly more closed wings than hot star bow shocks, while
  differences in the shape of the apex region are only marginally
  significant.  The Orion Nebula arcs, on the other hand, have both
  significantly more open wings and significantly flatter apices than
  the OhotB star bow shocks.  We discuss the implications of these
  differences for understanding the physics of the bow shock
  interaction in the three classes of sources.
\end{abstract}

\input{sec-intro}
\input{sec-generic-model}
\input{sec-projection}
\input{sec-conic}
\input{sec-thin-shell}
\clearpage
\input{sec-quadrics-observations}
\input{sec-conclusions}
\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\clearpage
\input{quadrics-appendix}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: