### 2017-09-21 papers/Paper1/sec-generic-model.tex
Expand some fig captions and move figures between sections

\section{Generic two-wind bow shock model}
\label{sec:generic-model}

\begin{figure}
  % \includegraphics[width=\linewidth]{2winds-scheme}
  \includegraphics[width=\linewidth]{figs/generic-bowshock}
  \caption{Quasi-stationary bow shock structure formed by the
    interaction of two supersonic winds.  Lower-left inset box shows
    the case where the inner wind is anisotropic. The streamlines
    (thin lines) are drawn so as to be qualitatively realistic: they
    are straight in regions of hypersonic flow, but curved in subsonic
    regions, responding to pressure gradients in the shocked shells.}
\label{fig:2-winds}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/bowshock-crw-variables}
  \caption[]{Schematic diagram of two-wind interaction problem,
    following \citet{Canto:1996}.}
  \label{fig:crw-schema}
\end{figure}

The general case of a two-wind interaction bow shock is illustrated in
Figure~\ref{fig:2-winds}.  If the winds are isotropic, then the bow
shock pattern wraps around the weaker of the two sources, in

The bow shocks we consider are originated by a source located at the
origin, emitting a wind with a mass loss rate of $\dot{M}_w$ and a
terminal (supersonic) velocity $v_w$. This wind interacts with another
wind originated by another source located at a distance $D$ from the
first one.  The mass loss rate of the second source is $\dot{M}_{w1}$
and the terminal velocity is $v_{w1}$. The momentum of the wind of the
second source is higher than the momentum of the first one, and the
resultant bow shock is stationary due to pressure balance.

\subsection{Characteristic Radii}
\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/characteristic-radii}
  \caption{Characteristic radii in the body frame.}
  \label{fig:characteristic-radii}
\end{figure}

In order to contrast different bow shock models, we derive a set of
measurable radii. Each model used should predict them and these
predictions can be compared with observations.
\begin{itemize}
\item Radius at axis of symmetry. Denoted as $R_0$.
\item Radius of Curvature at the axis of symmetry. Denoted as $R_c$
\item Radius at the perpendicular direction to the symmetry
  axis. Denoted as $R_{90}$
\item For open bow shocks, the asymptotic angle. Denoted as
  $\theta_\infty$
\end{itemize}

% \begin{figure}
%   \includegraphics[width=\linewidth]{ch-radii_ed2}
%   \caption{Schematic representation of the characteristic radii
%   $R_0$, $R_{90}$ and the radius of curvature at the symmetry axis
%   $R_c$}
% \end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End: