### 2017-10-13 papers/Paper1/obs-bowshocks.tex
Split observational material into its own paper !!!!!!!

%\RequirePackage{amsmath}
\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
%\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage[varg]{newtxmath}
\usepackage{newtxtext}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

% Define a fermata symbol for use inline with text
% 04 Jul 2017 WJH
% Based on info in Secs 18.4.1 and 24.1 of musixdoc.pdf
\usepackage{musixtex}
\makeatletter
\newcommand\textfermata{%
  {\let\extractline\relax
    \setlines10\smallmusicsize \nobarnumbers \nostartrule
    \staffbotmarg0pt \setclefsymbol1\empty \global\clef@skip0pt
    \raisebox{0ex}[0ex][0ex]{
      \startextract\addspace{-\afterruleskip}%
      \notes\fermataup{-2}\en
      \zendextract}}}
\makeatother

% % The \eye command is in the dingbat package, but we need to take care
% % of a naming conflict with the AMS package
% \usepackage{savesym}
% \savesymbol{checkmark}
% \usepackage{dingbat}

% The "eye" symbol is drawn with the \faEye command of the fontawesome
% package
\usepackage{fontawesome}
% The ``irregular'' symbol is drawn from the Icelandic Sorcery and Witchcraft 
\usepackage{staves}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

% Definitions needed in abstract
\newcommand\hii{\ion{H}{ii}}

\title[Stellar bow shocks]{Weather vanes and runaways: Bow
  shocks across the HR diagram}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Tarango Yong]{
  William J. Henney \& Jorge A. Tarango Yong \\
  \AddressCRyA
}
\begin{document}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.
  We recently proposed a new two-dimensional classification scheme for
  the shape of such bow shocks, which we here apply to three very
  different observational datasets.  These are: (1)~mid-infrared arcs
  around hot main-sequence stars (\(N = 227\)); (2)~far-infrared arcs
  around luminous cool stars (\(N = 7\)); and (3)~emission-line arcs
  around proplyds and other young stars in the Orion Nebula
  (\(N = 18\)).  We find significant differences between the three
  datasets: cool star bow shocks have markedly more closed wings than
  hot star bow shocks, while differences in the shape of the apex
  region are only marginally significant.  The Orion Nebula arcs, on
  the other hand, have both significantly more open wings and
  significantly flatter apices than the hot star bow shocks.  We
  discuss the implications of these differences for understanding the
  physics of the bow shock interaction in the three classes of
  sources.
\end{abstract}

\input{sec-obs-intro}
\clearpage
\input{sec-quadrics-observations}
\input{sec-obs-conclusions}
\clearpage
\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-p-values}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: