### 2017-10-13 papers/Paper1/sec-obs-conclusions.tex

\section{Conclusions}
\label{sec:conclusion}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "obs-bowshocks"
%%% End: