### 2017-11-01 papers/Paper1/sec-generic-model.tex
Add figure with empirical terminology to theory paper
Also, finish abstract and add keywords

\section{Generic two-wind bow shock model}
\label{sec:generic-model}

% \begin{figure}
%   \includegraphics[width=\linewidth]{ch-radii_ed2}
%   \caption{Schematic representation of the characteristic radii
%   $R_0$, $R_{90}$ and the radius of curvature at the symmetry axis
%   $R_c$}
% \end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End: