### 2017-11-03 papers/Paper1/sec-conic.tex
Convert first part of quadrics section to planitude-alatude notation

\section{Quadrics of revolution}
\label{sec:conic}

\newcommand\Sin{\ensuremath{\mathcal{S}}}
\newcommand\Cos{\ensuremath{\mathcal{C}}}
\newcommand\Cot{\ensuremath{\mathcal{T}}}

For an arbitrary surface of revolution, application of
equations~(\ref{eq:tanphi}, \ref{eq:tangential}) to determine the
projected shape of the tangent line is not straightforward and in
general requires numerical techniques.  However, analytical results
can be found for the important class of surfaces known as
\textit{quadrics of revolution} \citep{Goldman:1983a, Gfrerrer:2009a},
which are formed by rotating a conic section plane curve about its
symmetry axis.  Examples are the sphere, spheroids (oblate and
prolate), and right circular paraboloids and hyperboloids.\footnote{We
  consider only the case of a single sheet of a 2-sheet hyperboloid or
  paraboloid, since these are the versions that resemble a bow,
  whereas the 1-sheet versions resemble the waist of an hourglass.}
We ignore the degenerate cases of cylinders, cones, and pairs of
parallel planes.  While mathematically simple, these quadrics are
sufficiently flexible that they can provide a useful approximation to
more complex bow shock shapes.

%The 
%In this section we will analyze the case where the resultant shape  of the bow shock is a conic curve (circle, ellipse, parabola or hyperbola).
%These curves are mathematical simple to model and give us a good reference to understand the effects of the projection effects
%described in the last section on other bow shocks. The source of the inner wind is located at the origin, and the center of the conic is located at
%a distance $x_0$ from the source.

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure*}
\setlength\tabcolsep{2em}
\begin{tabular}{ll}
  (a) & (b) \\
  \includegraphics[height=0.45\linewidth]{figs/ellipse_edited}
      & \includegraphics[height=0.45\linewidth]{figs/hyperbola_edited}
\end{tabular}
\label{fig:conics}
\caption{Example off-center conic sections that can form quadrics of
  revolution: (a) ellipse, (b) hyperbola.  The relationship is shown
  between the conic section parameters \(a\), \(b\), \(\theta_c\) and the
  bowshock characteristic radii \(R_0\), \(R_{90}\), \(R_c\), as
  defined in Fig.~\ref{fig:characteristic-radii}. The origin (center
  of the weaker flow) is indicated by a red star, the center of
  curvature of the apex of the bow shock is indicated by a green dot,
  and the geometric center of the conic section is indicated by a blue
  cross.  }
\end{figure*}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{Example of a family of conic sections, all with the same
  normalized radius of curvature at their apex (marked by white dot):
  \(\widetilde{R}_c = R_c/R_0 = 2 \). The parameter
  $\theta_c = \pm \tan^{-1} b/a$ describes the family, with negative values
  corresponding to hyperbolae.  Differerent values of
  \(\widetilde{R}_c\) can be achieved by moving the origin along the
  \(x\)-axis while keeping the apex fixed at \(\tilde{x} = 1\).)}
\label{fig:conics-family}
\end{figure}

The shape of the quadric curves in the \(xy\) plane (\(\phi = 0\)) are
shown in Figure~\ref{fig:conics}(a) and (b) for the ellipse and
hyperbola case, respectively.  The conic section itself is fully
described by two lengths, \(a\) and \(b\), which are the
semi-axes.\footnote{Note that we do not require that \(a > b\), so
  either \(a\) or \(b\) may be the semi-major axis.}  However, the
curve can be translated along the \(x\) axis to an arbitrary point
with respect to the star, so that the apex distance \(R_0\) has no
necessary relation to \(a\) or \(b\) and therefore the star/bow
combination requires \emph{three} independent lengths for its
specification.  The displacement \(x_0\) from the star to the
``center'' of the conic section is
\begin{equation}
  \label{eq:conic-x0}
  x_0 = R_0 - \sigma a
  \quad \text{with} \quad
  \sigma = \begin{cases}
    +1 & \text{ellipse}\\
    -1 & \text{hyperbola}
  \end{cases} \ .
\end{equation}
For hyperbolas the center is ``outside'' of the bow and \(x_0\) is
always positive, whereas for ellipses the center is ``inside'' the bow
and \(x_0\) is usually negative, except when \(a < R_0\) (see
Figure~\ref{fig:conics}).

A general parametric form\footnote{%
  The special case of the parabola needs to be treated differently,
  see Appendix~\ref{app:parabola}.} %
for the \(xy\) coordinates of the quadrics (in the \(\phi = 0\)plane, and
with the star at the origin) as a function of \(t = [0, \pi]\) is then
\begin{gather}
  \label{eq:par-xy}
  \begin{aligned}
    x &= x_0 + \sigma a \Cos(t) \\ 
    y &= b\Sin(t) 
  \end{aligned}
\end{gather}
where
\begin{equation}
  \label{eq:sin-sinh-etc}
  \Sin(t), \Cos(t) = \begin{cases}
    \sin(t), \cos(t) & \text{ellipse}\\
    \sinh(t), \cosh(t) & \text{hyperbola}
  \end{cases}
\end{equation}
Except for the circle case (\(\sigma = +1\), \(a = b\)), the parametric
variable \(t\) is not actually an angle in physical space.  Instead,
the polar form of the bow shape \(R(\theta)\) must be found by substituting
equations~\eqref{eq:par-xy} into \(\theta = \tan^{-1} y/x\) and
\(R = (x^2 + y^2)^{1/2}\).

The ``type'' of quadric surface can be characterized by the
\textit{quadric parameter}:
\begin{equation}
  \label{eq:Tq}
  T_q \equiv \sigma \frac{b} {a} \ , 
\end{equation}
where \(T_q < 0\) corresponds to open surfaces (hyperboloids) and
\(T_q > 0\) corresponds to closed surfaces (oblate spheroids with
\(T_q > 1\) and prolate spheroids with \(T_q < 1\)).  Special cases
are the sphere (\(T_q = 1\)) and the paraboloid (\(T_q = 0\)).  The
set of parameters \((T_q, a, x_0)\) are then sufficient to
characterize the star/bow combination, where \(a\) is the quadric
scale and \(x_0\) is its center displacement from the star.

But, we can also characterize the star/bow by \((R_0, \Pi, \Lambda)\), where
\(\Pi\) and \(\Lambda\) are the planitude and alatude, see
\S~\ref{sec:plan-alat-bow}.  We now derive the relationship between
these two descriptions.  The apex radius of curvature for a conic
section is
\begin{equation}
  \label{eq:Rc-conic}
  R_c = \frac{b^2}{a} \ ,
\end{equation}
whereas the perpendicular radius, \(R_{90}\), is the value of \(y\)
when \(x = 0\), which can be found from equations (\ref{eq:conic-x0},
\ref{eq:par-xy}) as
\begin{equation}
  \label{eq:conic-R90}
  R_{90} = \left|T_q\right|
  \left(2aR_0 - \sigma R_0^2\right)^{1/2} \ .
\end{equation}
Combining equations (\ref{eq:planitude}, \ref{eq:alatude},
\ref{eq:conic-x0}, \ref{eq:Tq}--\ref{eq:conic-R90}) yields
\begin{align}
  \label{eq:Tq-from-Pi-Lambda}
  T_q &= \sigma \left| 2 \Pi - \Lambda^2 \right|^{1/2} \\
  \label{eq:a-from-R0-Pi-Lambda}
  a &= R_0 \frac{ \Pi  } { \left|  2 \Pi - \Lambda^2\right| } \\
  x_0 &= R_0 \frac{\Pi - \Lambda^2} {2 \Pi - \Lambda^2}
\end{align}
From these we see that it is the sign of \(2 \Pi - \Lambda^2\) that determines
\(\sigma\) and thus whether a quadric is a spheroid or a hyperboloid.  For
example, for a constant planitude, \(\Pi\), we can have a family of
different quadric types, with varying alatude, \(\Lambda\), that increases
from oblate to prolate to hyperboloid, as illustrated in
Figure~\ref{fig:conics-family}.

% We have set a whole family of curves with a singular parameter denoted as:
% \begin{align}
%   \tan\theta_c \equiv \pm\frac{b}{a}\label{eq:Thc}
% \end{align}
% Positive values of this parameter describe closed curves (i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas).
% Particular cases are $\theta_c =\frac{\pi}{4}$,
% which describes a circle and $\theta_c=0$, which describes a parabola.
% To determine uniquely a given conic in terms of measurable quantities, we use the set of characteristic radii $(R_0,R_c,R_{90})$ in such way the conic
% parameters $(a,b,\theta_c)$ could be derived from the former. Either way, the radius of curvature can be computed as:

% \begin{align}
% R_c = \frac{b^2}{a} \label{eq:Rc-conic}
% \end{align}

% For the rest of the parameters, the equation for ellipses and hyperbolas are very similar:

% \begin{align}
% R_{90} = \left|\tan\theta_c\right|\left[2aR_0\mp R_0^2\right]^{1/2}\label{eq:conic-R90}
% \end{align}

% Now, we can invert equations (\ref{eq:Rc-conic}) and (\ref{eq:conic-R90}) to estimate the parameters $a$ and $b$.
% %But first is useful to define the quantities $\tilde{R}_c \equiv \frac{R_c}{R_0}$ and $\equiv \frac{\tilde{R}_{90}}{R_0}$ for simplification purposes.
% Hereafter is useful to denote with a tilde quantities which are normalized with $R_0$:

% \begin{align}
%   \tilde{a}\equiv \frac{a}{R_0} &= \pm\frac{\tilde{R}_c}{\left(2\tilde{R}_c-\tilde{R}_{90}^2\right)} \label{eq:a-con}\\
%   \tilde{b}\equiv \frac{b}{R_0} &= \frac{\tilde{R}_c}{\left|2\tilde{R}_c-\tilde{R}_{90}^2\right|^{1/2}} \label{eq:b-con}
% \end{align}

% Dividing equations (\ref{eq:a-con}) and (\ref{eq:b-con}) and substituting into equation (\ref{eq:Thc}) we get:

% \begin{align}
% \tan\theta_c = \pm \left|2\tilde{R}_c-\tilde{R}_{90}^2\right|^{1/2} \label{eq:th_c}
% \end{align}

% Note that the term $2\tilde{R}_c-\tilde{R}_{90}^2$ acts as a discriminant which determine the quadric we are working with.

%The transformation between the conic parameters and the characteristic radii are given by:
%\begin{align}
%R_c &= \frac{b^2}{a} \\
%R_{90} &= b\left[\pm\left(1 - \frac{(a \mp R_0)^2}{a^2}\right)\right]^{1/2}\label{eq:r90-conic}
%\end{align}ut 
%In (\ref{eq:r90-conic}), the positive sign correspond to an elliptic quadric, while the negative sign to an hyperbolic quadric.
%The inverse transformation of (\ref{eq:Rc-conic}) and (\ref{eq:r90-conic}) is given by:
%\begin{align}
%\frac{a}{R_0} &= \frac{A}{2A-B^2} \label{eq:a-conic}\\
%\frac{|b|}{R_0} &= \frac{A}{\left|2A-B^2\right|^{1/2}}\\
%\tan\theta_c &= \frac{2A-B^2}{\left|B^2 - 2A\right|^{1/2}} \label{eq:thc-conic} \\
%\mathrm{where:~} B &\equiv \frac{R_{90}}{R_0} \\
%#A &\equiv \frac{R_c}{R_0}
%\end{align}

\subsection{Plane-of-sky projection of quadric surfaces}

\begin{figure*}
  \centering
  \setkeys{Gin}{width=0.48\linewidth}
  \begin{tabular}{@{}ll@{}}
    (a) & (b) \\
    \includegraphics{figs/projected-Rc-vs-i}
        & \includegraphics{figs/projected-R0-vs-i} \\
    (c) & (d) \\
    \includegraphics{figs/projected-R90-vs-i}
    & \includegraphics{figs/projected-R90-vs-Rc}
  \end{tabular}
  \caption[]{Effects of projection on quadrics of revolution:
    variation with inclination of characteristic radii.  Different
    line colors correspond to different quadric shapes (see key),
    while variation in line width corresponds to variation in the
    ``true'' dimensionless radius of curvature, \(\tilde{R}_c\), at
    the apex. Vertical gray rectangles show quartiles of \(|\sin i|\),
    which will be equally populated for an isotropic distribution of
    orientations. (a)~Projected radius of curvature, normalized by
    projected stand-off distance: \( \tilde{R}_c' = R'_c /
    R'_0\). (b)~Projected stand-off distance: \(R_0' / R_0\).
    (c)~Projected perpendicular radius, normalized by projected
    stand-off distance: \( \tilde{R}_{90}' = R'_{90} /
    R'_0\). (d)~Diagnostic diagram for dimensionless projected radii:
    \(\tilde{R}_{90}'\) versus \(\tilde{R}_c'\).  Shading indicates
    different classes of quadrics: hyperboloids (white), prolate
    spheroids (dark gray), and oblate spheroids (light gray), with the
    limiting case of paraboloids shown by the thin black line.}
  \label{fig:quadric-projection}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/projected-R90-Rc-snapshots}
  \caption{Variation with inclination angle of projected quadric parameters}
  \label{fig:projected-R90-Rc-snapshots}
\end{figure*}
Once the parametrization is done, we can find the apparent shape of
the shell in the observer's frame, following the procedure explained
in section \ref{sec:projection}.  First of all, the intrinsic 3D shape of the shell is given by:
\begin{align}
x &= a\Cos(t)+x_0 \\ 
y &= b\Sin(t)\cos\phi \\
z &=  b\Sin(t)\sin\phi
\end{align}
The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:alpha}), then:
\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:
\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
\cot t ~\mathrm{if~ellipse} \\
\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}

We could work in another frame where the origin is located at the conic's center: $(X,Y)$, where $X=x-X_0$ and $Y=y$.
In this frame, we use the transformations (\ref{eq:Trans})  to obtain the apparent shape of the shell:
\begin{align}
X' &= \frac{\Cos(t)}{a\cos i}\left(a^2\cos^2 i \pm b^2\sin^2 i\right)  \label{eq:conic-projected-x}\\
Y' &= b\Sin(t)\left(1-\frac{b^2}{a^2}\tan^2 i\Cot^2(t)\right)^{1/2}
\label{eq:conic-projected-y}
\end{align}

If the apparent shape of the shell in the observer's frame is still the same conic, then, we
can write equations(\ref{eq:conic-projected-x}) and (\ref{eq:conic-projected-y}) as follows:
\begin{align}
X' = a'\Cos( t') \label{eq:conic-projected-x-2} \\
Y' = b'\Sin (t') \label{eq:conic-projected-y-2} . 
\end{align}

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y-2}) we find that
\begin{align}
a' &= \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') &= \frac{a'\Cos(t)}{a\cos i} \\
b' &= b \label{eq:b-prime}
\end{align}

Finally, using $D' = D\cos i$ and substituting $a'$ instead of $a$ in equation (\ref{eq:x0}) we find that:

\begin{align}
R'_0 &= \pm a' + x_0\cos i \\
R'_0 &= \pm\left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} +  (R_0 \mp a)\cos i
\end{align}

Defining  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1\pm \tilde{R}_c\cot^2\theta_c(f(i;\theta_c)-1) \right)
\label{eq:qprime}
\end{align}
%Where $A\equiv \frac{R_c}{R_0}$
The Radius of curvature in the observer's frame is given by $R'_c=\frac{b'^2}{a'}$. Then:
\begin{align}
  \tilde{R}_c'\equiv  \frac{R'_c}{R'_0} &= \frac{\tilde{R}_c}{\cos^2 i f(i;\theta_c)\frac{q'}{q}}
  \label{eq:Aprime}\\
  \tan\theta'_c & = \frac{\tan\theta_c}{\cos i f(i;\theta_c)} \\
  \tilde{R}'_{90} &= \left(\frac{2\tilde{R}_cf(i;\theta_c) \mp \tan^2\theta_c\frac{q'}{q}}{\frac{q'}{q}}\right)^{1/2}\frac{\sec i}{f(i;\theta_c)}
\end{align}
where $q=\frac{R_0}{D}$ and $q' = \frac{R'_0}{D'}$.

When $R'_{90}$ is measurable, diagnostic diagrams like figure (\ref{fig:projected-R90-vs-Rc}) can be used to compare with observations.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End: