### 2017-11-19 papers/Paper1/dusty-bow-wave.tex
Split up the paper again!  Now there are two theory papers

\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\title[Bow wave shapes]{True versus apparent shapes of dusty radiative bow waves}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Tarango Yong]{
  William J. Henney \& Jorge A. Tarango Yong\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\Sin}{\mathcal{S}}
\DeclareMathOperator{\Cos}{\mathcal{C}}
\DeclareMathOperator{\Cot}{\mathcal{T}}
\DeclareMathOperator{\GammaFunc}{\Gamma}
\newcommand\w{\ensuremath{\mathrm{w}}}
\newcommand\C{\ensuremath{\mathrm{c}}}
\providecommand{\abs}[1]{\lvert#1\rvert}
\providecommand{\Abs}[1]{\left\lvert#1\right\rvert}
\newcommand\TODO[1]{%
  \begin{center}
    \framebox{\parbox{0.8\linewidth}{
        \texttt{\footnotesize\color{red} #1}}}
  \end{center}}

\newcommand\uvec[1]{\bm{\hat{#1}}}
\newcommand\T{_{\mathrm{\scriptscriptstyle T}}}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dusty stellar bow waves result from the action of a star's radiation
  pressure on a stream of dust grains that is flowing past it.  They
  are an alternative mechanism to hydrodynamic bow shocks for
  explaining the curved arcs of infrared emission seen around some
  stars.  We develop simple cylindrically symmetric models for the
  dust grain dynamics under the influence of radiation and gas--grain
  drag forces, which produce bow shapes that we call dragoids.  Using
  our recently developed two-dimensional classification scheme for
  projected bow shapes, we then analyze the inclination-dependent
  tracks of the dragoids in the planitude--alatude plane, finding
  interesting differences from the wilkinoid, cantoid, and ancantoid
  shapes of hydrodynamic bow shock models.  We analyze the prospects
  of using shape analysis of observational datasets to discriminate
  between different bow models, finding that sample sizes of order 100
  are required.
\end{abstract}

\begin{keywords}
  hydrodynamics -- stars: winds, outflows
\end{keywords}

\input{sec-dust-bow-intro}
\input{app-dust-wave}
\input{sec-standing-wave}
\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-dust-equations}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: