### 2017-11-20 papers/Paper1/sec-dust-bow-intro.tex

\section{Introduction}
\label{sec:introduction}

The paper is organized as follows.
%
In \S~\ref{sec:shape-dust-wave} we do the same for simple models of a
dusty radiation bow wave (dragoids), including the effects of
gas-grain drag.
%
In \S~\ref{sec:perturbed-bows} we investigate the effects on the
planitude--alatude plane of small-amplitude perturbations to the bow
shape.
%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: