### 2017-11-23 papers/Paper1/dusty-bow-wave.tex
Change title and intro for paper 2

\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{xr-hyper}
\usepackage{hyperref}
\externaldocument[Q-]{quadrics-bowshock}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\title[Dust waves, bow waves, bow shocks]
{Shapes of dust waves, bow waves, and bow shocks}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Tarango Yong]{
  William J. Henney \& Jorge A. Tarango Yong\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\Sin}{\mathcal{S}}
\DeclareMathOperator{\Cos}{\mathcal{C}}
\DeclareMathOperator{\Cot}{\mathcal{T}}
\DeclareMathOperator{\GammaFunc}{\Gamma}
\newcommand\w{\ensuremath{\mathrm{w}}}
\newcommand\C{\ensuremath{\mathrm{c}}}
\providecommand{\abs}[1]{\lvert#1\rvert}
\providecommand{\Abs}[1]{\left\lvert#1\right\rvert}
\newcommand\TODO[1]{%
  \begin{center}
    \framebox{\parbox{0.8\linewidth}{
        \texttt{\footnotesize\color{red} #1}}}
  \end{center}}

\newcommand\uvec[1]{\bm{\hat{#1}}}
\newcommand\T{_{\mathrm{\scriptscriptstyle T}}}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves and bow waves result from the action of a star's
  radiation pressure on a stream of dust grains that is flowing past
  it.  They are an alternative mechanism to hydrodynamic bow shocks
  for explaining the curved arcs of infrared emission seen around some
  stars.  We employ simple cylindrically symmetric models for the dust
  grain dynamics under the influence of radiation and weak gas--grain
  coupling, which produce bow shapes that we call dragoids.  We also
  show how thin-shell hydrodynamic models can be modified to treat the
  case of strong gas-grain coupling, producing bow shapes that we call
  trapoids.  Using our recently developed two-dimensional
  classification scheme for projected bow shapes, we then analyze the
  inclination-dependent tracks of the dragoids and trapoids in the
  planitude--alatude plane, finding interesting differences from the
  wilkinoid, cantoid, and ancantoid shapes of hydrodynamic bow shock
  models. We further show how these are modified in the presence of
  small-amplitude standing-wave oscillations of the basic shape, as
  may be produced by hydrodynamic instabilities or a time-varying
  source.  We analyze the prospects of using shape analysis of
  observational datasets to discriminate between different bow models,
  finding that sample sizes of order 100 are required.
\end{abstract}

\begin{keywords}
  circumstellar matter -- radiation: dynamics -- stars: winds, outflows
\end{keywords}

\input{sec-dust-bow-intro}
\input{app-dust-wave}
\input{sec-standing-wave}
\input{sec-dust-bow-discuss}
\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-dust-equations}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: