### 2017-11-24 papers/Paper1/sec-obs-conclusions.tex
Get some screenshots together for Meyer (2016) simulations
Still need to measure arcs in ds9

\section{Conclusions}
\label{sec:conclusion}

Difference in SEDs between FIR and MIR samples?  \citet{Meyer:2016a}
say emission peaks at 3--50 micron for O star bow shocks.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "obs-bowshocks"
%%% End: