### 2017-12-04 papers/Paper1/quadrics-bowshock.tex
Finish revision of discussion and abstract

%\RequirePackage{amsmath}
\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

% The following is needed to fix the margins if using Letter-size paper
% REMOVE if your LaTeX uses A4 paper by default
%\addtolength\topmargin{-1.8cm}

% Standard LaTeX packages
% \usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{dblfloatfix}
%\usepackage[section, above, below]{placeins}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{nicefrac}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{hyperref}
\usepackage{xr-hyper}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

% Definitions needed in abstract
\newcommand\hii{\ion{H}{ii}}

\title[Bow shock shapes]{True versus apparent shapes of bow shocks}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Tarango Yong \& Henney]{
  Jorge A. Tarango Yong \& William J. Henney\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2017}

\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\Sin}{\mathcal{S}}
\DeclareMathOperator{\Cos}{\mathcal{C}}
\DeclareMathOperator{\Cot}{\mathcal{T}}
\DeclareMathOperator{\GammaFunc}{\Gamma}
\DeclareMathOperator{\Depart}{\Delta}
\newcommand\ecc{\ensuremath{e}}
\newcommand\w{\ensuremath{\mathrm{w}}}
\newcommand\C{\ensuremath{\mathrm{c}}}
\providecommand{\abs}[1]{\lvert#1\rvert}
\providecommand{\Abs}[1]{\left\lvert#1\right\rvert}
\newcommand\TODO[1]{%
  \begin{center}
    \framebox{\parbox{0.8\linewidth}{
        \texttt{\footnotesize\color{red} #1}}}
  \end{center}}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Astrophysical bow shocks are a common result of the interaction
  between two supersonic plasma flows, such as winds or jets from
  stars or active galaxies, or streams due to the relative motion
  between a star and the interstellar medium.
  % Similarly shaped radiation bow waves are produced
  % by the relative motion between dust grains and a luminous source.
  For cylindrically symmetric bow shocks, we develop a general theory
  for the effects of inclination angle on the apparent shape. We
  propose a new two-dimensional classification scheme for bow shapes,
  which is based on dimensionless geometric ratios that can be
  estimated from observational images.  The two ratios are related to
  the flatness of the bow's apex, which we term \textit{planitude} and
  the openness of its wings, which we term \textit{alatude}.  We
  calculate the expected distribution in the planitude--alatude plane
  for a variety of simple geometrical and physical models: quadrics of
  revolution, wilkinoids, cantoids, and ancantoids.  We further test
  our methods against numerical magnetohydrodynamical simulations of
  stellar bow shocks and find that the apparent planitude and alatude
  measured from infrared dust continuum maps serve as accurate
  diagnostics of the shape of the contact discontinuity, which can be
  used to discriminate between different physical models.
\end{abstract}

\begin{keywords}
  circumstellar matter -- hydrodynamics -- stars: winds, outflows
\end{keywords}

\input{sec-intro}
% \input{sec-generic-model}
\input{sec-projection}
\input{sec-conic}
\input{sec-thin-shell}
% \input{app-dust-wave}
% \input{sec-standing-wave}
\input{sec-conclusions}
\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-parabola}
%\input{app-ancantoid}
\input{app-shape-parameters}
% \input{app-dust-equations}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: