### 2018-03-02 papers/Papers1234/app-rcurv-empirical.tex
Make table of the M42 bow fits

\section{Empirical determination of radius of curvature for a bow shock of unknown orientation}
\label{app:rcurv-empirical}

Consider a set of \(N\) points on the plane of the sky,\footnote{%
  In the main body of the paper, the prime symbol (\('\)) is used to
  distinguish projected from ``true'' quantities.  In this Appendix,
  for simplicity, we omit the primes since \emph{all} quantities are
  projected.} %
with Cartesian coordinates \(\bm{r}_k = (x_k, y_k)\) for
\(k = 1 \dots N\).  We wish to estimate the radius of curvature of the
smooth curve that the set of points is presumed to be sampled from.
To do this, we fit a circle to the points as follows.  The circle is
defined by its center, \(\bm{r}_{\C} = (x_{\C}, y_{\C})\), and radius,
\(R_{\C}\).  For a given circle, we define a mean radius of the set of
points from the circle center:
\begin{equation}
  \label{eq:rcurv-mean-radius}
  \bar{R}_\C(x_{\C}, y_{\C}) = \frac{1}{N}  \sum_{k = 1}^{N} \Abs{\bm{r}_k - \bm{r}_{\C}}  \ .
\end{equation}
We then optimize
% \footnote{%
%   For example, using a Levenberg-Marquardt algorithm, such as
%   \texttt{astropy.modeling.fitting.LevMarLSQFitter} from the AstroPy
%   library \citep{Astropy-Collaboration:2013a}.} %
to find best-fit values \((x_{\C}^*, y_{\C}^*)\), which minimize the
objective function
\begin{equation}
  \label{eq:rcurv-objective-function}
  f(x_{\C}, y_{\C}) = \sum_{k = 1}^{N} \left(
    \abs{\bm{r}_k - \bm{r}_{\C}}  - \bar{R}_\C (x_{\C}, y_{\C}) \right)^2 \ .
\end{equation}
The best-fit radius of curvature is then given by
\(R_{\C}^* = \bar{R}_\C(x_{\C}^*, y_{\C}^*)\).

If we also know the position, \(\bm{r}_0 = (x_0, y_0)\), of the bow's
central source, then we can find the unit vector in the direction of
the bow's projected axis as
\begin{equation}
  \label{eq:rcurv-axis-vector}
  \uvec{\xi} = \frac {\bm{r}_0 - \bm{r}_{\C}^*} { \abs{\bm{r}_0 - \bm{r}_{\C}^*}} \ , 
\end{equation}
and the apex distance from the source as\footnote{%
  This is only valid if the resultant \(R_0 < R_{\C}^*\), otherwise
  the opposite sign of \(\uvec{\xi}\) must be taken.}
\begin{equation}
  \label{eq:rcurv-apex-distance}
  R_0 = \Abs{\bm{r}_{\C}^* + R_{\C}^* \uvec{\xi} - \bm{r}_0} \ .
\end{equation}
A refinement of the method is then to iteratively repeat the circle
fit after restricting the set of points to those lying within a
certain angle \(\Delta\theta\) of the bow axis, where we find that best results
are obtained with \(\Delta\theta \approx \ang{60}\) to \ang{75}.  That is, 
\begin{equation}
  \label{eq:rcurv-delta-theta}
  \Abs{\theta_k} < \Delta\theta \ , 
\end{equation}
where the signed angle \(\theta_k\) of each point from the
axis,\footnote{%
  Although the sign of \(\theta_k\) is not relevant to
  equation~\eqref{eq:rcurv-delta-theta}, it is used below in
  calculating the perpendicular radii.  } %
measured at the source position \(r_0\), can be calculated as
\begin{equation}
  \label{eq:rcurv-theta-k}
  \theta_k = \arctan \left[  
    \frac{\left( \bm{r}_k - \bm{r}_0 \right) \cdot \uvec{\xi}^\perp}
    {\left( \bm{r}_k - \bm{r}_0 \right) \cdot \uvec{\xi}}
  \right] \ .
\end{equation}
In the preceding equation, the ``perpendicular'' operator (\(\perp\))
rotates its vector argument counter-clockwise by \ang{90}, so that
\((x, y)^\perp = (-y, x)\).

Two or three iterations are sufficient for convergence in most cases,
although in some cases it is possible that the process will converge
to a stable flip-flop oscillation between two different solutions.
This is due to the dependence of the \(\theta_k\), via \(\uvec{\xi}\), on
the \(\bm{r}_{\C}^*\) of the previous iteration, which can lead to
points entering and leaving the fitted set.  We have not found this to
be a serious problem in practice, since the two solutions tend to be
very close to one another.  It could be mitigated by averaging
\(\bm{r}_{\C}^*\) over two previous iterations.  The alternative of
measuring the angle with respect to the center of curvature,
\(\bm{r}_{\C}^*\), instead of the source, \(\bm{r}_0\), is found to be
much less stable.

If quantitative estimates exist for the uncertainties, \(\epsilon_k\), in the
measurements of \(\bm{r}_k\), then it is appropriate to incorporate
weights of \(\epsilon_k^{-2}\) in the objective function.  However, it is
rare for the \(\epsilon_k\) to be objectively quantifiable, since the
uncertainties are often systematic and/or subjective.  In cases
where the bow shape is traced by eye, based on real or synthetic
observations, a more practical approach is to maintain uniform
weighting but to place a greater density of points \(\bm{r}_k\) in
regions where the shape is well-determined and to place them more
sparsely in regions where the shape is less certain.

Since there is no guarantee of symmetry about the axis \(\uvec{\xi}\),
the perpendicular radius will in general be different in the two wings
of the bow, with values \(R_{90+}\) and \(R_{90-}\).  These can be
estimated by defining
\begin{equation}
  \label{eq:rcurv-R-k}
  R_k = \Abs{\bm{r}_k - \bm{r}_0}
\end{equation}
and linearly interpolating between the points \((\theta_k, R_k)\) at
\(\theta = \pm \ang{90}\).

Our Python language implementation of this algorithm is freely
available at \url{https://github.com/div-B-equals-0/circle-fit}.  An
example application to real data is given in \S~\ref{sec:obs} and
Figure~\ref{fig:000-400-fit}.  Note that this method is not necessary
if the orientation of the bow axis is known a~priori, in which case
the Taylor series method described in \S~\ref{sec:plan-alat-bow} is
more efficient and accurate.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "quadrics-bowshock"
%%% End: