### 2018-06-26 papers/Papers1234/app-dust-wave.tex
Finish text in appendix about inertia-confined dust waves

\section{Dust waves in the drag-free limit}
\label{sec:gas-free-bow}

% As an alternative to hydrodynamic or magnetohydrodynamic bow shocks,
% it is possible that some observed emission arcs may be bow waves due
% to the action of radiation pressure on dust grains.

\begin{figure}
  (a)\\
  \includegraphics[width=\linewidth]{figs/dust-trajectories}
  (b)\\
  \includegraphics[width=\linewidth]{figs/dust-divergent}
  \caption[Dust grain trajectories]{Dust grain trajectories under
    influence of a repulsive central \(r^{-2}\) radiative force.
    (a)~A parallel stream of dust grains approach from the right at a
    uniform velocity and with a variety of impact parameters (initial
    \(y\)-coordinate). The central source is marked by a red star at
    the origin, and its radiative force deflects the trajectories into
    a hyperbolic shape, each of which reaches a minimum radius marked
    by a small black square.  The incoming hyperbolic trajectories are
    traced in gray and the outgoing trajectories are traced in red.
    The locus of closest approach of the outgoing trajectories is
    parabolic in shape (traced by the thick, light gray line) and this
    constitutes the inner edge of the bow wave.  (b)~The same but for
    a divergent stream of dust grains that originates from a source on
    the \(x\) axis at a distance \(D = 10 R\starstar\) from the origin.  In
    this case, the inner edge of the bow wave is hyperbolic and the
    parallel stream result is also shown for comparison.}
  \label{fig:dust-trajectories}
\end{figure}

A dust grain of geometrical cross-section \(\xsec\) situated a
distance \(R\) from a point source of radiation with luminosity \(L\)
will experience a repulsive, radially directed radiative force
\citep[e.g.,][]{Spitzer:1978a}
\begin{equation}
  \label{eq:dust-rad-force}
  \frad = \frac{\xsec \Qp L} {4 \pi R^2 c} e^{-\tau}
\end{equation}
where \Qp{} is the frequency-averaged\footnote{%
  Frequency averages of any quantity \(x\) should be understood as
  weighted by the attenuated source spectrum:
  \(\langle x \rangle_\nu = (L \, e^{-\tau})^{-1} \int_0^\infty x(\nu)\, L_\nu \, e^{-\tau_\nu} \, d\nu
  \).  } %
radiation pressure efficiency\footnote{%
  \label{fn:Qp}
  For absorption efficiency \(Q_{\text{abs}}\), scattering efficiency
  \(Q_{\text{scat}}\), and asymmetry parameter (mean scattering
  cosine) \(g\), we have
  \(\Qp = Q_{\text{abs}} + (1 - g) Q_{\text{scat}}\)
  \citep[e.g., \S~4.5 of][]{Bohren:1983a}.} %
of the grain, \(c\) is the speed of light, and \(\tau\) is the
frequency-averaged optical depth between the source and the grain.
For simplicity, we will consider only the optically thin case,
\(\tau \to 0\).

% \subsection{Gas-free dust wave}

If \(\frad\) is the only force experienced by the grain, then it will
move on a \textit{ballistic} trajectory, determined by its initial
speed at large distance, \(v_\infty\), and its impact parameter, \(b\).
For \(b = 0\), the grain radially approaches the source with initial
radial velocity \(-v_\infty\), which is decelerated to zero at the distance
of closest approach, \(R\starstar\), given by energy conservation:
\begin{equation}
  \label{eq:dust-r0}
  R\starstar = \frac{\xsec \Qp L} {2 \pi c m\grain v_\infty^2} \ ,
\end{equation}
where \(m\grain\) is the grain mass.  The grain then turns round and
recedes from the source along the same radius, reaching a velocity of
\(+v_\infty\) at large distance.  Note that \(R\starstar\) as given by
equation~\eqref{eq:dust-r0} is the equivalent of
equation~\eqref{eq:rad:R0}, except for a single grain considered in
isolation, rather than a well-coupled dusty plasma.

For \(b > 0\), the trajectory, \(R\grain(\theta; b)\), is
found\footnote{%
  The problem is formally identical to that of Rutherford scattering,
  or (modulo a change of sign) planetary orbits.  The method of
  solution (via introduction of a centrifugal potential term and
  reduction to a 1-dimensional problem) can be found in any classical
  mechanics text \citep[e.g.,][\S~14]{Landau:1976a}.} %
to be hyperbolic, characterized by an eccentricity,
\(\varepsilon = \bigl( 1 + 4 b^2 / R\starstar^2\bigr)^{1/2}\), and polar angle
of closest approach, \(\thm = \cos^{-1} \varepsilon^{-1}\).  The trajectory is
symmetrical about \(\thm\) and can be written as
\begin{equation}
  \label{eq:dust-r-theta}
  \frac{R\grain(\theta; b)} {R\starstar} = 
  \frac{ \tfrac12 \bigl( \varepsilon^2 - 1 \bigr)} {\varepsilon \cos(\theta - \thm) - 1} \ , 
\end{equation}
with a total deflection angle of \(\ang{180} - 2 \thm\), which is equal to
\ang{90} when \(b = 0.5 R\starstar\).

\subsection{Parallel dust stream}
\label{sec:dust-parallel}

If the incoming dust grains initially travel along parallel
trajectories with varying \(b\), but the same \(v_\infty\), then deflection
by the radiative force will form a bow-shaped dust wave around the
radiation source, as shown in Figure~\ref{fig:dust-trajectories}.
However, the inner edge of the dust wave, \(R_{\text{in}}(\theta)\) is not
given by the closest approach along individual trajectories,
\(R\grain(\thm; b)\), but instead must found by minimizing
\(R\grain(\theta; b)\) over all \(b\) for each value of \(\theta\), which yields
\begin{equation}
  \label{eq:dust-r-in}
  \frac{R_{\text{in}}(\theta)} {R\starstar} = \frac{2}{1 + \cos\theta} \ .
\end{equation}
This is the polar form of the equation for the confocal parabola,
which we have already discussed in detail in \PaperI{}'s
\S~\ref{Q-sec:conic} and Appendix~\ref{Q-app:parabola}.  Its planitude
and alatude are \(\Gamma = \Lambda = 2\) and these are unchanged under projection
at any inclination.

\subsection{Divergent dust stream}
\label{sec:dust-divergent}

If the dust grains are assumed to originate from a second point
source, located at a distance \(D\) from the radiation source, then
the incoming stream will be divergent instead of plane parallel.  The
individual streamlines are not affected by this change and are still
described by equation~\eqref{eq:dust-r-theta}, except that the
trajectory axes for \(b > 0\) are no longer aligned with the global
symmetry axis, so we must make the substitution
\(\theta \to \theta + \theta_1(b)\), where \(\sin \theta_1 = b / D\) (see
Fig.~\ref{Q-fig:crw-schema} of \PaperI{}). We parametrize the degree
of divergence as \(\mu = R\starstar / D\) and, as before,
\(R\grain(\theta + \theta_1(b, \mu); b)\) is minimized over all trajectories to
find the shape of the bow wave's inner edge.  This time, the result is
a confocal hyperbola:
\begin{equation}
  \label{eq:dust-divergent-r-in}
  \frac{R_{\text{in}}(\theta; \mu)} {R\starstar} = \frac{1 + \varepsilon_\mu}{1 + \varepsilon_\mu\cos\theta} \ ,
\end{equation}
where the eccentricity is (to first order in \(\mu\))
\( \varepsilon_\mu = (1 - 2\mu)^{-1}\).  An example is shown in
Figure~\ref{fig:dust-trajectories} for \(\mu = 0.1\).  Unsurprisingly,
the resulting bow shape is more open than in the parallel stream case,
increasingly so with increasing \(\mu\).  The planitude and alatude are
both equal: \(\Pi = \Lambda = 1 + \varepsilon_\mu\).

\subsection{Tight magnetic coupling}
\label{sec:tight-magn-coupl}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/parallel-bfield-dust-wave-inertia}
  \caption{Dust wave formed by action of radiation forces on grains
    that are tightly coupled to a uniform parallel magnetic field.
    Two example trajectories, one with \(b > R\starstar\) (upper) and
    one with \(b < R\starstar\) (lower) are shown schematically in
    orange.  The orientation of the magnetic field lines is shown in
    blue.  The grayscale image shows the resultant dust density
    distribution.}
  \label{fig:inertia-thB0}
\end{figure}

As discussed in \S~\ref{sec:magn-effects-grain}, for small grains
(\(a < \SI{0.1}{\um}\)) the Larmor magnetic gyration radius, \(r\B\)
is typically small compared with other length scales of interest, and
so the grains are effectively tied to the magnetic field lines.  In
this approximation, we can calculate the grain dynamics in the
drag-free limit using \(f\rad\) as the sole force as above, but this
time allowing acceleration only along the field lines.  Assuming a
uniform magnetic field, the only extra parameter needed is
\(\theta\B\), the angle between the field direction and the direction of
the dust stream (assumed to be plane parallel).  The radiation force
will also produce an out-of-plane drift, given by
\begin{equation}
  \label{eq:perpendicular-drift}
  \bm{v}_{\text{drift}} = \frac{c}{e z\grain} \, \frac{\bm{f}\!\rad \times \bm{B}}{B^2}
  \ ,
\end{equation}
but from equations~(\ref{eq:Larmor}, \ref{eq:dust-rad-force},
\ref{eq:dust-r0}) it follows that
\begin{equation}
  \label{eq:vdrift-over-vinfinity}
  \frac{v_{\text{drift}}(R\starstar)}{v_\infty} = \frac{r\B}{2 R\starstar} \ ,
\end{equation}
so it is valid to ignore this drift in the limit of small \(r\B\).

We now calculate in detail the grain trajectories in this limit for
two cases, with the magnetic field oriented parallel and perpendicular
to the stream\footnote{In both cases, the stream trajectories are
  assumed parallel to one another, as in \S~\ref{sec:dust-parallel}.}
direction, respectively. These are sufficient to give a flavor of the
effects of a magnetic field on the dust wave structure.  Further
models at intermediate angles, and which also include the effects of
gas drag, are shown in \S~\ref{sec:magn-effects-grain}.

\subsubsection{Parallel magnetic field}
\label{sec:parall-magn-field}

For \(\theta\B = 0\), the \(b = 0\) trajectory is identical to the
non-magnetic case since the grain velocity and radiation force are
both parallel to the field, which yields zero Lorentz force and zero
\(\bm{f} \times \bm{B}\) drift. Therefore, the axial turnaround radius,
\(R\starstar\), is still given by equation~\eqref{eq:dust-r0},
whatever the value of \(r\B\).  For \(b > 0 \), \(f\rad\) has
a component perpendicular to \(\bm{B}\), which will induce a helical
gyromotion around the field lines, but the guiding center must move
parallel to the axis if \(r\B \ll R\starstar\).  Thus, the guiding
center motion can be found from conservation of potential plus kinetic
energy in one dimension:
\begin{equation}
  \label{eq:parallel-energy-balance}
  \frac{v^2}{v_\infty^2} + V\rad = 1 \ , 
\end{equation}
where \(V\rad\) is a suitably normalized potential of the projected
radiation force along the field lines (\(x\) axis, where
\(x = R \cos\theta = b \cot\theta\)):
\begin{equation}
  \label{eq:parallel-radiation-potential}
  V\rad = R\starstar \int_x^\infty \frac{\cos\theta}{R^2}\, dx
  = \frac{R\starstar\sin\theta}{b}
  = \frac{R\starstar}{R} \ .
\end{equation}
From equation~\eqref{eq:parallel-energy-balance}, the trajectory must
turn around when \(V\rad = 1\), and
equation~\eqref{eq:parallel-radiation-potential} shows that this
occurs at the same spherical radius, \(R = R\starstar\), for all
impact parameters, \(b\), so that the inner boundary of the dust wave
is hemispherical in shape:
\begin{equation}
  \label{eq:thB-0-shape}
  \frac{R_{\text{in}}(\theta)} {R\starstar} = 1 \ ,
\end{equation}
with \(\Pi = \Lambda = 1\).  Note, however, that this only applies to
streamlines with \(b \le R\starstar\).  For those with
\(b > R\starstar\), the maximum \(V\rad\), which occurs at \(x = 0\),
is smaller than unity, so that grains on these streamlines slow down
as they go past the star, but then speed up again without turning
round.

The grain density of the inflowing stream follows from mass continuity as:
\begin{equation}
  \label{eq:thB-0-density}
  n\grain(R) = \frac{n \bar{m} Z\grain}{m\grain} \left( 1 - R^{-1} \right)^{-1/2} \ ,
\end{equation}
where for simplicity we assume a single grain species of mass
\(m\grain\) and dust-gas ratio \(Z\grain\). The outflowing stream has
exactly the same velocity profile as the inflowing one, apart from a
change of sign for those streamlines that turn round.  Therefore, in
the region where the two streams co-exist (\(y \le R\starstar\),
\(x > 0\), and \(R > R\starstar\)), the total density is double that
given by equation~\eqref{eq:thB-0-density}.  This is illustrated in
Figure~\ref{fig:inertia-thB0}.

\subsubsection{Perpendicular magnetic field}
\label{sec:perp-magn-field}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/perp-bfield-dust-wave-inertia}
  \caption{Dust wave formed by action of radiation forces on grains
    that are tightly coupled to a uniform perpendicular magnetic
    field.}
  \label{fig:inertia-thB90}
\end{figure}

For \(\theta\B = \ang{90}\), the guiding center is forced to move at a
constant speed in the \(x\) direction, so that \(x = -v_\infty t\), while
the motion in the \(y\) direction obeys the ODE:
\begin{equation}
  \label{eq:ode-perp-bfield}
  \frac{d^2 Y}{d X^2} = \frac{Y}{2 \left( X^2 + Y^2\right)^{3/2}} \ ,
\end{equation}
where \(X = x / R\starstar\) and \(Y = y / R\starstar\).  We have been
unable to find an analytic solution to this equation, but a numerical
solution is shown in Figure~\ref{fig:inertia-thB90}.  When the impact
parameter is larger than \(b \sim R\starstar\), the trajectories are very
similar to in the non-magnetic case
(Fig.~\ref{fig:dust-trajectories}a).  As shown in
Figure~\ref{fig:inertia-thB90}b, the interaction of the grain with the
radiation field in this large-\(b\) regime can be approximated as an
impulsive acceleration of magnitude \(\sim b^{-2}\) and duration
\(\sim b\), producing a final \(y\) velocity of \(\sim b^{-1}\).  Since the
\(x\) velocity is constant, the total deflection angle is also of
order \(\sim b^{-1}\).  The overlap of the outgoing trajectories produces
a dense concentration of grains at the inner edge, which is roughly
parabolic in shape.  However, for \(b < R\starstar\) the remorseless
advance of the magnetic field does not allow the grains to slow down
and turn round, as they do in the non-magnetic and parallel-field
cases.  As a result, no dense shell forms in the apex region, but
instead there is a diffuse minimum in the density of grains around the
star due to the high grain velocities reached there.  This means that
the apparent morphology of a pure dust wave becomes very sensitive to
the radial dependence of the grain emissivity.  If this is
sufficiently steep, then the apex would coincide with the position of
the star, although in practice the presence of a wind-supported bow
shock, however small, will complicate the picture.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "dusty-bow-wave"
%%% End: