### 2019-02-26 papers/Papers1234/bs-bw-dw-02.tex

\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\sisetup{separate-uncertainty = true}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{xr-hyper}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{chemgreek}
\activatechemgreekmapping{newtx}
\usepackage{listings}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% The following hack solves a problem with
%% ERROR: \pdfendlink ended up in different nesting level than \pdfstartlink.
%% See https://tex.stackexchange.com/a/249743
\makeatletter
\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
  \errmessage{\noexpand\@combinedblfloats could not be patched}%
}%
\makeatother

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\title
{Bow shocks, bow waves, and dust waves. II. Beyond the rip point}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Arthur]{
  William J. Henney \& S. Jane Arthur\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

\input{bs-bw-dw-defs.tex}

\externaldocument[Q-]{quadrics-bowshock}
%\externaldocument[I-]{bs-bw-dw-01}
%\externaldocument[III-]{bs-bw-dw-03}

\defcitealias{Tarango-Yong:2018a}{Paper~I}
\newcommand\PaperI{\citetalias{Tarango-Yong:2018a}}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves are a result of gas--grain decoupling in a stream of
  dusty plasma that flows past a luminous star.  The radiation field
  is sufficiently strong to overcome the collisional coupling between
  grains and gas at a \textit{rip-point}, where the ratio of radiation
  pressure to gas pressure exceeds a critical value of roughly 1000.
  When the rip point occurs outside the hydrodynamic bow shock, a
  separate dust wave may form, decoupled from the gas shell, which can
  either be drag-confined or inertia-confined, depending on the stream
  density and relative velocity.  In the drag-confined case, there is
  a minimum stream velocity of roughly \SI{60}{km.s^{-1}} that allows
  a steady-state stagnant drift solution for the dust wave apex.  For
  lower relative velocities, the dust dynamics close to the axis
  exhibit a limit cycle behavior (rip and snap back) between two
  different radii.  Strong coupling of charged grains to the plasma's
  magnetic field can modify these effects, but for a quasi-parallel
  field orientation the results are qualitatively similar to the
  non-magnetic case. For a quasi-perpendicular field, on the other
  hand, the formation of a decoupled dust wave is strongly suppressed.
\end{abstract}

\begin{keywords}
  circumstellar matter -- radiation: dynamics -- stars: winds, outflows
\end{keywords}

\input{sec-weak-coupling}
\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-cloudy-dust-charging}
\input{app-dust-wave}
\input{app-dust-equations}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: