### 2019-03-08 papers/Papers1234/app-cloudy-dust-charging.tex
Major surgery on Paper II

% Inner shock \dots no dust \dots proton Larmor radius very small, so
% no penetration across contact discontinuity

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "bs-bw-dw-02"
%%% End: