### 2019-03-09 papers/Papers1234/bs-bw-dw-02.tex
Submit BS-BW-DW-01 paper

\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\sisetup{separate-uncertainty = true}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{xr-hyper}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{chemgreek}
\activatechemgreekmapping{newtx}
\usepackage{listings}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% The following hack solves a problem with
%% ERROR: \pdfendlink ended up in different nesting level than \pdfstartlink.
%% See https://tex.stackexchange.com/a/249743
\makeatletter
\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
  \errmessage{\noexpand\@combinedblfloats could not be patched}%
}%
\makeatother

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\title
{Bow shocks, bow waves, and dust waves. II. Beyond the rip point}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Arthur]{
  William J. Henney \& S. J. Arthur\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

\input{bs-bw-dw-defs.tex}

\externaldocument[Q-]{quadrics-bowshock}
%\externaldocument[I-]{bs-bw-dw-01}
%\externaldocument[III-]{bs-bw-dw-03}

\defcitealias{Tarango-Yong:2018a}{Paper~I}
\newcommand\PaperI{\citetalias{Tarango-Yong:2018a}}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Dust waves are a result of gas--grain decoupling in a stream of
  dusty plasma that flows past a luminous star.  The radiation field
  is sufficiently strong to overcome the collisional coupling between
  grains and gas at a \textit{rip-point}, where the ratio of radiation
  pressure to gas pressure exceeds a critical value of roughly 1000.
  When the rip point occurs outside the hydrodynamic bow shock, a
  separate dust wave may form, decoupled from the gas shell, which can
  either be drag-confined or inertia-confined, depending on the stream
  density and relative velocity.  In the drag-confined case, there is
  a minimum stream velocity of roughly \SI{60}{km.s^{-1}} that allows
  a steady-state stagnant drift solution for the dust wave apex.  For
  lower relative velocities, the dust dynamics close to the axis
  exhibit a limit cycle behavior (rip and snap back) between two
  different radii.  Strong coupling of charged grains to the plasma's
  magnetic field can modify these effects, but for a quasi-parallel
  field orientation the results are qualitatively similar to the
  non-magnetic case. For a quasi-perpendicular field, on the other
  hand, the formation of a decoupled dust wave is strongly suppressed.
\end{abstract}

\begin{keywords}
  circumstellar matter -- radiation: dynamics -- stars: winds, outflows
\end{keywords}

\section{Introduction}
\label{sec:rip-introduction}

In \citet{Henney:2019a} (hereafter, Paper~I) we studied the formation
of bows around OB stars that are moving supersonically relative to
their surrounding medium.  For the case of strong collisional coupling
between gas and dust grains, we showed that three different regimes of
interaction are possible: a taxonomy of bow types

The plan of the paper is as follows.
%
In \S~\ref{sec:recap-paper-i} we give some results from Paper~I that
we will need in later sections.
%
In \S~\ref{sec:gas-free-bow} we calculate analytical models of the
shape of dust waves in the drag-free limit.
% %
% In \S\ref{sec:cloudy-models-dust} we discuss the relevant dust physics
% and the Cloudy models that we employ.
%
In \S~\ref{sec:imperf-coupl-betw} we calculate in detail the grain
charging and dynamics, subject to radiation and drag forces, in order
to determine the \textit{rip point}, which is where gas--grain coupling
catastrophically breaks down.  This is used to determine existence
conditions for the presence of a decoupled dust wave.
%
In \S~\ref{sec:magn-effects-grain} we consider the coupling between
the grains and the plasma's magnetic field, and how this effects the
existence and structure of dust waves.
%
In Appendix~\ref{sec:equat-moti-grains} we provide additional
technical information on the numerical calculation of grain
trajectories.

\section{Important results from Paper~I}
\label{sec:recap-paper-i}

%% XREF \eqref{eq:Rstar}
\begin{equation}
  \label{eq:Rstar}
  R_* = \left(\frac{L}{4\pi c \rho v^2}\right)^{1/2} \ .
\end{equation}

%% XREF \eqref{eq:tau-star},
\begin{equation}
  \label{eq:tau-star}
  \tau_* = \rho \kappa R_* \ .
\end{equation}

\begin{equation}
  \label{eq:wind-efficiency}
  \eta\wind  = \frac{L}{c\, \dot{M} V\wind} \ , 
\end{equation}

%% XREF \ref{eq:x-cases}
\begin{equation}
  \label{eq:x-cases}
  R_0 \approx
  \begin{cases}
    \text{RBS (\(\tau_* > 1\)):} & (1 + \eta\wind)^{1/2} \, R_*  \\
    \text{RBW (\(\eta\wind < \tau_* < 1\)):} & 2 \, \tau_* \, R_* \\
    \text{WBS (\(\tau_* < \eta\wind^{1/2}\)):} & \eta\wind^{1/2} \, R_*  
  \end{cases}
\end{equation}

\begin{align*}
  \label{eq:stellar-parameters}
  \dot{M}_{-7} &= \dot{M} / \bigl(\SI{e-7}{M_\odot.yr^{-1}}\bigr) \\
  V_3 &= V / \bigl(\SI{1000}{km.s^{-1}}\bigr) \\
  L_4 &= L / \bigl(\SI{e4}{L_\odot}\bigr) \\
  v_{10} &= v_\infty / \bigl( \SI{10}{km.s^{-1}} \bigr) \\
  n &= (\rho / \bar{m}) / \bigl( \SI{1}{cm^{-3}} \bigr) \\
  \kappa_{600} &= \kappa / \bigl( \SI{600}{cm^2.g^{-1}} \bigr) \ ,
\end{align*}
where \(\bar{m}\) is the mean mass per hydrogen nucleon
(\(\bar{m} \approx 1.3 m_{\text{p}} \approx \SI{2.17e-24}{g}\) for solar
abundances).

%% XREF \ref{eq:wind-eta-typical}
\begin{equation}
  \label{eq:wind-eta-typical}
  \eta\wind = \num{0.495} \,\dot{M}_{-7} \,V_3  \,L_4^{-1}
\end{equation}

%% XREF \ref{eq:taustar-typical}
%% XREF \ref{eq:Rstar-typical}
\begin{align}
  \label{eq:Rstar-typical}
  R_* / \si{pc} &= \num{2.21} \, (L_4 / n)^{1/2} \,v_{10}^{-1} \\
  \label{eq:taustar-typical}
  \tau_* &= \num{0.0089} \,\kappa_{600} \, (L_4 \,n)^{1/2} \,v_{10}^{-1} \ .
\end{align}

\input{app-dust-wave}

\input{sec-weak-coupling}

\section*{Acknowledgements}
We are grateful for financial support provided by Dirección General de
Asuntos del Personal Académico, Universidad Nacional Autónoma de
México, through grants Programa de Apoyo a Proyectos de Investigación
e Inovación Tecnológica IN112816 and IN107019.  This work has made
extensive use of Python language libraries from the SciPy
\citep{Jones:2001a} and AstroPy \citep{Astropy-Collaboration:2013a,
  Astropy-Collaboration:2018a} projects.

\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-dust-equations}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: