### 2019-03-11 papers/Papers1234/sec-dust-wave-discussion.tex

\section{Discussion}
\label{sec:discussion}

In molecular clouds:  \citet{Hopkins:2016a, Lee:2017a} say they
decouple. \citet{Tricco:2017a} says they don't.  But then
\citet{Mattsson:2019a} say they do.

In \citep{Akimkin:2015a} they used the drift approximation (ignored
grain inertia effects) and also ignore back reaction on gas. In
\citet{Akimkin:2017a} they do include back reaction.

Instabilities: Resonant Drag Instability \citep{Squire:2018a, Hopkins:2018a}

Survival of dust grains.  Sputtering \citep{Draine:2011a}.  Radiative Torque Disruption \citep{Hoang:2018a}.

\section{Summary}
\label{sec:summary}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "bs-bw-dw-02"
%%% End: