### 2019-03-17 papers/Papers1234/bs-bw-dw-03.tex

\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\sisetup{separate-uncertainty = true}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{xr-hyper}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{chemgreek}
\activatechemgreekmapping{newtx}
\usepackage{listings}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% The following hack solves a problem with
%% ERROR: \pdfendlink ended up in different nesting level than \pdfstartlink.
%% See https://tex.stackexchange.com/a/249743
\makeatletter
\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
  \errmessage{\noexpand\@combinedblfloats could not be patched}%
}%
\makeatother

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\title
{Bow shocks, bow waves, and dust waves. III. Diagnostics}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Arthur]{
  William J. Henney \& S. J. Arthur\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

\input{bs-bw-dw-defs.tex}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Stellar bow shocks, bow waves, and dust waves all result from the
  action of a star's wind and radiation pressure on a stream of dusty
  plasma that flows past it.  The dust in these bows emits prominently
  at mid-infrared wavelengths in the range \SIrange{8}{60}{\um}. We
  propose a novel diagnostic method, the \(\tau\)--\(\eta\) diagram, for
  analyzing these bows, which is based on comparing the fractions of
  stellar radiative energy and stellar radiative momentum that is
  trapped by the bow shell.  This diagram allows the discrimination of
  radiation-supported bows from wind-supported bows.  For the latter,
  it allows the stellar wind mass loss rate to be determined.  We
  critically compare our method with a previous method that has been
  proposed for determining wind mass loss rates from bow shock
  observations. We find that the dust emissivity model used previously
  needs to be corrected to account for differences in the spectral
  energy distribution between OB stars and the interstellar radiation
  field, which results in a downward revision by a factor of two in
  the mass loss rates.  After correction, the methods complement one
  another to some degree, with differing sensitivities to systematic
  model uncertainties.  We point out that observations of bows are of
  limited utility in constraining mass loss rates of weak-wind stars,
  since these tend to have radiation-supported bows, which yield only
  upper limits to the mass loss rate.
\end{abstract}

\begin{keywords}
  circumstellar matter -- radiation: dynamics -- stars: winds, outflows
\end{keywords}

\section{Introduction}
\label{sec:introduction}

In \citet{Henney:2019a, Henney:2019b} (Paper~I and Paper~II) we
presented a taxonomy of stellar bows, which we divided into
wind-supported bow shocks (WBS) and various classes of
radiation-supported bows.  When the dust and gas remains well-coupled
(Paper~I), these are optically thin radiation-supported bow waves
(RBW) and optically thick radiation-supported bow shocks (RBS).  When
the dust decouples from the gas (Paper~II), one has inertia-confined
dust waves (IDW) and drag-confined dust waves (DDW).

In Paper~I we derived expressions for the bow radius \(R_0\)in each of
the three well-coupled regimes as a function of the parameters of the
star and the ambient medium.  In Paper~II, we found the criteria for
the dust to decouple from the gas to form a separate dust wave outside
of the hydrodynamic bow shock.  The most important of these is that
the ratio of radiation pressure to gas pressure should exceed a
critical value, \(\Xi_\dag \sim 1000\).

In order to provide an empirical anchor to our theoretical
calculations, we now consider how the parameters of our models might
be determined from observations.  The parameter-space diagrams, such
as Figure~2 of Paper~I, are not always useful in this regard, since in
many cases the ambient density and relative stellar velocity are not
directly measured.  Instead, we aim to construct diagnostics based on
the most common observations, which are of the infrared dust emission.
Key questions that we wish to address include
\begin{enumerate}[1.]
\item Can we distinguish observationally between radiation support
  (bow waves and dust waves) and wind support (bow shocks)?
\item Are there any clear examples of sources with radiation-supported bows?
\item In the case of wind support, can we reliably determine mass loss
  rates from mid-infrared observations?
\end{enumerate}

The outline of the paper is as follows. In
\S~\ref{sec:energy-trapp-vers} we describe how the optical depth and
the gas pressure in the bow shell can be estimated from a small number
of observed quantities.  In \S~\ref{sec:eta-tau-diagnostic} we place
observed bow sources on a diagnostic diagram of these two quantities
and discuss the influence of observational errors and systematic model
uncertainties. In \S~\ref{sec:cand-radi-supp} we use the diagram to
identify some candidates for radiation-supported bows. In
\S~\ref{sec:grain-temp-emiss} we calculate the grain emissivity as a
function of the stellar radiation field around OB stars.  In
\S~\ref{sec:stellar-wind-mass} we compare two different methods of
estimating the stellar wind mass loss rates from observations of the
bows.  In \S~\ref{sec:conclusions} we summarise out findings.

\input{sec-observational-diagnostics}
\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-k17-k18}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: