### 2019-03-22 papers/Papers1234/bs-bw-dw-03.tex
Lots more writing of discussion for Paper III

\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\sisetup{separate-uncertainty = true}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{xr-hyper}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{chemgreek}
\activatechemgreekmapping{newtx}
\usepackage{listings}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% The following hack solves a problem with
%% ERROR: \pdfendlink ended up in different nesting level than \pdfstartlink.
%% See https://tex.stackexchange.com/a/249743
\makeatletter
\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
  \errmessage{\noexpand\@combinedblfloats could not be patched}%
}%
\makeatother

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\title
{Bow shocks, bow waves, and dust waves. III. Diagnostics}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney \& Arthur]{
  William J. Henney \& S. J. Arthur\\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

\input{bs-bw-dw-defs.tex}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Stellar bow shocks, bow waves, and dust waves all result from the
  action of a star's wind and radiation pressure on a stream of dusty
  plasma that flows past it.  The dust in these bows emits prominently
  at mid-infrared wavelengths in the range \SIrange{8}{60}{\um}. We
  propose a novel diagnostic method, the \(\tau\)--\(\eta\) diagram, for
  analyzing these bows, which is based on comparing the fractions of
  stellar radiative energy and stellar radiative momentum that is
  trapped by the bow shell.  This diagram allows the discrimination of
  radiation-supported bows from wind-supported bows.  For the latter,
  it allows the stellar wind mass loss rate to be determined.  We
  critically compare our method with a previous method that has been
  proposed for determining wind mass loss rates from bow shock
  observations. We find that the dust emissivity model used previously
  needs to be corrected to account for differences in the spectral
  energy distribution between OB stars and the interstellar radiation
  field, which results in a downward revision by a factor of two in
  the mass loss rates.  After correction, the methods complement one
  another to some degree, with differing sensitivities to systematic
  model uncertainties.  We point out that observations of bows are of
  limited utility in constraining mass loss rates of weak-wind stars,
  since these tend to have radiation-supported bows, which yield only
  upper limits to the mass loss rate.
\end{abstract}

\begin{keywords}
  circumstellar matter -- radiation: dynamics -- stars: winds, outflows
\end{keywords}

\section{Introduction}
\label{sec:introduction}

In \citet[Paper~I and Paper~II]{Henney:2019a, Henney:2019b} we
presented a taxonomy of stellar bows, which we divided into
wind-supported bow shocks (WBS) and various classes of
radiation-supported bows.  When the dust and gas remains well-coupled
(Paper~I), these are optically thin radiation-supported bow waves
(RBW) and optically thick radiation-supported bow shocks (RBS).  When
the dust decouples from the gas (Paper~II), one has inertia-confined
dust waves (IDW) and drag-confined dust waves (DDW).

In Paper~I we derived expressions for the bow radius \(R_0\) in each of
the three well-coupled regimes as a function of the parameters of the
star and the ambient medium.  In Paper~II, we found the criteria for
the dust to decouple from the gas to form a separate dust wave outside
of the hydrodynamic bow shock.  The most important of these is that
the ratio of radiation pressure to gas pressure should exceed a
critical value, \(\Xi_\dag \sim 1000\).

In order to provide an empirical anchor to our theoretical
calculations, we now consider how the parameters of our models might
be determined from observations.  The parameter-space diagrams, such
as Figure~2 of Paper~I, are not always useful in this regard, since in
many cases the ambient density and relative stellar velocity are not
directly measured.  Instead, we aim to construct diagnostics based on
the most common observations, which are of the infrared dust emission.
Key questions that we wish to address include
\begin{enumerate}[1.]
\item Can we distinguish observationally between radiation support
  (bow waves and dust waves) and wind support (bow shocks)?
\item Are there any clear examples of sources with radiation-supported bows?
\item In the case of wind support, can we reliably determine mass loss
  rates from mid-infrared observations?
\end{enumerate}

The outline of the paper is as follows. In
\S~\ref{sec:energy-trapp-vers} we describe how the optical depth and
the gas pressure in the bow shell can be estimated from a small number
of observed quantities.  In \S~\ref{sec:eta-tau-diagnostic} we place
observed bow sources on a diagnostic diagram of these two quantities
and discuss the influence of observational errors and systematic model
uncertainties. In \S~\ref{sec:cand-radi-supp} we use the diagram to
identify some candidates for radiation-supported bows. In
\S~\ref{sec:grain-temp-emiss} we calculate the grain emissivity as a
function of the stellar radiation field around OB stars.  In
\S~\ref{sec:stellar-wind-mass} we compare two different methods of
estimating the stellar wind mass loss rates from observations of the
bows.  In \S~\ref{sec:conclusions} we summarise out findings.

\input{sec-observational-diagnostics}

\section{Summary}
\label{sec:conclusions}

\section*{Acknowledgements}
We are grateful for financial support provided by Dirección General de
Asuntos del Personal Académico, Universidad Nacional Autónoma de
México, through grants Programa de Apoyo a Proyectos de Investigación
e Inovación Tecnológica IN112816 and IN107019.  This work has made
extensive use of Python language libraries from the SciPy
\citep{Jones:2001a}, AstroPy \citep{Astropy-Collaboration:2013a,
  Astropy-Collaboration:2018a} and Pandas \citep{McKinney:2010a}
projects.  We thank Chip Kobulnicky for useful discussions.

\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-k17-k18}

\section{Notes on particular sources}
\label{sec:notes-part-sourc}

\subsection{Source~380, HD~53367, V750~Mon}
\label{sec:hd-53367-v750}

This is a Herbig Be star with spectral type B0V--B0III and mass 12 to
\SI{15}{M_\odot}, which shows long-scale irregular photometric
variability \citep{Tjin-A-Djie:2001a, Pogodin:2006a} together with
cyclic radial velocity variations, which are interpreted as an
eccentric binary with a \SI{5}{M_\odot} pre-main-sequence companion.
It is located outside the solar circle and, although it was originally
classified as part of the CMa OB association at about \SI{1}{kpc}
\citep{Tjin-A-Djie:2001a}, more recent estimates put it much closer.
K18 assume a distance of \SI{260}{pc}, based on Hipparcos measurements
\citep{van-Leeuwen:2007a}, but its Gaia DR2 parallax
\citep{Gaia-Collaboration:2016a, Gaia-Collaboration:2018a, Luri:2018a}
puts it closer still at \SI{140 \pm 30}{pc}.\footnote{%
  Median and 90\% confidence interval, estimated from Bayesian
  inference using an exponential density distribution with scale
  length of \SI{1350}{pc} as a prior (see
  \url{https://github.com/agabrown/astrometry-inference-tutorials/}). }

\citet{Quireza:2006b} report a kinematic distance to the associated
\hii{} region IC~2177 (G\num{223.70}\num{-1.90}) of \SI{1.6}{kpc},
based an LSR radio recombination line velocity of \SI{+16}{km.s^{-1}}
\citep{Quireza:2006a} and the outer Galaxy rotation curve of
\citet{Brand:1993a}.  However, given the dispersion in peculiar
velocities of star-forming clouds (\SIrange{7}{9}{km.s^{-1}},
\citealp{Stark:1984a}) and likely streaming motions of the ionized gas
(\(\approx \SI{10}{km.s^{-1}}\), \citealp{Matzner:2002a, Lee:2012a}),
this is still consistent with a much smaller distance, which would
also help in bringing the radio continuum-derived nebular electron
density into agreement with the value (\(\approx \SI{100}{cm^{-3}}\))
derived from the optical [\ion{S}{ii}] line ratio
\citep{Hawley:1978a}.

\citet{Fairlamb:2015a} derive a distance of \SI{340 \pm 60}{pc} from
combining a spectroscopically determined effective temperature,
gravity, and reddening with photometry and pre-main sequence
evolutionary tracks.  They also determine a luminosity of \SI{13000
  \pm 1000}{L_\odot}, which is only half that assumed by K18, and
would have to be even lower to bring the photometry into concordance
with the Gaia distance.  Alternatively, the luminosity could remain
the same if the total-to-selective extinction ratio were higher than
the \(R_V = 3.1\) assumed by \citet{Fairlamb:2015a}.  Taking
\(R_V = 5.5\) instead, as in Orion, would give \(A_V = 3.34\) and a
predicted \(V\) magnitude of 6.7 if we assume \(L_4 = 1.3\),
\(D = \SI{170}{pc}\) (furthest distance in Gaia 90\% confidence
interval) and \(T_{\text{eff}} = \SI{29.5}{kK}\)
\citep{Fairlamb:2015a} for a bolometric correction of \(-2.86\)
\citep{Nieva:2013a}.  The observed brightness varies between
\(V = 6.9\) and 7.2, which is still at least 20\% fainter than
predicted, but this is the best that can be achieved without rejecting
the Gaia parallax distance entirely.

A final sanity check can be performed by considering the free-free
radio continuum flux of V750~Mon's surrounding \hii{} region, which,
after converting to a luminosity, should be proportional to the total
recombination rate in the nebula and therefore, assuming
photoionization equilibrium and negligible dust absorption in the
ionized gas, also proportional to the ionizing photon luminosity of
the star.  \citet{Quireza:2006b} report a flux density of \SI{6}{Jy}
at \SI{8.6}{Ghz} for G\num{223.70}\num{-1.90}, as compared with a flux
density of \SI{260}{Jy} for the Orion Nebula using the same
instrumental setup.  Assuming an ionizing photon luminosity of
\(S_{49} = 1\) and distance \SI{410}{pc} for the ionizing Trapezium
stars in Orion, therefore implies
\(S_{49} = 0.0027 (D/\SI{140}{pc})^2\) for V750~Mon.  Using an
ionizing flux of \SI{3.2e22}{cm^{-2}.s^{-1}} from the curves in Fig.~4
of \citet{Sternberg:2003a}, this translates to a bolometric luminosity
of \(L_4 = 0.96 (D/\SI{140}{pc})^2\), which is consistent with the
\citet{Fairlamb:2015a} value if we take a distance towards the high
end of the Gaia range.

\subsection{Source 407, HD~93249}
\label{sec:source-407}

TODO

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/cygnus-bows-midir-combo}
  \caption{Two bows in Cygnus}
  \label{fig:cygnus-bows}
\end{figure}

\subsection{Source 342}
\label{sec:source-342}

Source~342 is anomalous in both methods. Has high mass loss for low luminosity.

\subsection{Source 344}
\label{sec:source-344}

This has big disagreement between K18 and me.  My mass-loss rate seems too low, given its high luminosity. Why?

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: