### 2019-03-22 papers/Papers1234/sec-observational-diagnostics.tex

% Are there any objects that might not be bow shocks?

% Progression in density:
% \begin{gather*}
%   \text{Increasing density} \longrightarrow \\
%   \text{WBS} \to \text{WBS} + \text{IDW} \to \text{WBS} + \text{DDW} \to \text{(RBW)} \to \text{RBS}
% \end{gather*}

% Chief diagnostic for radiation supported bows (RBW or RBS cases) is
% infrared luminosity of bow.  Favored by high densities.

% Dust waves favored by high velocities and intermediate densities.

\section{Optical depth and pressure of the bow shell}
\label{sec:energy-trapp-vers}

\begin{figure*}
  \centering
  \includegraphics[width=0.8\linewidth]{figs/All-sources-eta-tau}
  \caption[Observational diagnostic diagram]{Observational diagnostic
    diagram for bow shocks.  The shell optical depth \(\tau\) (\(x\)
    axis) and momentum efficiency \(\eta\shell\) (\(y\) axis) can be
    estimated from observations of the bolometric stellar luminosity,
    infrared shell luminosity, and shell radius, as described in the
    text.  Results are shown for the 20 sources (circle symbols) from
    \citet{Kobulnicky:2018a} plus three further sources (square
    symbols), where we have obtained the measurements ourselves (see
    Tab.~\ref{tab:observations}).  The color of each symbol indicates
    the stellar luminosity (dark to light) as indicated by the scale
    bar. The shell pressure is determined assuming a gas temperature
    \(T = \SI{e4}{K}\), an absorption opacity
    \(\kappa = \SI{600}{cm^2.g^{-1}}\), and a thickness-to-radius ratio
    \(H/R = 0.25\).  The sensitivity of the results to a
    factor-of-three change in each parameter is shown in the upper
    inset box.  Exceptions are the two Orion Nebula sources, \thD{}
    and LP~Ori, where the small dim squares show the results of
    assuming the standard shell parameters, while the large squares
    show the results of modifications according to the peculiar
    circumstances of each object, as described in the text.  The lower
    inset box shows the sensitivity of the results to a factor-of-two
    uncertainty in each observed quantity: distance to source \(D\);
    stellar luminosity \(L_*\), shell infrared flux \(F\IR\); shell
    angular size \(\theta\).  Lines and shading indicate different
    theoretical bow regimes (see Paper~I and Paper~II).  The dashed
    blue diagonal line corresponds to radiation-supported bows, while
    the upper left region corresponds to wind-supported bows.  The
    upper right corner (purple) corresponds to optically thick bow
    shocks, while the lower left corner (yellow) is the region where
    grain--gas separation \textit{may} occur, leading to a potential
    dust wave.  However, the existence of a dust wave in this region
    is not automatic, since it only includes one of the four necessary
    conditions (\S\S~4.4 and 5.1 of Paper~II). The lower-right region
    is strictly forbidden, except in case of violation of the
    assumption that dust heating be dominated by stellar radiation. }
  \label{fig:All-sources-eta-tau}
\end{figure*}

A fundamental parameter is the optical depth, \(\tau\), of the bow
shell to UV radiation, which determines what fraction of the stellar
photon momentum is available to support the shell (see
\S~2.1 of Paper~I).  But the same photons also heat the
dust grains in the bow, which re-radiate that energy predominantly at
mid-infrared wavelengths (roughly \SIrange{10}{100}{\um}) with
luminosity \(L\IR\).  Assuming that Ly\(\alpha\) and mechanical
heating of the dust shell is negligible (see
\S~\ref{sec:unimp-other-heat}) and that the emitting shell subtends a
solid angle \(\Omega\), as seen from the star, then the optical depth
can be estimated as
\begin{equation}
  \label{eq:tau-empirical}
  \tau = -\ln \left( 1 - \frac{4\pi}{\Omega} \frac{L\IR}{L_*} \right)
  \approx \frac{2 L\IR}{L_*} \ ,
\end{equation}
where the last approximate equality holds if \(\tau \ll 1\) and the shell
emission covers one hemisphere.  Note that the \(\tau\) of Paper~I is not
exactly the same as the \(\tau\) of equation~\eqref{eq:tau-empirical},
but is larger by a factor of
\(Q_P / Q_{\text{abs}} = 1 + \varpi (1 - g)/(1 - \varpi)\), where
\(\varpi\) is the grain albedo and \(g\) the scattering asymmetry.
This is plotted in Figure~6a of Paper~II for Cloudy's default ISM
grain mixture (see \S~\ref{sec:grain-temp-emiss} below), where it can
be seen that \(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at
EUV/FUV wavelengths (\(\lambda = \text{\SIrange{0.05}{0.2}{\um}}\)), so the
correction is small.

A second important parameter is the thermal plus magnetic pressure in
the shocked shell, which is doubly useful since in a steady state it
is equal to \emph{both} the internal supporting pressure (wind ram
pressure plus absorbed stellar radiation) \emph{and} the external
confining pressure (ram pressure of ambient stream).  The shell
pressure is not given directly by the observations, but can be
determined by the following three steps:
\begin{enumerate}[P1.]
\item \label{P1} The shell mass column (\si{g.cm^{-2}}) can be
  estimated from the optical depth by assuming an effective UV
  opacity: \(\Sigma\shell = \tau / \kappa\)
\item \label{P2} The shell density (\si{g.cm^{-3}}) can be found from
  the mass column if the shell thickness is known:
  \(\rho\shell = \Sigma / h\shell\).  In the absence of other information, a
  fixed fraction of the shell radius can be used.  In particular, we
  normalize by a typical value of one~quarter the star--apex distance:
  \(h_{1/4} = h\shell / (0.25 R_0)\).  This corresponds to a Mach
  number \(\M_0 = \surd 3\) if the stream shock is radiative, or
  \(\M_0 \gg 1\) if non-radiative (see \S~3.2 of Paper~I). Further
  discussion is given in \S~\ref{app:bow-shock-data} below. 
\item \label{P3} Finally, the pressure (\si{dyne.cm^{-2}}) follows by
  assuming values for the sound speed and Alfvén speed:
  \(P\shell = \rho\shell (\sound^2 + \frac12 v\alfven^2) \).
\end{enumerate}
It is natural to normalize this pressure to the stellar radiation
pressure at the shell, so we define a shell momentum efficiency
\newcommand\pc{\ensuremath{_{\text{pc}}}}
\begin{equation}
  \label{eq:eta-shell}
  \eta\shell \equiv \frac{P\shell}{P\rad}
  = \frac{4\pi R_0^2\, (\sound^2 + \frac12 v\alfven^2)\, \tau}{L_*\, \kappa\, h\shell}
  \approx 245 \frac{R\pc \, T_4 \, \tau}{L_4 \, \kappa_{600} \, h_{1/4}} \ , 
\end{equation}
where in the last step we have assumed ionized gas with negligible
magnetic support (\(v\alfven \ll \sound\)) and written the stellar
luminosity and shell parameters in terms of typical values, which we
summarize below.  Note that the shell momentum efficiency is simply
the reciprocal of the radiation parameter from Paper~II's equation~(23):
\(\eta\shell = \Xi\shell^{-1}\), which provides yet a third use for
\(\eta\shell\), since \(\Xi\) is paramount in determining whether the
grains and gas remain well-coupled (see \S~4.4 of Paper~II).

In this section and the remainder of the paper, we employ
dimensionless versions of the stellar bolometric luminosity, \(L_*\),
wind mass-loss rate, \(\dot{M}\), and terminal velocity, \(V\wind\),
together with the ambient stream's mass density, \(\rho\), relative
velocity \(v_\infty\), and effective dust opacity, \(\kappa\).  These are
defined as follows:
\begin{align*}
  % \label{eq:stellar-parameters}
  \dot{M}_{-7} &= \dot{M} / \bigl(\SI{e-7}{M_\odot.yr^{-1}}\bigr) \\
  V_3 &= V\wind / \bigl(\SI{1000}{km.s^{-1}}\bigr) \\
  L_4 &= L_* / \bigl(\SI{e4}{L_\odot}\bigr) \\
  v_{10} &= v_\infty / \bigl( \SI{10}{km.s^{-1}} \bigr) \\
  n &= (\rho / \bar{m}) / \bigl( \SI{1}{cm^{-3}} \bigr) \\
  \kappa_{600} &= \kappa / \bigl( \SI{600}{cm^2.g^{-1}} \bigr) \ ,
\end{align*}
where \(\bar{m}\) is the mean mass per hydrogen nucleon
(\(\bar{m} \approx 1.3 m_{\text{p}} \approx \SI{2.17e-24}{g}\) for solar
abundances).

\section[The eta-tau diagnostic diagram]
{\boldmath The \(\eta\shell\)--\(\tau\) diagnostic diagram}
\label{sec:eta-tau-diagnostic}

\begin{table}
  \centering
  \caption[Observational]{Key observational parameters for star/bow systems}
  \label{tab:observations}
  \begin{tabular}{l S S S}
    \toprule
    Star & {\(L_* / \si{L_\odot}\)} & {\(L_{\text{IR}} / \si{L_\odot}\)} & {\(R_0 / \si{pc}\)} \\
    \midrule
    \thD & 2.95e4 & 620 & 0.003 \\
    LP~Ori & 1600 & 240 & 0.01 \\
    \(\sigma\)~Ori & 6e4 & 15 & 0.12 \\[\smallskipamount]
    K18 Sources & \numrange{1.4e4}{8.7e5} & \numrange{8}{2800} & \numrange{0.02}{1.35} \\
    \bottomrule
  \end{tabular}
\end{table}

In Figure~\ref{fig:All-sources-eta-tau} we show the resultant
diagnostic diagram: \(\eta\shell\) versus \(\tau\).  The horizontal axis
shows the fraction of the stellar radiative \emph{energy} that is
reprocessed by the bow shell, while the vertical axis shows the
fraction of stellar radiative \emph{momentum} that is imparted to the
shell, either directly by absorption, or indirectly by the stellar
wind (which is itself radiatively driven).  Radiatively supported bows
(DW, RBW, or RBS, or cases) should lie on the diagonal line
\(\eta\shell = ( Q_P / Q_{\text{abs}}) \tau \approx 1.25 \tau\), where we have used
the ratio of grain radiation pressure efficiency to absorption
efficiency found in the FUV band for the dust mixture shown in
Paper~II's Figure~6.  Wind-supported bows should lie above this line
and no bows should lie below the \(\eta\shell = \tau\) line, since
\(Q_P\) cannot be smaller than \(Q_{\text{abs}}\).

We have calculated \(\eta\shell\) and \(\tau\) using the above-described
methods for the 20 mid-infrared sources studied by
\citet{Kobulnicky:2018a} (K18) and plotted them on our diagnostic
diagram.  Details of our treatment of this observational material are
provided in the following subsection.  In order to expand the range of
physical conditions, we have included three additional sources (data
in Table~\ref{tab:observations}): bows around \thD{}
\citep{Smith:2005a} and LP~Ori \citep{ODell:2001c} in the Orion
Nebula, which show larger optical depths, plus the inner bow around
\(\sigma\)~Ori, which illuminates the Horsehead Nebula and has previously
been claimed to be a dust wave \citep{Ochsendorf:2014b,
  Ochsendorf:2015a}.  Details of the observations of these additional
sources will be published elsewhere.

\subsection{Treatment of sources from \citeauthor{Kobulnicky:2018a}}
\label{sec:kobulnicky}

In a series of papers \citeauthor{Kobulnicky:2018a} provide an
extensive mid-infrared-selected sample of over 700 candidate stellar
bow shock nebulae (\citealp{Kobulnicky:2016a, Kobulnicky:2017a,
  Kobulnicky:2018a}, hereafter K16, K17, and K18).  For 20 of these
sources, reliable distances and spectral classifications are provided
in Table~5 of K17 and Tables~1 and~2 of K18. In this section, we
outline how we obtain \(\tau\) and \(\eta\shell\) from the data in these
catalogs, while further aspects of the \citeauthor{Kobulnicky:2018a}
material are discussed in \S~\ref{app:bow-shock-data}.

The UV optical depth of the bow shell is obtained
(eq.~[\ref{eq:tau-empirical}]) from the ratio of infrared shell
luminosity to stellar luminosity.  The inverse of this ratio is given
in Table~5 of K17, but we choose to re-derive the values since the
spectral classification of some of the sources was revised between K17
and K18.  Although K17 found the total shell fluxes from fitting dust
emission models to the observed SEDs, we adopt the simpler approach of
taking a weighted sum of the flux densities \(F_\nu\) (in \si{Jy}) in
three mid-infrared bands:
\begin{multline}
  \label{eq:total-ir-flux}
  F_{\text{IR}}  \approx \bigl[  2.4\,(F_8 \text{ or } F_{12})
    + 1.6\,(F_{22} \text{ or } F_{24})  \\
  + 0.51\,F_{70}\bigr]
  \,\times \SI{e-10}{erg.s^{-1}.cm^{-2}} \ ,
\end{multline}
where \(F_8\) is Spitzer IRAC \SI{8.0}{\um}, \(F_{24}\) is Spitzer
MIPS \SI{23.7}{\um}, \(F_{12}\) and \(F_{22}\) are WISE bands 3 and 4,
and \(F_{70}\) is Herschel PACS \SI{70}{\um}.  The weights are chosen
so that the integral \(\int_0^\infty F_\nu \,d\nu\) is approximated by the sum
\(\Sigma_k F_k\, \Delta\nu_k\), under the assumption that fluxes in shorter (e.g.,
IRAC \SI{5.8}{\um}) and longer (e.g., PACS \SI{150}{\um}) wavebands
are negligible.  Shell fluxes are converted to luminosities using the
assumed distance to each source, and stellar luminosities are taken
directly from K18 Table~2, based on spectroscopic classification and
the calibrations of gravity and effective temperature from
\citet{Martins:2005a}.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/K17-tau-comparison}
  \caption{Comparison between shell-to-star luminosity ratios
    calculated as described in the text (\(y\) axis) with those given
    in K17 (\(x\) axis).  The blue dashed line signifies equality and
    the gray band shows ratios between 1/2 and 2.}
  \label{fig:k17-k18-comparison}
\end{figure}

In Figure~\ref{fig:k17-k18-comparison} we compare the \(\tau\) obtained
using the shell luminosity as described above with that obtained using
the luminosity ratios directly from K17 Table~5.  It can be seen that
for the majority of sources the two measurements are consistent within
a factor of two (gray band).  The four furthest-flung outliers can be
understood as follows:
\begin{description}
\item[\textit{Source 67}] This has a very poor-quality spectral fit in
  K17 (see lower left panel of their Fig.~12) and so 
  \(F_{\text{IR}}\) is overestimated by them by a factor of 10.
\item[\textit{Sources 341 and 342}] The spectral classes changed from
  B2V in K17 to O9V and B1V, respectively, in K18, increasing the
  derived \(L_*\), which lowers \(\tau\).
\item[\textit{Source 411}] The luminosity class changed from Ib (K17)
  to V (K18), so \(L_*\) has been greatly reduced, which increases
  \(\tau\).
\end{description}

\subsection{Random uncertainties due to observational errors}
\label{sec:rand-syst-uncert}

The fundamental observational quantities that go into determining
\(\tau\) and \(\eta\shell\) for each source are distance, \(D\); stellar
luminosity, \(L_*\); total infrared flux, \(F_{\text{IR}}\); and bow
angular apex distance, \(\theta\).  From these, the shell radius and
infrared luminosity are found as \(R_0 = \theta D\) and
\(L_{\text{IR}} = 4\pi D^2 F_{\text{IR}}\).  Rather than clutter the
diagram with error bars, we instead show the sensitivity to
observational errors in the lower-right box, where each arrow
corresponds to a factor of two increase (0.3~dex) in each quantity:
\(D\), \(L_*\), \(F_{\text{IR}}\), and \(\theta\).  We now calculate
uncertainty estimates for individual observational quantities that are
used in deriving not only \(\tau\) and \(\eta\shell\), but also mass-loss
rates, as discussed later in \S~\ref{sec:stellar-wind-mass}.

\subsubsection{Distance}
\label{sec:distance}

Most sources are members of known high-mass clusters with distance
uncertainties less than 20\% (0.08~dex). The only exception is
Source~329 in Cygnus, for which the distance uncertainty is roughly a
factor of~2 \citep{Kobulnicky:2018a}.

\subsubsection{Stellar luminosity}
\label{sec:stellar-luminosity}

The stellar luminosity is determined from spectral classification,
which makes it independent of distance.  Taking a \SI{2000}{K}
dispersion in the effective temperature scale \citep{Martins:2005a}
gives an uncertainty of 25\% in the luminosity, and adding in possible
errors in gravity and the effect of binaries, we estimate a total
uncertainty in \(L_*\) of 50\% (0.45~mag or 0.18~dex).

\subsubsection{Shell flux and surface brightness}
\label{sec:shell-flux-surface}

We estimate the uncertainty in shell bolometric flux,
\(F_{\text{IR}}\), by comparing two different methods: model fitting
\citep{Kobulnicky:2017a} and a weighted sum of the 8, 24, and
\SI{70}{\um} bands (eq.~[\ref{eq:total-ir-flux}]), giving a standard
deviation of 17\% (0.07~dex).  To this, we add the estimate of 25\%
for the effects of background subtraction uncertainties on individual
photometric measurements \citep{Kobulnicky:2017a}. The absolute flux
calibration uncertainty for both Herschel PACS \citep{Balog:2014a} and
Spitzer MIPS \citep{Engelbracht:2007a} is less than 5\%, which is
small in comparison. Combining the 3 contributions in quadrature gives
a total uncertainty of 0.12~dex.  We adopt the same uncertainty for
the \SI{70}{\um} surface brightness.

\subsubsection{Angular sizes}
\label{sec:angular-sizes}

For the angular apex distance, \(\theta\), the largest uncertainty for
well-resolved sources is due to the unknown inclination.
\citet{Tarango-Yong:2018a} show that the dispersion in true to
projected distances can introduce an uncertainty of \(30\%\)
(0.11~dex) in unfavorable cases (e.g., their Fig.~26).  For 5 of the
20 sources from \citet{Kobulnicky:2018a}, \(\theta\) is of order the
Spitzer PSF width at \SI{24}{\um}, so the errors may be larger.

\subsubsection{Stellar wind velocity}
\label{sec:stell-wind-veloc}

Although this is not strictly an observed quantity for the K18 sample,
we will treat it as such since it is estimated per star, based on the
spectral type.  K18 estimate 50\% uncertainty, and we adopt the same
here (0.18~dex).

\subsubsection{Combined effect of uncertainties on the
  \(\tau\)--\(\eta\shell\) diagram}
\label{sec:comb-effect-uncert}

Assuming that the uncertainty in each observational quantity is
independent, we can now combine them using the techniques described in
Appendix~\ref{sec:comb-uncert-covar} to find the \(\pm 1~\sigma\) error
ellipse, shown in blue in the figure.  It can be seen that
observational uncertainties in \(\tau\) and \(\eta\shell\) are highly
correlated: the dispersion is \SI{0.7}{dex} in the product
\(\eta\shell \tau\) but only \SI{0.16}{dex} in the ratio
\(\eta\shell/\tau\), with stellar luminosity errors dominating in both
cases.  Observational uncertainties are therefore relatively
unimportant in determining whether a given source is wind-driven or
radiation-driven, which depends only on \(\eta\shell/\tau\).  On the other
hand, they do significantly effect the question of whether a source
has a sufficiently high radiation parameter \(\Xi\) to possibly be a
dust wave.

\subsection{Systematic uncertainties due to assumed shell parameters}
\label{sec:syst-uncert-due}

A further source of uncertainty arises from the parameters of the
shocked shell that are assumed in steps P\ref{P1}--P\ref{P3}. Namely,
the relative shell thickness, \(h\shell/R_0\), the ultraviolet grain
opacity per mass of gas, \(\kappa\), and the shell temperature, \(T\).
These parameters effect only \(\eta\shell\), not \(\tau\), with a
sensitivity shown by arrows in the upper left box of
Figure~\ref{fig:All-sources-eta-tau}.

\subsubsection{Shell thickness}
\label{sec:shell-thickness}
For maximally efficient post-shock radiative cooling, the shell
thickness depends on the Mach number of the shock as
\(h / R_0 \sim M_0^{-2}\).  However, for ambient densities less than
about \SI{10}{cm^{-3}}, the minimum thickness is only about ten times
smaller than the \(h / R_0 = 0.25\) that we are assuming.  In
photoionized gas, this occurs at \(v \approx \SI{60}{km.s^{-1}}\),
corresponding to the peak in the cooling curve at \SI{e5}{K} (see
\S~3.2 of Paper~I),
% XREF: Paper I \S~\ref{sec:radi-cool-lengths}
since the thickness is set by the cooling length at higher speeds.  In
the case that the Alfv\'en speed is a significant fraction of the
sound speed, this will also tend to increase the thickness.  In
principle, the shell thickness can be measured observationally if the
source is sufficiently well resolved \citep{Kobulnicky:2017a},
although this is complicated by projection effects.  We return to the
issue of the shell thickness in the discussion below.

\subsubsection{Dust opacity}
\label{sec:dust-opacity}
The dust opacity will depend on the total dust-gas ratio and on the
composition and size distribution of the grains.  Our adopted value of
\SI{600}{cm^2.g^{-1}}, or \SI{1.3e-21}{cm^2.H^{-1}}, is appropriate
for average Galactic interstellar grains in the EUV and FUV (e.g.,
\citealp{Weingartner:2001a}), but there is ample evidence for
substantial spatial variations in grain extinction properties
\citep{Fitzpatrick:2007a}, both on Galactic scales
\citep{Schlafly:2016a} and within a single star forming region
\citep{Beitia-Antero:2017a}.  The properties of grains within
photoionized regions are very poorly constrained observationally
because the optical depth is generally much lower than in overlying
neutral material.  In the Orion Nebula, there is some evidence
\citep{Salgado:2016a} that the FUV dust opacity in the ionized gas may
be as low as \SI{90}{cm^2.g^{-1}}, although the uncertainties in this
estimate are large and different results are obtained in other
regions, such as W3(A) \citep{Salgado:2012a}.  It is even possible
that the FUV dust opacity may be larger than the ISM value if the
abundance of very small grains is enhanced through radiative torque
disruption of larger grains \citep{Hoang:2018a}.

\subsubsection{Shell gas temperature}
\label{sec:shell-gas-temp}
For bows around O~stars, the shell temperature should be close to the
photoionization equilibrium value of \(\approx \SI{e4}{K}\), since the
post-shock cooling length is short in ambient densities above
\SI{0.1}{cm^{-3}} (\S~3.2 of Paper~I)
% XREF: Paper I \ref{sec:radi-cool-lengths}
and the shell does not trap the ionization front for ambient densities
below \SI{e4}{cm^{-3}} (\S~3.1 of Paper~I).
% XREF: Paper I \\ref{sec:trapp-ioniz-front}
For B~stars, on the other hand, these two density limits move closer
together (cf.~the smaller gap between the blue and red lines in
% XREF: Paper I \ref{fig:zones-v-n-plane}a
% XREF: Paper I \ref{fig:B-supergiant}
% XREF: Paper I \ref{fig:zones-v-n-plane}bc
Paper~I's Figs.~2a and 4, as compared with 2b and~c), making it more
likely that a bow will lie in a different temperature regime.  The
only source for which we have evidence that this occurs is LP~Ori, as
discussed in the next section.

\subsection{Special treatment of particular sources}
\label{sec:spec-treatm-part}

For two of the additional bows listed in Table~\ref{tab:observations},
we are forced to deviate from the default values for the shell
parameters.  For LP~Ori, the bow shell appears to be formed from
neutral gas \citep{ODell:2001c} and its relatively high \(\tau\) value
is more than sufficient to trap the weak ionizing photon output of a
B3 star.  We therefore move its point in
Figure~\ref{fig:All-sources-eta-tau} downward to reflect a temperature
of \SI{1000}{K} (or, equivalently, a magnetically supported shell with
\(v\alfven \approx \SI{3}{km.s^{-1}}\)).  For the case of the Orion
Trapezium star \thD{}, we find that using the default parameters
results in a placement well inside the forbidden zone of
Figure~\ref{fig:All-sources-eta-tau} (indicated by fainter symbol).
For this object there is no reason to suspect anything but the usual
photoionized temperature of \SI{e4}{K}, but its placement could be
resolved either by decreasing the shell thickness, or decreasing the
UV dust opacity, or both. Given the moderate limb brightening seen in
the highest resolution images of the Ney--Allen nebula
\citep{Robberto:2005a, Smith:2005a}, the shell thickness is unlikely
to be less than half our default value.  But, if this were combined
with a factor 5 decrease in \(\kappa\), as suggested by
\citet{Salgado:2016a}, then this would be sufficient to move the
source up to the RBW line, or slightly above.

\subsection{Candidate radiation-supported bows}
\label{sec:cand-radi-supp}

Four sources are sufficiently close to the diagonal line
\(\eta\shell = 1.25 \tau\) in Figure~\ref{fig:All-sources-eta-tau}
that they should be treated as strong candidates for
radiation-supported bows. These are K18 sources~380 (HD~53367,
V750~Mon) and 407 (HD~93249 in Carina) plus \thD{} and LP~Ori.  Of the
four, source 380 is the only one that is also a candidate for
grain-gas decoupling.  Further details of the two K18 sources are
presented in Appendix~\ref{sec:notes-part-sourc}, where for source~380
we show that reducing both luminosity and distance by a factor of
roughly 2 with respect to the values used by K18 would provide a
better fit to the totality of observational data.  However, the ratio
\(\eta\shell /\tau\) is proportional to \(D / L_*\) so it would not be
affected by such an adjustment and the bow remains
radiation-supported.  On the other hand \(\eta\shell\) (proportional
to \(D / L_*^2\)) would increase by 2, making the classification as
dust wave candidate more marginal.

Three additional K18 sources (409, 410, 411) are within a factor of 3
of the radiation-supported line, so they too must be considered as
potential candidates, given our estimated systematic uncertainty in
the shell parameters (\S~\ref{sec:syst-uncert-due}).

\input{app-cloudy-dust-heating}

\section{Stellar wind mass-loss rates}
\label{sec:stellar-wind-mass}

Various methods have been proposed to derive stellar wind mass-loss
rates from observations of stellar bow shocks \citetext{for example,
  \citealp{Kobulnicky:2010a, Gvaramadze:2012a, Kobulnicky:2018a}}.  In
this section, we will show how the \(\tau\)--\(\eta\shell\) diagram can be
used to derive the mass-loss rate for bows in the wind-supported
regime.  We then compare our method with the method used by
\citet{Kobulnicky:2018a}, which is a refinement of that originally
proposed in \citet{Kobulnicky:2010a}.  The method used by
\citet{Gvaramadze:2012a} uses combined measurements of the bow shock
and the surrounding \hii{} region.  It has the advantage of depending
on fewer free parameters than the other methods, but can only be used
in the case of isolated stars, whereas the majority of the sources
considered here are in cluster environments.

\subsection{Mass loss determination from the \boldmath \(\tau\)--\(\eta\shell\) diagram}
\label{sec:mass-loss-determ}
\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/Mdot-vs-lum-combo-edited}
  \caption{Wind mass-loss rates as a function of stellar luminosity,
    derived from (a)~our trapped energy/momentum method and (b)~the
    grain emissivity method of \citet{Kobulnicky:2018a}, with
    corrections as described in our \S~\ref{app:bow-shock-data}.
    Circle symbols show the sources from K18, colored according to the
    stellar effective temperature (see key at far right). In panel a,
    squares show two of our additional sources
    (Tab.~\ref{tab:observations}). Upper limits to the mass loss are
    given for sources that lie close to the radiation-supported line
    in Fig.~\ref{fig:All-sources-eta-tau}, represented by faint
    symbols and downward-pointing arrows.  Sources that lie less than
    a factor of three above the line have enhanced downward
    uncertainties and are also shown by slightly fainter symbols.  For
    \(\sigma\)~Ori, the large symbol corresponds to the sum of the
    luminosities of the triple OB system Aa, Ab, and B
    \citep{Simon-Diaz:2015a}, while the small symbol corresponds to
    the luminosity of only the most massive component Aa.  Lines show
    the predictions of stellar wind models: red lines are the commonly
    used recipes from \citet{Vink:2000a} for dwarfs (solid), giants
    (dashed), and supergiants (dotted), while black lines show
    eq.~(11) of \citet{Krticka:2017a} for O~stars (solid) and models
    of \citet{Krticka:2014a} for B stars.  Orange plus symbols show
    mass-loss measurements from NUV lines for weak-wind O~dwarfs
    \citep{Marcolino:2009a}, while the green plus symbol shows the
    measurement from infrared H recombination lines for \(\sigma\)~Ori
    \citep{Najarro:2011a}.  Boxes show the sensitivity of the results
    to observational uncertainties (lower right) and assumed shell
    parameters (upper left).}
  \label{fig:mass-loss-vs-luminosity}
\end{figure*}

Taking into account the support from stellar wind ram pressure and the
absorbed stellar radiation pressure, the pressure of the bow shell can
be written
\begin{equation}
  \label{eq:eta-shell-components}
  \eta\shell = \eta\wind +  \left(  1 - e^{-\Qp \tau / Q_{\text{abs}}}\right) \ ,
\end{equation}
which is valid in all 3 regimes: WBS, RBW, and RBS.\@ The wind
momentum efficiency is defined in terms of the stellar wind parameters
(eq.~[13] of Paper~I):
\begin{equation}
  \label{eq:wind-eta-typical}
  \eta\wind = \num{0.495} \,\dot{M}_{-7} \,V_3  \,L_4^{-1} \ .
\end{equation}
Therefore, assuming \(\Qp/Q_{\text{abs}} = 1.25\) as in
\S~\ref{sec:energy-trapp-vers}, the mass-loss rate can be estimated as
\begin{equation}
  \label{eq:mdot-tau-eta}
  \dot{M}_{-7} = 2.02 \, L_4 \, V_3^{-1} \,
  \left[ \eta\shell - \left(  1 - e^{-1.25 \tau} \right) \right]  \ .
\end{equation}
In the wind-supported WBS regime
(\(\tau \ll \eta\wind \approx \eta\shell\)), the second term in the brackets is
negligible, and we can use equations~(\ref{eq:tau-empirical},
\ref{eq:eta-shell}) to write
\begin{equation}
  \label{eq:mdot-eta-wbs}
  \dot{M}_{-7} \approx 990 \frac{R\pc \, T_4 \, L_{\text{IR},4}}
  {L_4 \, V_3 \, \kappa_{600} \, h_{1/4}} \ .
\end{equation}
On the other hand, if \(\eta\shell\) does not greatly exceed \(\tau\), then
the full equation~\eqref{eq:mdot-tau-eta} should be used, although the
uncertainties in \(\eta\shell\) and \(\tau\), combined with the partial
cancellation of two similar-sized terms, mean that the mass loss will
be poorly constrained in such cases.

Resulting mass-loss rates are shown as a function of stellar
luminosity in Figure~\ref{fig:mass-loss-vs-luminosity}a for the K18
and Orion sources.  As in Figure~\ref{fig:All-sources-eta-tau} we show
separately the uncertainties in the measurements due to observational
errors (lower right box) and systematic model uncertainties (upper
left box).  Typical observational uncertainties are combined into an
error ellipse using the techniques of
Appendix~\ref{sec:comb-uncert-covar}, which is shown in light blue.
Sources that are close to the RBW line in the \(\tau\)--\(\eta\shell\)
diagram yield only upper limits to the mass-loss rate and these are
shown as faint symbols in the figure.

Our results are compared with the commonly used mass-loss recipes of
\citet{Vink:2000a} (red lines) and more recent whole-atmosphere
simulations of wind formation in B~stars \citep{Krticka:2014a} and
O~stars \citep{Krticka:2017a}.  Additionally, we show observational
mass loss determinations for weak wind O~dwarfs
\citep{Marcolino:2009a} as orange plus symbols.  It can be seen that
the majority of the K18 sources fall below the \citeauthor{Vink:2000a}
line and are more consistent with the \citet{Krticka:2017a} models.

\textit{Write about sigma Ori}

\subsection{Mass loss determination method of \protect\citet{Kobulnicky:2018a}}
\label{app:bow-shock-data}

% \begin{figure*}
%   \centering
%   \includegraphics[width=\linewidth]{figs/K18-pairplot-edited}
%   \caption[K18 pair plot]{Pair plots of correlations between observed
%     and derived parameters of bows from the \citep{Kobulnicky:2018a}
%     sample.}
%   \label{fig:K18-pairplot-edited}
% \end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/K18-emissivity-vs-U}
  \caption{Discrepancies in \SI{70}{\um} emissivities that we have
    identified in K18.  Red crosses show the emissivities given in
    K18's Table~2 as a function of the radiation field \(U\), while
    the solid gray line shows the emissivities that they claim to be
    using from \citet{Draine:2007a}.  The dashed line shows the
    emissivities that we believe they should have been using, which
    correct for the marked difference in spectral energy distribution
    between OB stars and the Galactic interstellar radiation field
    (see App.~\ref{sec:grain-temp-emiss}).}
  \label{fig:k18-emissivity}
\end{figure}

\newcommand\LOS{\ensuremath{_{\text{los}}}}

K18 derive mass loss rates for their sources using a method that is
different from the one that we employ above.  Both methods are based
on determining the stellar wind ram pressure that supports the bow
shell, but K18 do so via the following steps:
\begin{enumerate}[K1.]
\item \label{K1} The line-of-sight mass column through the shell is
  calculated by combining the peak surface brightness at \SI{70}{\um},
  \(S_{70}\), with a theoretical emissivity per nucleon,
  \(j_{70}(U)\), from \citet{Draine:2007a}:
  \(\Sigma\LOS = S_{70} / \bar{m} j_{70}(U)\).  This depends on
  knowledge of the stellar radiation field at the shell:
  \(U \propto L_* / R_0^2\).
\item \label{K2} The shell density is found from the line-of-sight
  mass column using an observationally determined ``chord diameter'',
  \(\ell\), which is assumed to be equal to the depth along the line
  of sight: \(\rho\shell = \Sigma\LOS / \ell\).
\item \label{K3} The internal ram pressure is equated to the external
  ram pressure, which is found by assuming a stream velocity of
  \SI{30}{km.s^{-1}} and a compression factor of 4 across the outer
  shock:
  \(P_{\text{stream}} = 0.25 \rho\shell \times
  (\SI{30}{km.s^{-1}})^2\).
\end{enumerate}
There are clear parallels but also differences between steps
K\ref{K1}--K\ref{K3} and our own steps P\ref{P1}--P\ref{P3}.  Our
step~P\ref{P1} depends on the total observed infrared flux of the bow
combined with an assumption about the grain opacity at ultraviolet
wavelengths, while step~K\ref{K1} depends on the peak brightness at a
single wavelength combined with an assumption about the grain
emissivity at infrared wavelengths.  Our step~P\ref{P2} requires an
assumption about the relative thickness of the shell, while
step~K\ref{K2} is more directly tied to observations.
On the other hand, step~K\ref{K3} makes a roughly equivalent
assumption about the shock compression factor,\footnote{%
  In reality, the compression factor may be larger or smaller than 4,
  depending on the efficiency of the post-shock cooling (see
  % XREF Paper I \ref{sec:radi-cool-lengths}
  \S~3.2 of Paper~I).  For instance, for \(v = \SI{30}{km.s^{-1}}\) as
  assumed by K18 and \(T = \SI{e4}{K}\), one has a Mach number of
  \(\M_0 = 2.63\) and a compression factor of 2.8 for a non-radiative
  shock (by
  % XREF Paper I \ref{eq:shock-n-jump}
  eq.~[27] of Paper~I) or a factor of
  \(\M_0^2 =6.9\) for a strongly radiative one.} %
and a further assumption about the stream velocity. These assumptions
are not necessary for our step~P\ref{P3}, but we do need to assume a
value for the shell gas temperature.

In principle, both methods are valid and their different assumptions
and dependencies on observed quantities and auxiliary parameters
provide an important cross check on one another.  However, as
explained in detail in \S~\ref{sec:grain-temp-emiss}, the
\(j_\nu(U)\) relation depends on the shape of the illuminating SED,
which means that the \citet{Draine:2007a} models require modification
when applied to grains around OB stars.  In addition, when attempting
to replicate the K18 values of \(j_{70}\) we find that they only
follow the \citet{Draine:2007a} values for \(U > \num{e4}\), tending
to a constant value for weaker radiation fields.  The situation is
summarized in Figure~\ref{fig:k18-emissivity}, where the values taken
directly from Table~2 of K18 are shown by red crosses, the values they
claim to be using are shown by the gray solid line (this curve is
consistent with that shown in K18 Fig.~2), and the values they
\textit{should} have been using are shown by the gray dashed line.

% If we replaced K3 with M3, then mass loss rates would be reduced by 1.86 across the board

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/K18-mdot-Ux8-comparison}
  \caption{Effects on mass-loss determination of correcting the K18
    emissivities.  The mass-loss rates from Table~2 of K18 are shown
    on the \(x\) axis, while the corrected values are shown on the
    \(y\) axis.  Symbols are color coded by the strength of the
    radiation field, \(U\). The corrected mass-loss rates are
    predominantly lower by a factor of roughly 2.}
  \label{fig:k18-mdot-corrected-emissivity}
\end{figure}

After correcting the \SI{70}{\um} emissivities in this way, we
re-derive the mass loss rates, following the same steps as in K18,
which are then used in Figure~\ref{fig:mass-loss-vs-luminosity}b
above.  The difference between these corrected mass-loss rates and
those published in K18 is shown in
Figure~\ref{fig:k18-mdot-corrected-emissivity}.  It can be seen that
sources with \(U \approx \num{e3}\) (darker shading) are relatively
unaffected but that sources with stronger radiation fields (lighter
shading) have their mass-loss increasingly reduced, as could be
surmised from Figure~\ref{fig:k18-emissivity}.  The average reduction
is by a factor of about two.

% Different scenarios for producing velocities: dynamic ejection from
% young clusters \citep{Hoogerwerf:2001a, Oh:2016c} produce high
% velocities, dissolution of binary systems following core-collapse SN
% \citep{Renzo:2018a} tend to produce lower velocities for the unbound
% MS companion (walkaways, slower than 30 km/s).  Also, champagne flows
% have low velocities.

% \subsection{Diagnosing bow type from shell emission profile}
% \label{sec:diagnosing-bow-type}

% Given that the predicted shell density profile varies between
% wind-supported and radiation-supported bows (Paper~I) it is worth
% considering whether it may be possible to use observed emission
% profiles to distinguish between the two.

\section{Discussion}
\label{sec:discussion}

\subsection{Why the two mass-loss methods differ}
\label{sec:underst-diff-betw}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/K18-mdot-corrected-comparison-R0-edited}
  \caption{Comparison of the two mass-loss methods: K18 corrected
    method (\(x\) axis) versus our method (\(y\) axis).  Error bars on
    the \(y\) axis correspond to a factor-three uncertainty in
    \(\eta\shell\).  Sources for which these error bars overlap with
    the RBW zone are only upper limits for the wind mass-loss rate,
    and are indicated by faint symbols.}
  \label{fig:mass-loss-comparison}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/mdot_correlations_edited}
  \caption{Correlations of derived mass loss rates with star and bow
    parameters. First row is for mass-loss rate derived using the
    method of this paper (\S~\ref{sec:mass-loss-determ}).  Second row
    is for mass-loss rate derived using the corrected K18 method
    (\S~\ref{app:bow-shock-data}).  Third row is the ratio of these
    two methods.  Points show 18 of the 20 K18 sources, omitting the
    two strongest candidates for radiation support.  The correlation
    coefficient, \(r\), and linear regression slope, \(m\), are shown
    for selected pairs of interest (fainter text indicates weaker
    correlations).  These were calculated using the Python library
    function \texttt{scipy.stats.linregress}.}
  \label{fig:correlations}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/H_versus_ell_annotated}
  \caption{Relative chord length versus shell thickness.  The line and
    gray shading shows the theoretical relation (eq.~\eqref{eq:h-R0})
    for planitude of \(\Pi = 1.8 \pm 0.9\).  The short vertical lines
    at the top of the graph show the chord lengths for each of the 20
    K18 sources.  The sources with the four smallest and four largest
    values of \(\ell/R_0\) are individually labelled. The mean value
    of \(\ell / R_0 = 1.8\) is indicated by a dashed line, which
    corresponds to \(H/R_0 = 0.25\), which is the common value that we
    use in deriving \(\eta\) and therefore \(\dot{M}\) by our method.
    Inset diagram at lower right shows the geometry that leads to
    eq.~\eqref{eq:h-R0}.  }
  \label{fig:H-versus-ell}
\end{figure}

From comparing the two panels in
Figure~\ref{fig:mass-loss-vs-luminosity} it is clear that there is a
broad average agreement between the \(\dot{M}\) values from the two
methods.  Nevertheless, there is considerable disparity for individual
sources.  Obviously this is to be expected for sources where the
\(\tau\)--\(\eta\shell\) method implies radiation support, since the
K18 method assumes stellar wind support in all cases, and so has no
way of identifying these.  However, discrepancies remain even for
sources that are well above the diagonal line in
Figure~\ref{fig:All-sources-eta-tau}.  This is shown more clearly in
Figure~\ref{fig:mass-loss-comparison}, where we compare the two
mass-loss determinations. Plot symbols are colored according to the
physical size of the bow, \(R_0\), and radiation-support candidates
are shown fainter.  It is apparent that there is only weak correlation
between the two techniques (Pearson correlation coefficient
\(r = 0.67\)).  Interestingly, the smaller bows (red/orange shading)
show much better agreement than the larger bows (yellow shading).  The
five bows with \(R_0 > \SI{0.4}{pc}\) (339, 344, 369, 381, 382) show a
difference of nearly an order of magnitude, in the sense that our
method consistently predicts lower mass-loss rates than K18.  The
error ellipse for the combined observational errors (see
Tab.~\ref{tab:error-ellipse} in App.~\ref{sec:comb-uncert-covar}) is
highly elongated along the leading diagonal in
Figure~\ref{fig:mass-loss-comparison} due to the fact that most of the
observational uncertainties effect both mass-loss methods in a similar
way.  It is therefore unlikely that observational errors are a
significant contribution to the \textit{difference} in the two
\(\dot M\) methods, which must instead be due to the systematic model
uncertainties.  These are represented in the figure by the vertical
error bars, which show the effect of a factor of three uncertainty in
\(\eta\shell\) due to variations in shell thickness, dust opacity, and
shell gas temperature.  It can be seen that systematic errors of this
magnitude could indeed account for the observed differences, but the
questions remain: which parameter is causing the problem? and can we
correct for it?

Although we have seen (Fig.~\ref{fig:mass-loss-comparison}) that the
mass-loss discrepancy increases with \(R_0\), it is hard to understand
why the physical size per~se might cause this.  We have therefore
looked at the correlations between all of the observed and derived
quantities, showing some of the more interesting results in
Figure~\ref{fig:correlations}.  Each row of plots shows the dependence
(\(y\) axis) of the two different derived mass-loss rates, and their
ratio, on different parameters (\(x\) axis) of the star (leftmost four
columns) and bow shell (remaining six columns).  The most significant
correlations are marked in red with the value of the Pearson
correlation coefficient, \(r\) (calculated after taking the log of all
quantities).  The values shown with dark text have \(r^2 > 0.5\),
which means that at least 50\% of the variance in the \(y\) quantity
is ``explained'' by the variance in the \(x\) quantity (although this
cannot necessarily be taken to imply causation in either
direction). For some correlations, we also mark the linear regression
slope, \(m\).  Since we are working in logarithmic space, this is
equal to the power index in the relation \(y \propto x^m\).  Some
selected weaker correlations are also shown (fainter text).

For our \(\eta\)--\(\tau\shell\) mass-loss method (top row) the only
significant correlation is a positive one with the shell infrared
luminosity \(L\IR\) (rightmost column), with slope
\(m = 0.6 \pm 0.1\).  For the K18 mass-loss method (middle row), the
situation is very different, showing significant correlation with a
trio of quantities: a positive correlation with bow radius, \(R_0\),
and negative correlations with relative chord length, \(\ell/R_0\),
and radiation field at the shell, \(U\).  The same three correlations
(with inverted sense) are seen for the ratio of the two methods
(bottom row).  This is due to the fact that the
\(\eta\)--\(\tau\shell\) method shows much shallower slopes in its
(weak) correlations with these quantities.  As an example, we show the
values for \(\ell/R_0\) on the figure (seventh column):
\(m = -2.3 \pm 0.4\) for the K18 method but \(m = -0.7 \pm 0.4\) for
our method, which leaves a significant slope in the ratio of
\(m = 1.6 \pm 0.4\).  For the shell luminosity, on the other hand, the
ratio does not show any significant correlation at all.  This is
because both methods have essentially the same slope in their
correlation with \(L\IR\): \(m = 0.6 \pm 0.1\) for our method, and a
weak correlation with \(m = 0.5 \pm 0.3\) for the K18 method.

% From https://www.socscistatistics.com/pvalues/pearsondistribution.aspx
% With 18 points: r^2 > 0.22 for p < 0.05 (|r| > 0.469)
% With 18 points: r^2 > 0.35 for p < 0.01 (|r| > 0.590)
% With 18 points: r^2 > 0.50 for p < 0.001 (|r| > 0.71)

Out of the three quantities, \(R_0\), \(\ell/R_0\), and \(U\) (which
are highly correlated between themselves) we will concentrate on the
relative chord length \(\ell/R_0\) as the most likely culprit for the
discrepancy between the two mass-loss estimates.  This is because it
contains observational information that is used in the K18 method
(step K\ref{K2} of \S~\ref{app:bow-shock-data}), but is \textit{not
  used} in the \(\tau\)--\(\eta\shell\) method.  The closest
equivalent of \(\ell\) in our method is the shell thickness,
\(h\shell\), which enters in step~P\ref{P2} of
\S~\ref{sec:energy-trapp-vers}, but crucially we use a fixed fraction
of the bow radius, \(h\shell / R_0 = 0.25\), rather than basing it on
any observations.  In principle, we could derive \(h\shell\) from
\(\ell\) if we knew the radius of curvature of the shell,
\(R_{\text{c}}\), or equivalently the planitude:
\(\Pi = R_{\text{c}} / R_0\)\citep{Tarango-Yong:2018a}.  This is
illustrated in the inset to Figure~\ref{fig:H-versus-ell}, from which
we find
\begin{equation}
  \label{eq:h-R0}
  \frac{h\shell}{R_0} = \frac{1}{8\Pi} \left(\frac{\ell}{R_0}\right)^2 \ .
\end{equation}
This is plotted in Figure~\ref{fig:H-versus-ell} for planitudes of
\(\Pi = 1.8 \pm 0.9\), which are typical of OB star bows (this will be
shown in detail in the upcoming Paper~IV).  Also shown in the figure
are the individual relative chord lengths for the K18 sources (rug
plot at top).  Although the mean value of \(\ell/R_0 \approx 1.8\)
corresponds closely to the \(h\shell/R_0 = 0.25\) that we have used,

\subsection{Shared properties of candidates for radiation support}
\label{sec:comm-prop-cand}

For five of the K18 sources (380, 407, 409, 410, 411), our method of
\S~\ref{sec:mass-loss-determ} gives only an upper limit to the wind
mass-loss rate if one assumes a factor-three uncertainty in
\(\eta\shell\), raising the possibility that they may be
radiation-supported instead of wind-supported (see Paper~I).  It is
notable that these sources are among the smallest in the K18 sample,
all with \(R_0 < \SI{0.1}{pc}\).  The two Orion Nebula bows, LP~Ori
and \thD{} also fall into this category, and they are smaller still,
with \(R_0 < \SI{0.01}{pc}\).  Six of these 7 stars also occupy a
narrow range of spectral type\footnote{%
  However, this is not particularly statistically significant.  For
  instance, out of the twenty K18 sources, 7 have spectral type O8 or
  earlier.  Therefore, taking as the null hypothesis that the 5
  sources were randomly selected from the 20, then the chance that
  they should all be O9 or later is
  \(p = (1 - \frac{7}{20})^5 = 0.116\), which fails to meet the
  conventional \(2\sigma\) significance level of \(p = 0.05\).} %
between O9 and B0, corresponding to
\(T_{\text{eff}} = \text{\SIrange{28}{31}{kK}}\) (LP~Ori is a much
cooler B2 star with \(T_{\text{eff}} = \SI{20}{kK}\),
\citealp{Petit:2008a, Alecian:2013a}).

The small bow sizes are fully consistent with the expectations from
theory developed in Paper~I.  Radiation-supported bows are predicted
to occur either in high-density environments
(\(n > \SI{100}{cm^{-3}}\)) or for stars with particularly weak winds.
In either case, the bow is expected to be smaller than \SI{0.1}{pc}
for \(v_\infty \ge \SI{20}{km.s^{-1}}\), see Figures~2a,b and~6 of
Paper~I.\@

Another common property of these sources is that they all have high
ratios of chord length to stand-off radius, \(\ell/R_0\).  For
instance, from Figure~\ref{fig:H-versus-ell}, the four sources with
the highest \(\ell/R_0\) values are all among the five
radiation-supported candidates.\footnote{%
  In this case, the statistical significance is far higher, since
  \(p = (\frac{5}{20})^4 = 0.0039\) for the null hypothesis.}  From
equation~\eqref{eq:h-R0} this means they also should have thicker than
average shells.  This is consistent with our finding from Paper~I's
\S~3.3 that the shell tends to be broad in the radiation-supported bow
wave regime.  Furthermore, if the larger than average \(H/R_0\) were
to be taken account in the calculation of \(\eta\shell\)
(eq.~\eqref{eq:eta-shell}), then it would strengthen the case for
radiation support in these objects.

\subsection{The anomalous B~star KGK2010~2 in Cygnus: strong wind or
  trapped ionization front?}
\label{sec:anomalous-b-star}

Using emissivity at different wavelength, such as \SI{24}{\um}.  No
good reason to fear stochastic heating.  They mention not getting 70/24 ratio right.  Probably be better with our correction.

Correlations between observed parameters and with mass loss rates.
Pair plot from my jupyter notebook.

How both the UV opacity and the IR emissivity depend on the same VSG grains.

Effects of grain evolution.  RATD destroying larger grains.  How would
this effect the abundance of smaller grains? Calculate maximum grain
size for the \(U\) distribution of the K18 sample.  Do we see any
effects that seem to vary with \(U\)?

Evidence for small grains around Arches cluster near Galactic center \citep{Hankins:2017a}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "bs-bw-dw-03"
%%% End: