### 2019-03-30 papers/Papers1234/sec-obs-intro.tex
Finish intro for Paper III

\section{Introduction}
\label{sec:introduction}

In the Orion Nebula, at least three different classes of stellar bow
shocks have been identified. As well as small number of OB bow shocks \citep{Smith:2005a, ODell:2001c},

HMXRB in external galaxies, possible bowshock in LMC~X-1 \citep{Hyde:2017a}.

Non-thermal radio emission from BD+43~3654 \citep{Benaglia:2010a}, has
been searched for but

Cooler red supergiant and asymptotic giant branch stars
\citep{Ueta:2008a, Sahai:2010a, Cox:2012a}.

Instability of bow \citep{Blondin:1988a}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "obs-bowshocks"
%%% End: