### 2019-05-01 papers/Papers1234/figs/mipsgal-summary-stats-table-body.tex

%%% MIPSGAL summary statistics
%%% Table automatically generated from mipsgal-summary-stats.tab
%%% 2019-04-30 22:48:49.149121
%%%
% \Width is used to align number under col header in first column
	\newlength\Width\settowidth\Width{Comparison}
	\begin{tabular}{
	  @{} ll @{\quad }
	  % Mean
	  S[round-mode=places]S[round-mode=places]
	  % Std Dev
	  S[round-mode=places]S[round-mode=places]
	  % Obs Disp
	  SS
	  % Effect sizes
	  @{\quad} SSS[round-mode=places]
	  % p values
	  @{\quad}
	  S[table-format = +1.2e1]
	  S[table-format = +1.2e1]
	  S[table-format = +1.2e1] @{}
	  }\toprule
	  & {Dependent}
	  & \multicolumn{2}{c}{Mean}
	  & \multicolumn{2}{c}{Std.\ Dev.}
	  & \multicolumn{2}{c}{Obs.\ Disp.}
	  & \multicolumn{3}{c @{} }{\dotfill Effect sizes\dotfill }
	  & \multicolumn{3}{c @{}}{Non-parametric test \(p\)-values} \\ 
	  {Comparison} & {Variable}
	  & {\(\langle \text{A} \rangle\)} & {\(\langle \text{B} \rangle\)}
	  & {\(\sigma_{\text{A}}\)} & {\(\sigma_{\text{B}}\)}
	  & {\(\langle \epsilon_{\text{A}} \rangle\)} & {\(\langle \epsilon_{\text{B}} \rangle\)}
	  & {\(r_b\)} & {Cohen \(d\)} & {\(\sigma_{\text{A}}/\sigma_{\text{B}}\)}
	  & {K--S} & {Rank} &  {B--F}\\
	  {\makebox[\Width]{(1)}} & \multicolumn{1}{c@{\quad}}{(2)}
	  & {(3)} & {(4)}
	  & {(5)} & {(6)}
	  & {(7)} & {(8)}
	  & {(9)} & {(10)} & {(11)}
	  & {(12)} & {(13)} & {(14)}  \\  
	  \midrule\multicolumn{10}{@{} l @{}}{\itshape Median split of continuous independent variables}\\
\addlinespace
Faint/bright & \(\Pi\) & 1.655 & 1.917 & 0.631 & 1.045 & 0.097 & 0.078 & 0.123 & 0.303 & \itshape 1.654 & 0.254 & 0.111 & \itshape 0.0335\\
\(H\) magnitude & \(\Lambda\) & 1.677 & 1.768 & 0.269 & 0.316 & 0.231 & 0.236 & \itshape 0.174 & \itshape 0.308 & 1.175 & \itshape 0.0156 & \itshape 0.0235 & 0.125\\
\(n_{\text{A}} =  n_{\text{B}} = 113\) & \(\Delta \Lambda\) & 0.183 & 0.198 & 0.161 & 0.163 &   &   & 0.07 & 0.093 & 1.013 & 0.191 & 0.365 & 0.761\\
\addlinespace
Low/high & \(\Pi\) & 1.766 & 1.803 & 0.975 & 0.755 & 0.114 & 0.062 & 0.1 & 0.043 & 0.774 & 0.358 & 0.192 & 0.599\\
bow shock size, \(R_0\) & \(\Lambda\) & 1.707 & 1.739 & 0.25 & 0.336 & 0.256 & 0.212 & 0.061 & 0.11 & \bfseries 1.342 & 0.0815 & 0.428 & \bfseries 0.00139\\
\(n_{\text{A}} =  n_{\text{B}} = 113\) & \(\Delta \Lambda\) & 0.175 & 0.204 & 0.157 & 0.165 &   &   & 0.091 & 0.176 & 1.054 & \itshape 0.042 & 0.238 & 0.193\\
\addlinespace
Low/high & \(\Pi\) & 1.725 & 1.846 & 0.822 & 0.917 & 0.091 & 0.085 & 0.082 & 0.139 & 1.116 & 0.254 & 0.285 & 0.982\\
extinction, \(A_K\) & \(\Lambda\) & 1.703 & 1.742 & 0.267 & 0.323 & 0.233 & 0.235 & 0.04 & 0.132 & 1.213 & 0.524 & 0.602 & 0.123\\
\(n_{\text{A}} =  n_{\text{B}} = 113\) & \(\Delta \Lambda\) & 0.186 & 0.195 & 0.138 & 0.183 &   &   & -0.039 & 0.057 & 1.326 & 0.421 & 0.61 & 0.112\\
\addlinespace
Low/high & \(\Pi\) & 1.706 & 1.862 & 0.727 & 0.988 & 0.085 & 0.091 & 0.069 & 0.181 & 1.358 & 0.0568 & 0.368 & 0.0842\\
\(\vert{}b\vert\) & \(\Lambda\) & 1.722 & 1.724 & 0.328 & 0.261 & 0.234 & 0.234 & 0.02 & 0.008 & 0.796 & 0.323 & 0.795 & 0.0534\\
\(n_{\text{A}} =  n_{\text{B}} = 113\) & \(\Delta \Lambda\) & 0.188 & 0.191 & 0.161 & 0.162 &   &   & 0.009 & 0.021 & 1.005 & 0.95 & 0.907 & 0.694\\
\addlinespace
High/low & \(\Pi\) & 1.807 & 1.751 & 0.946 & 0.742 & 0.09 & 0.084 & -0.0 & -0.064 & 0.785 & 0.998 & 0.999 & 0.549\\
\(\cos \ell\) & \(\Lambda\) & 1.734 & 1.707 & 0.279 & 0.321 & 0.241 & 0.223 & -0.049 & -0.093 & 1.152 & 0.554 & 0.532 & 0.159\\
\(n_{\text{A}}, n_{\text{B}} = 137, 90\) & \(\Delta \Lambda\) & 0.182 & 0.201 & 0.155 & 0.171 &   &   & 0.054 & 0.122 & 1.1 & 0.791 & 0.491 & 0.365\\
\midrule
\multicolumn{10}{@{} l @{}}{\itshape Categorical independent variables}\\
\addlinespace
Environment: & \(\Pi\) & 1.757 & 1.852 & 0.854 & 0.899 & 0.087 & 0.083 & 0.042 & 0.11 & 1.053 & 0.67 & 0.676 & 0.377\\
Isolated vs Facing & \(\Lambda\) & 1.735 & 1.693 & 0.283 & 0.338 & 0.238 & 0.218 & -0.07 & -0.142 & 1.194 & 0.877 & 0.49 & 0.392\\
\(n_{\text{A}}, n_{\text{B}} = 170, 41\) & \(\Delta \Lambda\) & 0.19 & 0.195 & 0.161 & 0.172 &   &   & -0.019 & 0.034 & 1.066 & 0.883 & 0.85 & 0.438\\
\addlinespace
Environment: & \(\Pi\) & 1.757 & 1.907 & 0.854 & 0.955 & 0.087 & 0.105 & 0.024 & 0.174 & 1.118 & 0.587 & 0.875 & 0.255\\
Isolated vs \hii & \(\Lambda\) & 1.735 & 1.68 & 0.283 & 0.309 & 0.238 & 0.233 & -0.13 & -0.193 & 1.092 & 0.67 & 0.391 & 0.782\\
\(n_{\text{A}}, n_{\text{B}} = 170, 16\) & \(\Delta \Lambda\) & 0.19 & 0.175 & 0.161 & 0.138 &   &   & -0.048 & -0.095 & 0.855 & 0.825 & 0.754 & 0.799\\
\addlinespace
Single/multiple & \(\Pi\) & 1.767 & 1.833 & 0.825 & 0.988 & 0.09 & 0.08 & 0.027 & 0.076 & 1.198 & 0.965 & 0.756 & 0.605\\
source candidate & \(\Lambda\) & 1.709 & 1.762 & 0.289 & 0.315 & 0.23 & 0.243 & 0.074 & 0.177 & 1.09 & 0.431 & 0.396 & 0.338\\
\(n_{\text{A}}, n_{\text{B}} = 167, 60\) & \(\Delta \Lambda\) & 0.184 & 0.206 & 0.162 & 0.16 &   &   & 0.093 & 0.136 & 0.987 & 0.293 & 0.284 & 0.97\\
\addlinespace
With/without & \(\Pi\) & 1.714 & 1.802 & 0.598 & 0.926 & 0.091 & 0.087 & -0.012 & 0.101 & 1.547 & 0.948 & 0.904 & 0.2\\
\SI{8}{\um} emission & \(\Lambda\) & 1.734 & 1.721 & 0.289 & 0.298 & 0.223 & 0.237 & -0.042 & -0.044 & 1.031 & 0.559 & 0.667 & 0.563\\
\(n_{\text{A}}, n_{\text{B}} = 45, 182\) & \(\Delta \Lambda\) & 0.203 & 0.186 & 0.205 & 0.149 &   &   & 0.021 & -0.106 & 0.726 & 0.985 & 0.826 & 0.219\\
\addlinespace
3-star vs (4+5)-star & \(\Pi\) & 1.632 & 2.0 & 0.91 & 0.764 & 0.106 & 0.061 & \bfseries 0.386 & \bfseries 0.431 & 0.84 & \bfseries 5e-06 & \bfseries 7.66e-07 & 0.762\\
 & \(\Lambda\) & 1.661 & 1.812 & 0.288 & 0.286 & 0.247 & 0.216 & \bfseries 0.328 & \bfseries 0.525 & 0.99 & \bfseries 8.45e-05 & \bfseries 2.63e-05 & 0.403\\
\(n_{\text{A}}, n_{\text{B}} = 133, 94\) & \(\Delta \Lambda\) & 0.189 & 0.19 & 0.162 & 0.161 &   &   & -0.007 & 0.009 & 0.994 & 0.676 & 0.927 & 0.561\\
\midrule
\multicolumn{10}{@{} l @{}}{\itshape Intercomparison with other datasets}\\
\addlinespace
MIPS vs Orion & \(\Pi\) & 1.784 & 3.09 & 0.871 & 1.666 &   &   & \bfseries 0.57 & \bfseries 1.37 & \bfseries 1.911 & \bfseries 8.55e-05 & \bfseries 5.73e-05 & \bfseries 0.00133\\
 & \(\Lambda\) & 1.723 & 2.532 & 0.297 & 0.749 &   &   & \bfseries 0.71 & \bfseries 2.308 & \bfseries 2.526 & \bfseries 6.62e-07 & \bfseries 5.49e-07 & \bfseries 1.49e-09\\
\(n_{\text{A}}, n_{\text{B}} = 227, 18\) & \(\Delta \Lambda\) & 0.19 & 0.658 & 0.162 & 0.513 &   &   & \bfseries 0.642 & \bfseries 2.243 & \bfseries 3.173 & \bfseries 4.98e-07 & \bfseries 5.95e-06 & \bfseries 1.07e-10\\
\addlinespace
MIPS vs RSG & \(\Pi\) & 1.784 & 1.479 & 0.871 & 0.227 &   &   & -0.159 & -0.362 & \itshape 0.26 & \itshape 0.0129 & 0.288 & \itshape 0.0151\\
 & \(\Lambda\) & 1.723 & 1.41 & 0.297 & 0.101 &   &   & \bfseries -0.719 & \bfseries -1.089 & \itshape 0.339 & \bfseries 2.89e-06 & \bfseries 1.56e-06 & \itshape 0.004\\
\(n_{\text{A}}, n_{\text{B}} = 227, 16\) & \(\Delta \Lambda\) & 0.19 & 0.147 & 0.162 & 0.076 &   &   & -0.072 & -0.267 & 0.469 & 0.518 & 0.701 & 0.12
\\
\bottomrule
\multicolumn{14}{@{}p{\linewidth}@{}}{
  \textit{Description of columns:}
  (Col.~1)~How the two A/B source sub-samples are defined, also giving the size of each sub-sample, \(n_{\text{A}}\) and~\(n_{\text{B}}\).
  (Col.~2)~Dependent variable whose distribution is compared between the two sub-samples.
  (Cols.~3--6)~Mean and standard deviation, \(\sigma\), of the dependent variable for each of the two sub-samples.
  (Cols.~7--8)~Mean over each sub-sample of the observational dispersion (\(\epsilon\), standard deviation) of radii that contribute to the dependent variable for each individual source, as in steps~\ref{step:R0} and \ref{step:R90} of \S~\ref{sec:autom-trac-fitt}.  Note that in the case of \(\Pi\), this is \(\epsilon(R_0)\), and so is not a direct measure of the observational uncertainty in \(\Pi\). 
  % (Cols.~9--10)~``Standard error of the mean'' (s.e.m.) of the dependent variable for each of the two sub-samples. 
  (Cols.~9--11)~Standardized ``effect sizes'', which are dimensionless measures of the difference in the distribution of the dependent variable between the two sub-samples.
  (Col.~9)~Rank biserial correlation coefficient \citep{Cureton:1956a}, which is obtained by considering all \(n_{\text{A}} n_{\text{B}}\) pair-wise comparisons of the dependent variable between a source in sub-sample~A and a source in sub-sample~B.  It is the difference between the fraction of such comparisons ``won'' by sub-sample~A and those ``won'' by sub-sample~B, and thus may vary between \(-1\) and \(+1\). 
  (Col.~10)~Cohen's \(d\), which is a dimensionless mean difference: \(d = (\langle \text{A} \rangle - \langle \text{B} \rangle) / \sigma_{\text{pool}} \), where \(\sigma_{\text{pool}} = (n_{\text{A}} \sigma_{\text{A}}^2 + n_{\text{B}} \sigma_{\text{B}}^2)^{1/2} / \sqrt{n_{\text{A}} + n_{\text{B}}}\) is the pooled standard deviation.
  (Col.~11)~Ratio of standard deviations between the two sub-samples.
  (Cols.~12--14)~Probabilities (\(p\)-values) of the two sub-samples being as different as observed if they were to be drawn from the same population, according to three different non-parametric tests.
  (Col.~12)~Anderson--Darling 2-sample test, which is a general test of similarity between two distributions that is designed to retain sensitivity to differences in the tails of the distributions.
  (Col.~13)~Mann--Whitney--Wilcoxon \(U\) test \citep{Mann:1947a}, which is sensitive to differences in the central value of the distributions.
  (Col.~14)~Brown--Forsythe test for equality of variance \citep{Brown:1974a}
}
\end{tabular}