### 2019-05-09 papers/Papers1234/sec-obs-conclusions.tex
Finish discussion for Paper IV - now just summary to go!

\section{Summary}
\label{sec:conclusion}

We have presented a statistical study of the shapes of three different
classes of stellar bow shocks, as characterized by their planitude
(flatness of apex) and alatude (openness of wings), following the
terminology of \citet[Paper~0]{Tarango-Yong:2018a}.  Our principal
findings are as follows:
\begin{enumerate}[1.]
\item 
\end{enumerate}

% Difference in SEDs between FIR and MIR samples?  \citet{Meyer:2016a}
% say emission peaks at 3--50 micron for O star bow shocks.

% Al other things being equal, larger bows will be at higher
% inclinations (we should estimate how much variation in size we should
% get due to inclination -- it will be more for flatter bows).  Higher
% inclinations means less variation from the standing wave oscillations,
% which goes against our result that larger bows have greater dispersion
% in \(\Lambda'\).  Although the standing waves mainly give dispersion in
% \(\Pi'\) anyhow.

% \citet{Meyer:2014a} say that in cool stars the dust emission comes
% from the shocked inner wind, rather than from near the contact
% discontinuity for hot stars.  This could explain the difference in
% shape between the Herschel and Spitzer samples.

% How different regions of the \(\Pi\)--\(\Lambda\) plane are populated.
% Bottom-right quadrant hard to get to (except for standing wave
% oscillations), but may be due to finite shell thickness, which (for
% low Mach number) will be more apparent in the wings, which might
% decrease \(\Lambda\) more than \(\Pi\).  Fact that thin-shell solutions should
% trace the contact discontinuity, but in some cases it may be only the
% inner or the outer shell that is visible.

% Justification for standing waves: Fig.~3 of \citet{Meyer:2016a} shows
% a time sequence of thin-shell instability, which looks a bit like a
% standing wave. But much larger amplitude than we are considering.

% Deviations from axisymmetry as an alternative to oscillations.

% \subsection{The case of inside-out bows}
% \label{sec:case-inside-out}

% So far, we have considered the case where the inner source dominates
% the radiation, while dust is present only in the outer stream, which
% applies to hot stars interacting with the ISM.  However, in the case
% of cool stars, the inner wind will also be dusty.  Examples are the
% red supergiant (RSG) phase of high-mass evolution, or the asymptotic
% giant branch (AGB) stage of low/intermediate-mass evolution.  In both
% these cases, it is still the inner source that provides the radiation
% field.  However, not all winds are radiatively driven and in those
% cases it is conceivable that it is the outer source that dominates the
% radiation field.  An example is the case of photoevaporating
% protoplanetary disks (proplyds) in the Orion Nebula and other \hii{}
% regions \citep{ODell:1994a}.  In the proplyds, the inner wind is a
% thermally driven photoevaporation flow \citep{Henney:1998b, Henney:1999a},
% while the outer stream is the stellar wind from an O~star
% \citep{Garcia-Arredondo:2001a}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "obs-bowshocks"
%%% End: