### 2019-05-10 papers/Papers1234/obs-bowshocks.tex

%\RequirePackage{amsmath}
\documentclass[useAMS, usenatbib, a4paper]{mnras}
\pdfsuppresswarningpagegroup=1

% Standard LaTeX packages
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{fixltx2e}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage{color}
\usepackage{enumerate}
\usepackage{pdflscape}
\usepackage{rotating}
\usepackage{hyperref}

\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}

% Fonts 
\usepackage{newtxtext}
% Note: newtxmath must come AFTER newtxtext
\usepackage[varvw,smallerops]{newtxmath}

\usepackage{chemgreek}
\activatechemgreekmapping{newtx}

% Define a fermata symbol for use inline with text
% 04 Jul 2017 WJH
% Based on info in Secs 18.4.1 and 24.1 of musixdoc.pdf
\let\ORIGna\na
\usepackage{musixtex}
\makeatletter
\newcommand\textfermata{%
  {\let\extractline\relax
    \setlines10\smallmusicsize \nobarnumbers \nostartrule
    \staffbotmarg0pt \setclefsymbol1\empty \global\clef@skip0pt
    \raisebox{0ex}[0ex][0ex]{
      \startextract\addspace{-\afterruleskip}%
      \notes\fermataup{-2}\en
      \zendextract}}}
\makeatother
\let\na\ORIGna

% % The \eye command is in the dingbat package, but we need to take care
% % of a naming conflict with the AMS package
% \usepackage{savesym}
% \savesymbol{checkmark}
% \usepackage{dingbat}

% The "eye" symbol is drawn with the \faEye command of the fontawesome
% package
\usepackage{fontawesome}
% The ``irregular'' symbol is drawn from the Icelandic Sorcery and Witchcraft
\usepackage{staves}

\hypersetup{colorlinks=True, linkcolor=blue!50!black, citecolor=black,
  urlcolor=blue!50!black}

\usepackage{etoolbox}
\robustify\bfseries
\robustify\itshape

%% The following hack solves a problem with
%% ERROR: \pdfendlink ended up in different nesting level than \pdfstartlink.
%% See https://tex.stackexchange.com/a/249743
\makeatletter
\patchcmd\@combinedblfloats{\box\@outputbox}{\unvbox\@outputbox}{}{%
  \errmessage{\noexpand\@combinedblfloats could not be patched}%
}%
\makeatother

%% Bold italic
\newcommand\hmmax{0}            % we don't need heavy fonts
\newcommand\bmmax{1}            % reduce use of math alphabets for bold
\usepackage{bm}

%% Bundled custom packages
\usepackage{aastex-compat}

\defcitealias{Tarango-Yong:2018a}{Paper~0}
\newcommand\PaperI{\citetalias{Tarango-Yong:2018a}}

% \title[Stellar bow shocks]{Weather vanes and runaways: Bow
%   shocks across the HR diagram}
\title[Bow shocks, bow waves, and dust waves. IV.]{Bow shocks, bow
  waves, and dust waves. IV. Shell shape statistics}

\newcommand\AddressCRyA{Instituto de Radioastronom\'{\i}a y Astrof\'{\i}sica,
  Universidad Nacional Aut\'onoma de M\'exico, Apartado Postal 3-72,
  58090 Morelia, Michoac\'an, M\'exico}
\author[Henney et al.]{
  William J. Henney,\thanks{Email: w.henney@irya.unam.mx}
  Jorge A. Tarango-Yong,
  Luis \'Angel Guti\'errez-Soto,\thanks{%
    Current address: Observatório do Valongo,
    Universidade Federal do Rio de Janeiro,
    Ladeira Pedro Antonio 43, 20080-090 Rio de Janeiro, Brazil}
  \& S. J. Arthur
  \\
  \AddressCRyA
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2019}

\input{bs-bw-dw-defs}

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Stellar bow shocks result from stars that move with respect to their
  environment. The interaction of the stellar wind and radiation with
  gas and dust in the interstellar medium produces curved arcs of
  emission at optical, infrared, and radio wavelengths.  We recently
  proposed a new two-dimensional classification scheme for the shape
  of such bow shocks, which we here apply to three very different
  observational datasets: (1)~mid-infrared arcs around hot OB stars;
  (2)~far-infrared arcs around luminous cool stars; and
  (3)~H\(\alpha\) emission-line arcs around proplyds and other young stars
  in the Orion Nebula.  For OB stars, the average shape is consistent
  with simple thin-shell models for the interaction of a spherical
  wind with a parallel stream, but the diversity of observed shapes is
  many times larger than such models predict.  We propose that this
  may be caused by time-dependent oscillations in the bow shocks, due
  to either instabilities or wind variability.  Cool star bow shocks
  have markedly more closed wings than hot star bow shocks, which may
  be due to the dust emission arising in the shocked stellar wind
  instead of the shocked interstellar medium.  The Orion Nebula arcs,
  on the other hand, have both significantly more open wings and
  significantly flatter apexes than the hot star bow shocks.  We test
  several possible explanations for this difference: divergent ambient
  stream, low Mach number, observational biases, and influence of
  collimated jets.  Further observations are required to decide
  between these.
\end{abstract}

\begin{keywords}
  circumstellar matter -- methods: statistical -- stars: winds, outflows
\end{keywords}

\input{sec-obs-intro}
\input{sec-quadrics-observations}
\input{sec-standing-wave}
\input{sec-obs-conclusions}

\section*{Acknowledgements}
We are grateful for financial support provided by Dirección General de
Asuntos del Personal Académico, Universidad Nacional Autónoma de
México, through grants Programa de Apoyo a Proyectos de Investigación
e Inovación Tecnológica IN111215 and IN107019.

\bibliographystyle{mnras}
%% All references should be put in the BibTeX file bowshocks-biblio.bib
\bibliography{bowshocks-biblio}
\appendix
\input{app-p-values}

% Don't change these lines

\bsp	% typesetting comment
\label{lastpage}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: