### 2019-06-27 papers/bs-bw-dw-03-submit/reply-to-referee.txt
re-submitted version of Paper III

			 ━━━━━━━━━━━━━━━━━━━━━━
			  REPLY TO THE REFEREE
			 ━━━━━━━━━━━━━━━━━━━━━━

We thank the referee for a most helpful report, which is full of useful
recommendations that we have tried to follow.  We comment on the
specific requests below.  We include a version of the manuscript with
differences highlighted, which was generated automatically using
latexdiff.  New material is in red and underlined.  Removed material is
in gray and struck through.

┌────
│ 
│ 1) My main comment is whether you can define terms where they are
│ first used.  This is especially acute in section 2 - what is R_0,
│ R_pc. The "Relative velocity" of what to what. What is the difference
│ between capital xi_{sh} and capital xi. I had assumed this was just a
│ typo, but then see Section 4.0
└────

This is a fair point.  We have tried to address these complaints by more
carefully defining the terms and relying less on the previous papers.
The "relative velocity" is between the star and the ambient stream.  The
symbol Xi refers to the general concept of the radiation parameter,
which in Paper II was applied in the ambient stream (outside the shell),
whereas Xi_sh refers to the value of Xi in the shell, which is generally
lower (since the gas pressure goes up in the shock).

┌────
│ Other examples - what is n_\infinity - is it just n.
└────

Yes, we have removed references to n_∞.  This was a hold-over from a
distinction that we made in Paper I that is not important for this
paper.

┌────
│ Use of v_\infinity - not for the wind terminal velocity (a bit
│ confusing as it is the standard symbol for wind terminal
│ velocity). Though I recognise the need for consistency through the
│ papers (as much as possible).
└────

Again, the difference was important in Paper I, but not for this paper.
So we agree that it is cleaner to remove the subscript ∞.

┌────
│ 2) As part of this, can you define tau and kappa a bit more
│ closely. These quantities are usually frequency dependent, but here
│ you are using an effective (or averaged) values. There is the footnote
│ in Section 2, but even that had to read closely as to determine
│ whether tau (Paper I) or tau (Paper II) was the larger. Where do the
│ assumed values of w and g come from for Q_p/Q_abs.
└────

Yes. On reflection, these do deserve a much fuller explanation.  Rather
than litter the text with footnotes, we have added a new appendix A in
order to address this. There we discuss the derivation of monochromatic
opacities from the dust model, including the differences between
scattering, absorption, and radiation pressure.  We describe the
frequency dependence of w, g and Q_p/Q_abs. Then we give the formal
equation for the frequency averaging process.

┌────
│ 
│ I know what was meant after re-reading it a few times, but the overall
│ effect was confusing (and involved reading parts of Paper I and Paper
│ II to understand what was going on).
└────

┌────
│ Minor points - the values of f_UV in Sect 4.0 - can you provide reference.
└────

The values were calculated from TLUSTY atmosphere models.  References
now given.

We made two further changes to the paper, which we hope are acceptable
to the referee.

First, we have slightly toned down the critique of the Kobulnicky+
(2018) paper in sec 5.2.  It turns out that the authors were partially
aware of their mistake about the emissivities and are addressing it in a
follow-up paper (in preparation).  We have therefore removed the
corresponding figure (Fig 7 in original MS).

Second, we changed the wording of conclusion 8 following conversations
with Thiem Hoang (lead author of the Radiative Torque Disruption
papers).