### 2019-06-28 papers/Papers1234/sec-quadrics-observations.tex
Paper 4 Jane comments up to end of section 4

\section{Mid-infrared arcs around early-type stars}
\label{sec:mid-infrared-arcs}
\label{sec:comp-with-observ}

\begin{figure*}
  \includegraphics[width=\linewidth]{figs/mipsgal-example-panels}
  \caption[]{Examples of typical fits to the bow shock shapes of
    MIPSGAL sources.  %
    (a)~Scatter plot of the bow shock sizes determined from our fits,
    as compared with those tabulated by \citet[K16]{Kobulnicky:2016a}
    for the MIPSGAL sources.  Different symbol sizes and colors
    correspond to different star ratings, which indicate our
    subjective judgement of the quality of the fit.  The horizontal
    dotted line shows the MIPS \SI{24}{\um} point spread function
    FWHM of \(5.5\arcsec\).  %
    (b)~Histograms of the difference in bow shock position angles
    between our fits and the K16 values for sources with 3- to 5-star
    ratings.  %
    (c--e)~Source K510, with a 3-star rating. %
    (f--h)~Source K506, with a 4-star rating. %
    (i--k)~Source K517, with a 5-star rating. %
    Panels~(c, f, i) show a \SI{24}{\um} gray-scale image of a
    \(160'' \times 160''\) field of view, centered on each source.
    Contours are ten linearly spaced levels between the median
    brightness of the entire image and the maximum brightness of the
    bow shock arc. Grids of galactic coordinates (light blue lines,
    parallel to the box sides) and equatorial coordinates (tilted
    magenta lines) are shown.  The stellar source and the bow shock
    axis, as determined by \citet{Kobulnicky:2016a} are indicated by
    an orange star and an orange line, respectively, where the line
    extends from \(-2 R_0\) to \(+2 R_0\).  The automatically traced
    arc shapes using the ``mean'' and ``peak'' methods (see text) are
    shown by blue and red dots, respectively.  The magenta circle
    shows the fit to the arc points within \(\pm 45^\circ\) of the nominal
    bowshock axis, with the magenta dot showing the center of
    curvature and the magenta line showing the fitted bow shock axis,
    which is the line passing through the source and the center of
    curvature. %
    Panels~(d, g, j) show plots of the radius measured from the source
    of the arc points, plotted as a function of angle \(\theta\) from the
    nominal bow shock axis, and with the same color coding as used on
    the image. Angular ranges of \(\theta = \pm 45\degr\) and
    \(\pm 90\degr\) are shown by gray shaded boxes.  The \(R_0\) value
    tabulated by \citet{Kobulnicky:2016a} is shown by a horizontal
    blue line. %
    Panels~(e, h, k) show equivalent plots for the surface brightness.
  }
  \label{fig:mipsgal-examples}
\end{figure*}

The most extensive observational sample of stellar bow shock nebulae
to date is a catalog of 709 arcs \citep{Kobulnicky:2016a} detected in
mid-infrared surveys of the Galactic Plane by the \textit{Spitzer
  Space Telescope} (\textit{SST}, \citealp{Werner:2004a}) and
\textit{Wide-field Infrared Survey Explorer} (\textit{WISE},
\citealp{Wright:2010a}).  These sources are believed to be powered by
the winds of early-type stars, which are either moving supersonically
through the interstellar medium (runaway stars,
\citealp{Gvaramadze:2008a}), or are interacting with a local bulk
flow, such as the champagne flow from a nearby \hii{} region (weather
vanes, \citealp{Povich:2008a}).

\subsection{Automatic tracing and fitting of bow shocks}
\label{sec:autom-trac-fitt}

In order to study the shapes of these bow shocks, we downloaded data
from the NASA/IPAC Infrared Science Archive archive\footnote{
  \url{http://irsa.ipac.caltech.edu/docs/program_interface/api_images.html}}
and extracted 4\arcmin{} square images in the \SI{24}{\um} bandpass of
the Multiband Imaging Photometer for \textit{Spitzer} (MIPS) centered
on each of the 471 \citet{Kobulnicky:2016a} sources that are covered
by the MIPSGAL \citep{Carey:2009a} survey, which includes most of the
sources with Galactic longitude within \(\pm 60\degr\) of the Galactic
center.

We have developed a method for automatically tracing the arcs and
determining their planitude and alatude, which is an extension of the
method described in Appendix~E of Paper~0.  The steps of the method
are as follows:
\begin{enumerate}[1.]
\item Calculate arrays of celestial coordinates, \(C\), for each pixel
  of the image. In our implementation of the method we use functions
  from the Python library \texttt{astropy.coordinates}
  \citep{Astropy-Collaboration:2018a}.
\item Using the central source coordinates, \(C_0\) and nominal
  bowshock radius, \(R_0\) from \citet{Kobulnicky:2016a}, construct a
  pixel mask that includes only those pixels with separations from the
  source that satisfy \(\frac12 R_0 \le |C - C_0| \le 3 R_0\).  This mask
  will be used for all subsequent operations, which serves to help
  avoid confusion from the star itself and other bright sources in the
  field of view.
\item Define a ``step-back'' point, \(C_1\), which is located at a
  separation \(2 R_0\) from the source, but in the opposite direction
  from the apex of the bow shock. That is, along a position angle
  180\degr{} from the nominal position angle, \(\text{PA}_0\), of the
  bow shock axis.  This point is at one end of the orange line shown
  superimposed on the bow shock images in
  Figure~\ref{fig:mipsgal-examples}.
\item Looping over a grid of 50 position angles, \(\text{PA}_k\),
  within \(\pm 60\degr\) of \(\text{PA}_0\), estimate the location of
  the arc along rays cast from the step-back point, using two
  different methods:
  \begin{enumerate}[(a)]
  \item The pixel with the peak brightness, with coordinates
    \(C_{k,\text{peak}}\) (red dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \item The mean brightness-weighted separation from \(C_1\), with
    coordinates \(C_{k,\text{mean}}\) (light blue dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \end{enumerate}
  For each \(\text{PA}_k\) in the grid, the calculation is performed
  over only those pixels that satisfy
  \(|\text{PA}(C, C_1) - \text{PA}_k| < \frac12 \delta\theta\), where
  \(\delta\theta = 120/50 = 2.4\degr\), which defines a thin radial wedge from
  \(C_1\).  The results are shown as red and blue dots superimposed on
  the images in Figure~\ref{fig:mipsgal-examples}. Each of the two
  methods, ``peak'' and ``mean'', works better in some objects and
  worse in others (according to the subjective judgment of
  ``correctly'' tracing the bow shock shape).  We therefore take the
  average by amalgamating all the \(C_{k,\text{peak}}\) and
  \(C_{k,\text{mean}}\) points into a single set, \(C_{k}\), for the
  following steps.  We have found that the use of the step-back point
  greatly improves the reliability of the method, as compared with the
  simpler option of tracing rays from the stellar source.
\item For each of the points \(C_{k}\), determine the radial
  separation from the central source, \(R_k = |C_k - C_0|\) and the
  angle from the bow shock axis about the central source
  \(\theta_k = \text{PA}(C_k, C_0) - \text{PA}_0\).  These are plotted in
  the upper left panels of Figure~\ref{fig:mipsgal-examples}.  Note
  that, even though the rays are cast from the step-back point \(C_1\)
  within \(\pm 60\degr\) of \(\text{PA}_0\), the angles \(\theta_k\) are
  measured from the source, \(C_0\), which is closer to the bow shock
  than \(C_1\) and therefore \(|\theta_k|\) can be much larger than
  \(60\degr\).
\item Make our own estimate of the axial size, \(R_0\), of the bow
  shock by calculating the mean of \(R_k\) over all points \(C_k\)
  with \(|\theta_k| \le 10\degr\).  Note that this is distinct from the
  nominal value of \(R_0\) given in the \citet{Kobulnicky:2016a}
  catalog, which was ``measured by eye''.  We denote by
  \(\epsilon(R_0)\) the standard deviation of the \(R_k\) that go into
  calculating \(R_0\).\label{step:R0}
\item Estimate the radius of curvature, \(R_c\), by fitting a circle
  to all those points within \(\pm 45\degr\) of the nominal axis
  (\(|\theta_k| < 45\degr\)), but after excluding any point with
  \(R_k < \frac12 R_m\) or \(R_k > 2 R_m\), where \(R_m\) is the median
  \(R_k\) for \(|\theta_k| < 45\degr\).\label{step:Rc}
\item Determine two separate estimates, \(R_{90+}\) and \(R_{90-}\),
  of the perpendicular radius, \(R_{90}\), by taking the mean of
  \(R_k\) over all points \(C_k\) with
  \(|\theta_k - 90\degr| \le 10\degr\) for \(R_{90+}\), and with
  \(|\theta_k + 90\degr| \le 10\degr\) for \(R_{90-}\).  The average of the
  two standard deviations of the \(R_k\) that contribute to
  \(R_{90+}\) and \(R_{90-}\) is denoted by \(\epsilon(R_{90})\).\label{step:R90}
\item The planitude is determined as \(\Pi = R_c / R_0\) and the mean
  alatude as \(\Lambda = \frac12 (R_{90+} + R_{90-}) / R_0\).  We also
  calculate an alatude asymmetry as
  \(\Delta\Lambda = \frac12 (R_{90+} - R_{90-}) / R_0\). Note that in this paper
  we simplify the notation of Paper~0 by no longer using the prime
  symbol (\('\)) to distinguish projected from intrinsic quantities.
  All planitudes and alatudes should be understood as specifying the
  projected bow shock shape on the plane of the sky unless noted
  otherwise.\label{step:Pi-Lambda}
\end{enumerate}
% \subsubsection{Subjective evaluation of the fit quality}
\label{sec:subj-eval-fit}
After these automatic steps, we subjectively evaluate the resulting
fit quality by giving a star rating to each source:

\paragraph*{0 stars} The fitting algorithm failed for some reason.

\paragraph*{1 star} The fit was formally successful, but the results
for \(R_c\) or \(R_{90}\) are far removed from what a human would
predict by looking at the image.  For example, in the smallest
bowshocks, which are only marginally resolved by Spitzer's 6\arcsec{}
beam, the dispersion in \(R_k\) can be a significant fraction of
\(R_0\), in which case our algorithm tends to erroneously favor
\(R_c < R_0\).

\paragraph*{2 stars} The fit results are not totally outlandish, but
nonetheless some problem is apparent that casts doubt on their
reliability.  For example, a double-shell structure to the bow shock
that leads to large differences between the ``peak'' and ``mean''
methods, or point sources near to the bow shock that interfere with
the tracing procedure.

\paragraph*{3 stars} A good fit, but where the dispersion in \(R_k\)
and/or the asymmetry in the bow shock reduces the precision in the
determination of \(R_c\) and \(R_{90}\), giving subjectively estimated
uncertainties around the 20\% level.  An example of a 3-star fit is
shown in Figure~\ref{fig:mipsgal-examples}a.

\paragraph*{4 stars} A high quality fit, with subjectively estimated
uncertainties in \(R_c\) and \(R_{90}\) around the 10\% level. An
example of a 4-star fit is shown in
Figure~\ref{fig:mipsgal-examples}b.

\paragraph*{5 stars} The highest-quality fit, usually corresponding to
large, sharply defined bow shocks, whose shape is determined with high
precision. An example of a 5-star fit is shown in
Figure~\ref{fig:mipsgal-examples}c.
\bigskip

% 
Figure~\ref{fig:mipsgal-examples}d compares the bow shock size,
\(R_0\), determined by our fits (vertical axis) with the corresponding
value given in the \citet{Kobulnicky:2016a} catalog (horizontal axis).
For most sources with 3-star or higher rating, the two estimates agree
to within \(\pm 20\%\), but there are a small number of sources with a
discrepancy of more than a factor of two.  In all cases that we
checked, we believe that our estimates of \(R_0\) are more accurate
than those in the catalog.  It is apparent that the star ratings are
correlated with the bow shock size, with larger bow shocks tending to
receive higher ratings, although there is considerable overlap.  In
particular, most of the 1- and 2-star sources are close to the
resolution limit of the MIPSGAL \SI{24}{\um} images (\(6\arcsec\),
indicated by the dotted horizontal line in the figure).

\begin{figure*}
  \centering
  \setlength\tabcolsep{0pt}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.5\textwidth]
    {figs/mipsgal-Rc-R90-zoom-annotated}
        & \includegraphics[width=0.5\textwidth]
          {figs/mipsgal-Rc-R90-thumbnails} 
  \end{tabular}
  \vspace*{-\baselineskip}
  \caption[]{MIPSGAL sources on the bow shock shape diagnostic diagram
    of dimensionless radius of curvature versus perpendicular radius.
    The regions corresponding to different classes of quadrics are
    shown by shading (see \S~4 of Paper~0): oblate spheroids (light
    gray background); prolate spheroids (darker gray background);
    paraboloids (black line); hyperboloids (white background).
    (a) Individual sources with bow shock fit quality rating of 3-star
    or above.  All 5-star sources plus those 4-star sources with
    \(\Pi < 1\) are labelled with their \cite{Kobulnicky:2016a} catalog
    number.  Horizontal error bars do not directly reflect the
    uncertainty in \(\Pi\) but are instead simply the standard deviation
    from the circle fit of bowshock points \(R_k\) within
    \(\pm 45\degr\) of the axis.  Values on the vertical axis
    represent the average of \(\Lambda_{+}\) and \(\Lambda_{-}\), with thin
    vertical error bars showing the difference between \(\Lambda_{+}\) and
    \(\Lambda_{-}\), and thick vertical error bars showing the rms
    dispersion of \(R_k\) about these values for bow shock points
    within \(\pm 10\degr\) of the \(+90\degr\) and \(-90\degr\)
    directions.  (b) Kernel density estimator (KDE) of the
    distribution for 3-star sources (blue, filled contours) and 4-
    plus 5-star sources (orange/brown, unfilled contours).  The KDE
    uses an anisotropic gaussian kernel with bandwidths of
    \(0.06 \times 0.04\) in \(\log_{10}\) units. Thumbnail images of
    representative 4- and 5-star sources at different points on the
    \(\Pi\)--\(\Lambda\) plane are also shown. The angular scale of each image
    is indicated by a yellow circle of diameter \(7.5''\), centered on
    the stellar source.}
  \label{fig:mipsgal-shapes}
\end{figure*}

In the following analysis, only those sources with a 3-star or higher
rating are used.  These comprise approximately half (227 out of 471)
of all the MIPSGAL arc sources.  In some cases of poor and failed
fits, there is nothing apparently ``wrong'' with the source itself,
and it is likely that minor tweaks to the methodology would improve
matters, but we have elected not to do so, in order to maintain a
uniform methodology across all sources.

The inset of Figure~\ref{fig:mipsgal-examples}d shows histograms of
the difference between the position angle, \(\text{PA}_0\) determined
by our fits and that listed in \citet{Kobulnicky:2016a}.  Although
observational uncertainties undoubtedly contribute in part, the
differences are mainly due to real asymmetries in the bow shocks,
especially for the 4- and 5-star sources.  The
\citeauthor{Kobulnicky:2016a} catalog \(\text{PA}_0\) values are
mostly sensitive to the orientation of the bow shock wings, whereas
our fitted \(\text{PA}_0\) values are determined by the point in the
bow shock head that is closest to the stellar source.  For this
reason, we use the catalog \(\text{PA}_0\) values for defining the
axis when measuring \(R_{90+}\) and \(R_{90-}\). On the other hand,
the fitted values of \(\text{PA}_0\) are better correlated with the
position of the bow shock's brightness peak, as is apparent in the
lower left panels of Figure~\ref{fig:mipsgal-examples}a and c.

\subsection{OB bow shock shapes on the diagnostic plane}
\label{sec:ob-shapes}

The derived bowshock shapes of all the 3-, 4-, and 5-star sources are
shown in Figure~\ref{fig:mipsgal-shapes} on the \(\Pi\)--\(\Lambda\) plane,
which was discussed extensively in Paper~0.
Figure~\ref{fig:mipsgal-shapes}a shows each individual source with
horizontal error bars that represent the dispersion of traced points
from the circle fit (step~\ref{step:Rc} of the previous section) and
vertical error bars that show the wing asymmetry \(\Delta\Lambda\)
(step~\ref{step:Pi-Lambda}).  The horizontal error bars, which are a
proxy for observational uncertainties, become very small for the 4-
and 5-star sources, whereas the vertical error bars remain at roughly
the \(10\%\) level since they reflect real asymmetries in the shapes
of the bow shocks.

The horizontal axis corresponds to the shape of the head of the bow
shock near its apex, ranging from sharper, pointier shapes with
\(\Pi < 1\) to flatter, snubber shapes with \(\Pi \gg 1\), where it must be
understood that all judgments of sharpness/flatness are with respect
to the axial separation, \(R_0\), between the source and the bow shock
apex (see Fig.~\ref{fig:obs-shape-terminology}).  The vertical axis
corresponds to the shape of the bow shock wings, ranging from closed
``C'' shapes for smaller values of \(\Lambda\) to open ``V'' shapes for
larger values of \(\Lambda\).  The background shading shows the regions of
the \(\Pi\)--\(\Lambda\) plane occupied by simple quadric shapes (see \S~4 of
Paper~0): oblate spheroids (light gray), prolate spheroids (dark
gray), and hyperboloids (white). The boundary between closed and open
corresponds to the paraboloids, and is shown by the solid line
that divides the dark gray and white regions of the graph.

In Figure~\ref{fig:mipsgal-shapes}b, we present a smoothed version of
the same data.  The contours show the kernel density estimator (KDE,
see \citealp{Leiva-Murillo:2012a, Scott:2015a}) of the two-dimensional
distribution of points on the \(\Pi\)--\(\Lambda\) plane.  The KDE contours
indicate that the distribution of 3-star sources is very similar to
that of 4- and 5-star sources, although the higher-rated sources are
shifted slightly to the upper right.  Possible reasons for this are
discussed in \S~\ref{sec:corr-shape} below.
% This is probably due to higher-rated sources
% being on average bigger, as will be discussed further below.
The bulk of the sources are concentrated around the paraboloid line,
with \(1 < \Pi < 3\), and \(1.2 < \Lambda < 2\).  But significant minorities
are found in three other regions: (1)~a clump with \(\Pi \la 1\); (2)~a
vertical spur towards higher \(\Lambda\) at \(\Pi \approx 1.8\); and (3)~a broad
horizontal tail towards higher \(\Pi\) at \(\Lambda \la 2\).  Thumbnail images
of the \SI{24}{\um} emission of selected sources from different
regions of the distribution are shown, in order to illustrate the
range of bow shock morphologies encountered.

\subsection{Correlations between non-shape parameters of the OB sources}
\label{sec:corr-size}

\begin{figure*}
  \centering
  \includegraphics[width=0.86\textwidth]{figs/mipsgal-pairplot}
  \vspace*{-\baselineskip}
  \caption[]{Matrix of pair plots that illustrate distributions of and
    correlations between the non-shape parameters of all MIPSGAL bow
    shock sources from \citet{Kobulnicky:2016a}.  Plots on the leading
    diagonal show histograms of the following parameters: bow shock
    angular size, \(\log_{10} R_0\); Galactic latitude,
    \(\log_{10}|b|\); Galactic longitude, \(\cos \ell\);
    extinction-corrected \(H\)-band magnitude of the stellar source,
    \(H_0\); \(K\)-band extinction, \(A_K\).  Scatter plots in the
    upper triangle show the joint distribution of each pair of
    parameters.  These are repeated in the lower triangle but showing
    the KDEs of the joint distributions, which are annotated with the
    Pearson linear correlation coefficient, \(r\), for each pair. The
    straight lines shown superimposed on the plots of stellar magnitude
    versus bowshock size correspond to toy model results for the same star
    at a sequence of distances (dashed lines) and a sequence of
    stellar luminosities at a fixed distance (dotted lines).  See text
    for details. }
  \label{fig:mipsgal-pairplot}
\end{figure*}

In Figure~\ref{fig:mipsgal-pairplot} we show the distributions over
all MIPSGAL bow shock sources of the bow shock size, Galactic
coordinates, extinction-corrected stellar source magnitude, and dust
extinction.  For the bow shock size, \(R_0\), we use the results from
our model fitting rather than the values given in the
\citet{Kobulnicky:2016a} catalog, but the distribution is very
similar, as can be seen by comparing the top-left plot of
Figure~\ref{fig:mipsgal-pairplot} with \citeauthor{Kobulnicky:2016a}'s
Figure~8.

The catalog gives the \(K\)-band extinction, \(A_K\), derived using
the method of \citet{Majewski:2011a}. However, that assumes an
intrinsic color of \((H - [\SI{4.5}{\um}])_0 = +0.08\) magnitudes,
which is too red if the sources are assumed to be OB stars.  We
therefore re-derive \(A_K\) from the catalog magnitudes combined with
the \citet{Indebetouw:2005a} reddening law, but assuming
\((H - [\SI{4.5}{\um}])_0 = -0.1\) magnitudes, which is more typical
of early type stars.  This does not make very much difference (compare
the top-right plot of our Fig.~\ref{fig:mipsgal-pairplot} with
\citeauthor{Kobulnicky:2016a}'s Fig.~9), but it does eliminate some of
the apparent negative extinctions that are found in the catalog.  The
same reddening law gives \(A_H = 1.55 A_K\), and this is used to
derive extinction-corrected \(H\)-band apparent magnitudes, \(H_0\).

\begin{figure*}
  \setlength\tabcolsep{0pt}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.5\textwidth]{figs/mipsgal-Rc-R90-environment} &
    \includegraphics[width=0.5\textwidth]{figs/mipsgal-Rc-R90-candidates} 
  \end{tabular}
  \vspace*{-\baselineskip}
  \caption[]{Example comparisons between the distribution of bowshock
    shapes when the sources are divided into two or more sub-samples
    according to the value of a categorical parameter.  Contours show
    the KDE of the distribution of each sub-sample for all 3-, 4-, and
    5-star sources, while filled circle symbols show 5-star sources
    only. Inset histograms show the marginal distributions on the two
    shape axes. (a)~Source environment, divided into three
    sub-samples: ``Isolated'' (purple symbols and purple filled
    contours), ``Facing \hii{} region or \SI{8}{\um} bright-rimmed
    cloud'' (orange symbols and orange-brown hollow continuous
    contours), and ``Within \hii{} region'' (green symbols and green
    hollow dashed contours). (b)~Uncertainty in stellar source
    identification, divided into two sub-samples: ``Multiple
    candidates for stellar source'' (green symbols and green filled
    contours) and ``Single candidate for stellar source'' (orange
    symbols and orange-brown hollow continuous contours).}
  \label{fig:mipsgal-uncorrelated}
\end{figure*}
\begin{figure*}
  \setlength\tabcolsep{0pt}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[width=0.5\textwidth]{figs/mipsgal-Rc-R90-R0} &
    \includegraphics[width=0.5\textwidth]{figs/mipsgal-Rc-R90-Mag} 
  \end{tabular}
  \vspace*{-\baselineskip}
  \caption[]{As Fig.~\ref{fig:mipsgal-uncorrelated} but for median
    splits of continuous parameters. (a)~Bow shock angular size,
    \(R_0\), divided into two equal-sized sub-samples: large (blue
    symbols and blue filled contours) and small (red hollow contours).
    (b)~Extinction-corrected \(H\)-band magnitude of the stellar
    source, divided into two equal-sized sub-samples: bright (red
    symbols and orange filled contours) and faint (green symbols and
    green hollow contours).}
  \label{fig:mipsgal-correlated}
\end{figure*}

The most significant linear correlation between any pair of parameters
in Figure~\ref{fig:mipsgal-pairplot} is that between bow shock size
and stellar source brightness: \(H_0\) versus \(\log_{10} R_0\), with
correlation coefficient \(r = -0.43\).  The distribution of \(H_0\)
depends on the absolute magnitude, \(M_H\), and the distance, \(d\),
to the source.  It is likely that variation in \(d\) is the more
important of the two because \(M_H\) changes relatively little for
main-sequence OB stars, ranging from \(M_H \approx -4\) (early-O) to
\(M_H \approx -1.5\) (mid-B).  This is because part of the increase in
bolometric luminosity, \(L\), as one ascends the main sequence is
offset by an increase in the effective temperature,
\(T_{\text{eff}}\), which shifts the peak of the stellar spectrum
farther away from the \(H\)~band, resulting in
\(L_H \propto L / T_{\text{eff}}^3 \sim L^{0.53}\), where the last step uses
the upper main-sequence mass--luminosity and mass--radius scalings from
\citet{Eker:2015a}. It is true that evolved OB supergiants can be much
brighter, reaching \(M_H \approx -7\), but such stars are expected to be
relatively rare.  Assuming a B2V star (\(M_H = -2\)), then the
observed range \(H_0 = 5\)--\(12\) corresponds to distances
\(d = 100\)--\SI{6300}{pc}, and the histogram peak at
\(H_0 \approx 9.5\) corresponds to \(d \approx \SI{2000}{pc}\), which is all
perfectly reasonable.

Turning now to the distribution of bow shock angular size, \(R_0\),
this will also be affected by distance to at least some degree, since
for a constant physical size the angular size will vary as
\(R_0 \propto d^{-1}\).  For instance, if we assume that the physical size
of all bow shocks is \SI{0.1}{pc} and the absolute magnitude of all
stars is \(M_H = -2\), as above, then we find the relation
\(H_0 = 14.57 - 5 \log_{10} R_0\) if \(R_0\) is measured in
arcseconds. This is shown as a dashed line on the relevant panels of
Figure~\ref{fig:mipsgal-pairplot} for values of \(R_0\) that
correspond to \(d = \SI{300}{pc}\) to \SI{8000}{pc}.  It can be seen
that this relation is in excellent agreement with the linear trend in
the data. On the other hand, the correlation coefficient of
\(r = -0.43\) means in broad terms that only a fraction
\(r^2 \approx 20\%\) of the total variance in \(H_0\) is ``explained'' by
changes in \(R_0\), and vice~versa, implying that one or both of
\(H_0\) and \(R_0\) is only a very imperfect proxy for \(d\).  We have
already seen that the spread in \(H_0\) probably \emph{is} mostly due
to a spread in distance, rather than a spread in \(H\)-band stellar
luminosity.  If this is true, it follows that it is \(R_0\) that
depends only weakly on \(d\) and is more influenced by other factors.

One such factor is the stellar/environmental momentum-loss ratio,
\(\beta\), between the two supersonic flows that form the bow shock.  All
other things being equal, we have \(R_0 \propto \beta^{1/2}\) for
\(\beta \ll 1\), as is typically the case.  If the environment flow is
constant and the OB star wind has mass-loss rate \(\dot{M}\) and
terminal velocity \(V_\infty\), then
\(\beta \propto \dot{M}{V_\infty}\).  Empirical and theoretical studies of hot star
winds (e.g., \citealp{Puls:1996a}) imply
\(\dot{M}{V_\infty} \sim L^{1.88} R^{-1/2} \sim L^{1.80} \sim L_H^{3.40}\), where
the final two steps apply only to main-sequence stars and again use
the relations of \citet{Eker:2015a}.  If we assume as above that a B2V
star with \(M_H = -2\) has a bow shock physical size of \SI{0.1}{pc},
and consider a sequence of stars with varying \(H\)-band luminosities
but all at a fixed distance of \(d = \SI{2000}{pc}\), then we find the
relation \(H_0 = 8.02 - 1.49 \log_{10} R_0\).  This is shown as a
dotted line in the relevant panels of
Figure~\ref{fig:mipsgal-pairplot} for the absolute magnitude range
\(M_H = -3.6\) (O6V) to \(M_H = -1.5\) (B5V).  It can be seen that
this relation does not match the linear trend in the data, and
predicts a much larger spread in \(R_0\) over a narrow range in
\(H_0\) than is observed.  This could mean one of two things: first,
it may be that the range of stellar luminosities is significantly
narrower than we have supposed, implying that B stars vastly outnumber
O stars among the sources.  Alternatively, there may be a positive
correlation between the stellar luminosity and the momentum of the
environmental flow, with the result that \(\beta\) varies less steeply
with \(L_H\) than we have assumed.  That could arise if more luminous
stars were preferentially found in denser environments, or, in the
case of runaways, if more luminous stars tended to be faster moving.

A third factor that may influence \(R_0\) is the inclination, \(i\),
of the bow shock axis with respect to the plane of the sky.
Figure~11b of Paper~0 shows that for \(\Pi > 1\), then the projected
\(R_0'\) becomes larger than the true \(R_0\) as \(|i|\) increases
(see also Figs.~26 and 27 of Paper~0).  It can be seen that the effect
is relatively modest, with an increase in \(R_0\) of no more than a
factor of 2 to 3.

%% Maybe talk about the other correlations later if we have time

\subsection{Correlation between bow shock shape and other parameters}
\label{sec:corr-shape}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]
  {figs/mipsgal-boxplot-Rc-R90-versus-R0}
  \vspace*{-\baselineskip}
  \caption{Box plots of the distributions of bow shock shape
    parameters, \(\Pi\) and \(\Lambda\), together with a measure of the tail
    asymmetry, \(2 |\Delta\Lambda| / \Lambda\), after partitioning on the bow shock
    size, \(R_0\).  All 3-, 4-, and 5-star sources are sorted
    according to \(R_0\) and divided into 5 non-overlapping
    sub-samples of roughly equal size, each labelled by their median
    value of \(R_0\) in arcseconds, as illustrated in the lower panel.
    The colored boxes show the interquartile range of the data, with
    the median indicated by a horizontal line and the notch showing
    the 90\% confidence limits of the median, as determined from 1000
    bootstrap resamplings.  The 5th-to-95th centile range of the data
    is indicated by error bars.  A 5-sample Anderson--Darling test and
    Brown--Forsythe test is performed for each dependent variable, with
    resultant \(p\)-value given at the bottom of each panel.  It is
    apparent that the dispersion in \(\Lambda\) (upper panel) increases
    systematically with \(R_0\), although the central value is roughly
    constant.  No clear systematic changes are apparent in \(\Pi\)
    (second panel).}
  \label{fig:mipsgal-boxplot}
\end{figure}

We now investigate if the bow shock shapes of the MIPSGAL sources are
correlated with any other parameters via the following methodology:
\begin{enumerate}[1.]
\item For each of the parameters in the \citet{Kobulnicky:2016a}
  catalog, we divide the sources into two or more sub-samples,
  according to the value of the parameter.  For quantitative
  parameters, such as those discussed in the previous section, we use
  two sub-samples of equal size, with membership determined by whether
  the parameter is larger or smaller than the median value.  But we
  also use categorical parameters, such as the type of bow shock
  environment, or the presence/absence of \SI{8}{\um} emission, in
  which case the sub-samples are of unequal size.
\item We plot the KDEs of the two sub-samples separately on the
  \(\Pi\)--\(\Lambda\) plane (see Figs.~\ref{fig:mipsgal-uncorrelated} and
  \ref{fig:mipsgal-correlated}) to check for any obvious
  differences.
\item We check if there is any statistically significant difference
  between the bow shock shapes of the sub-samples by applying three
  different non-parametric tests to \(\Pi\) and \(\Lambda\)
  separately.  
\end{enumerate}

The principal test used is Kuiper's variant of the two-sample
Kolmogorov--Smirnov test \citep[e.g.,][]{Stephens:1970a, Paltani:2004a}, as implemented by the Python library function \texttt{astropy.stats.kuiper\_two} \citep{Astropy-Collaboration:2018a}.   This is a
general test of the \textit{null hypothesis} that the two sub-samples
are drawn from the same distribution.  It returns a \(p\)-value, which
is the estimated probability that the observed difference between the
two sub-samples would be as large as it is purely by chance if they
were all were drawn from the same distribution.  We consider two
different thresholds for significance: \(p < 0.05\) (approximately
2-\(\sigma\) for a normal distribution) and \(p < 0.003\) (approximately
3-\(\sigma\)).  We show in Appendix~\ref{sec:distr-p-values} that, given
the large number of potential relationships that we are testing, the
more stringent \(p < 0.003\) condition is required in order to
confidently reject the null hypothesis and declare a ``significant''
difference between the two sub-samples, but we also consider the less
strict threshold of \(p < 0.05\) as an indicator of ``possible''
difference.  We supplement the general-purpose Kuiper test
with two tests that probe specific features of the sample
distributions: the Mann--Whitney--Wilcoxon \(U\) test
\citep{Mann:1947a}, which is sensitive to differences in the central
value (e.g., median) and the Brown--Forsythe test \citep{Brown:1974a},
which is sensitive to differences in the variance, or width, of the
distribution (see Appendix~\ref{sec:distr-p-values} for details).

As shown in detail in Table~\ref{tab:big-p}, there is remarkably
little variation in the bow shock shape distributions as a function of
most of the other parameters.  Two examples in which there is
\textit{no} significant shape difference between the sub-samples are
shown in Figure~\ref{fig:mipsgal-uncorrelated}.  This lack of
difference is interesting because in both examples there are a priori
grounds to suspect that differences might exist.  The first example
(Fig.~\ref{fig:mipsgal-uncorrelated}a) is the bow shock environment,
which was categorized in \citet{Kobulnicky:2016a} as ``Isolated'' (I),
``Facing a large \hii{} region'' (FH), ``Facing a \SI{8}{\um}
bright-rimmed cloud'' (FB), and ``Within a giant \hii{} region'' (H),
and where we have merged the FH and FB categories, labelled ``Facing''
in the figure.\footnote{Similar results are also found for the FB and
  FH categories separately.}  The shapes might be expected to vary
with environment because the FB and FH categories should be associated
with ``weather vane'' interactions \citep{Povich:2008a} between the
stellar wind and a divergent photoevaporation flow.  This is expected
to give a more open bow shock than in the ``runaway'' case of
interaction of a moving star with a static environment.  In the
thin-shell approximation, the predicted shapes are a cantoid for
weather vanes and a wilkinoid for runaways, see \S~6 of Paper~0.  The
fact that no such difference is detected could be explained in one of
two ways: (i)~the momentum ratio \(\beta\) for the weather vanes could be
small, since the cantoid becomes indistinguishable from the wilkinoid
as \(\beta \to 0\), or (ii)~many of the bow shocks classified as
``Isolated'' might also be weather vanes rather than runaways.

The second example (Fig.~\ref{fig:mipsgal-uncorrelated}b) divides the
sources according to whether or not \citet{Kobulnicky:2016a} judged
there to be multiple candidates for the identity of the central star
that drives the bow shock.  If the central star were to be
misidentified in a significant number of sources, then the measured
value of \(R_0\) for those sources would be erroneous, which would
increase the scatter in both \(\Pi\) and \(\Lambda\).  The fact that no
significant difference is seen in the shape distributions between
sources with/without multiple candidates implies that such mistakes in
identification of the central star must be rare.

We also tested all the other source parameters listed in
\citeauthor{Kobulnicky:2016a}'s catalog, finding no significant shape
differences for sources with/without \SI{8}{\um} emission, with
low/high extinction, closer/farther from the Galactic plane, or
closer/farther from the Galactic center. Details are given in
Table~\ref{tab:big-p}.  In all these cases, differences in mean or
median \(\Lambda\) less than \(0.06\) and in \(\Pi\) less than
\(0.16\) are found, corresponding to rank biserial correlation
coefficients (a non-parametric dimensionless measure of the difference
between two samples, see Appendix~\ref{sec:distr-p-values}) of
\(r_b < 0.15\), which are not significant even at the 2-\(\sigma\) level.

The only parameters that \emph{do} show a possible correlation with
the bowshock shape are the bow shock angular size, \(R_0\), and the
extinction-corrected magnitude of the central star, \(H_0\), which are
illustrated in Figure~\ref{fig:mipsgal-correlated}.  The general
purpose Kuiper test indicates differences in the sub-sample
distributions of \(\Lambda\) at the 2-\(\sigma\) level for \(H_0\) and at the
3-\(\sigma\) level for \(R_0\).  It is apparent from the histograms shown
in the right-hand inset graphs of Figure~\ref{fig:mipsgal-correlated}a
that in the case of the small/large \(R_0\) sub-samples it is the
width rather than the central tendency of the distributions that is
different, which is confirmed by the more specific Brown--Forsythe
test, which indicates a difference between the sub-sample dispersions
at the 3-\(\sigma\) level.  In the case of the faint/bright \(H_0\)
sub-samples (Fig.~\ref{fig:mipsgal-correlated}b), it is less clear
what feature of the distributions differ.

In order to investigate these effects in more detail and look for
systematic trends, we divide the independent parameter (\(R_0\) or
\(H_0\)) into 5 rather than 2 equal-sized sub-samples, with results in
the case of \(R_0\) shown as box plots in
Figure~\ref{fig:mipsgal-boxplot}.  This time the \(k\)-sample
Anderson--Darling test \citep{Anderson:1952a, Scholz:1987a,
  Makarov:2017a} is used to determine the statistical significance of
any observed differences.  A systematic increase with \(R_0\) in the
dispersion of \(\Lambda\) (upper panel of Fig.~\ref{fig:mipsgal-boxplot}a)
is apparent from both the interquartile range (colored boxes) and
interdecile range (error bars).  As a check on whether observational
uncertainties might be contributing to this trend, the third and
fourth rows of box plots show the statistics for, respectively, the
fractional asymmetry of the bowshock wings and the standard deviation,
\(\epsilon(\Lambda)\), of the individual points on the bow shock that go into
determining \(\Lambda\) (see step~\ref{step:R90} of the tracing/fitting
methodology described in \S~\ref{sec:autom-trac-fitt}).  It can be
seen that neither of these quantities tends to increase with \(R_0\),
and in fact there is a significant \emph{decrease} in
\(\epsilon(\Lambda)\) with \(R_0\).  This implies that the increase with angular
size of the diversity of bow shock wing shapes is real, and not due to
observational uncertainties.  On the other hand, we find that no clear
trends are evident as a function of source magnitude \(H_0\) (not
illustrated), which leads us to suspect that the \(p < 0.05\) result
obtained for the 2-sample Kuiper test was a \emph{false positive}, as
is also supported by the analysis in Appendix~\ref{sec:distr-p-values}
and Figure~\ref{fig:histo-p-values}.

% So, the only correlation that survives scrutiny is the increase in
% dispersion of \(R_{90}/R_0\) with increasing \(R_0\), which is finally
% presented in the greatest detail in
% Figure~\ref{fig:mipsgal-heteroscedastic}, where we eschew any binning
% and perform linear regressions of the individual sources as a function
% of \(\log_{10} R_0\).  As expected from the previous 2-sample and
% 5-sample analyses, the regression of \(R_{90}/R_0\) is almost flat,
% with a low correlation coefficient, whereas a regression of the
% absolute residuals to this fit against \(\log_{10} R_0\) shows a
% significant positive slope.  This increase in dispersion of shape with
% angular size is only evident for \(R_0 < 30''\), becoming flat at the
% largest sizes (see middle panel of
% Fig.~\ref{fig:mipsgal-heteroscedastic}).

As mentioned in \S~\ref{sec:ob-shapes} there is also a shape
difference between the 3-star sources and the 4/5-star sources (see
Fig.~\ref{fig:mipsgal-shapes}b).  The \(p\)-values of statistical
tests (see Table~\ref{tab:big-p}) indicate that this is much more
significant than any correlation with the other parameters discussed
in this section (\(p < 10^{-4}\) for \(\Lambda\) and \(p < 10^{-5}\) for
\(\Pi\)).  This means that it cannot be simply due to the tendency of
the higher-rated sources to have larger angular sizes.  However, the
subjective nature of the star ratings makes this result hard to
interpret.

\section{Far-infrared arcs around late-type stars}
\label{sec:far-infrared-arcs}

\begin{figure*}
  \setlength\tabcolsep{0pt}
  \setkeys{Gin}{width=0.5\linewidth}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/alphaori-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/mucep-imageplot}\\
    (c) & (d) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/rhya-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/rleo-imageplot}\\
    (e) & (f) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/uuaur-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/v1943sgr-imageplot}\\
    (g) & (h) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/xpav-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/rcas-imageplot}
  \end{tabular}
  \caption{RSG and AGB shells observed with Herschel.  Selected
    Class~I sources (``fermata-like'') from the MESS survey
    \citet{Cox:2012a}, where the arc structure is particularly clear
    and symmetrical.  Left panels show PACS \SI{70}{\um} surface
    brightness (blue) and \SI{160}{\um} (orange).  Right panels show
    tracing of the bow shock arc (red symbols) and circle fit (magenta
    lines and symbols) superimposed on a low-contrast image of the
    \SI{70}{\um} surface brightness. (a)~\(\alpha\)~Ori. (b)~\(\mu\)~Cep.
    (c)~R~Hya. (d)~R~Leo. (e)~UU~Aur. (f)~V1934~Sgr. (g)~X~Pav.
    (h)~R~Cas.}
  \label{fig:herschel-arc-fits}
\end{figure*}

\begin{figure*}
  \setlength\tabcolsep{0pt}
  \setkeys{Gin}{width=0.5\linewidth}
  \begin{tabular}{ll}
    (a) & (b) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/cwleo-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/epaqr-imageplot} \\
    (c) & (d) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/khicyg-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/rtvir-imageplot} \\
    (e) & (f) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/waql-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/wpic-imageplot} \\
    (g) & (h) \\
    \includegraphics[trim=10 0 65 20, clip]{figs/rscl-imageplot}
    & \includegraphics[trim=10 0 65 20, clip]{figs/tetaps-imageplot}
  \end{tabular}
  \caption{As Fig~\ref{fig:herschel-arc-fits}, but for a further 8
    Class~I sources from \citet{Cox:2012a} where the arc structure is
    more diffuse, weak, and/or asymmetric. (a)~CW~Leo. For this source
    only, the right-hand panel shows a grayscale image of the
    \SI{160}{\um} rather than \SI{70}{\um}
    emission. (b)~EP~Aqr. (c)~\(\chi\)~Cyg. (d)~RT~Vir. (e)~W~Aql. (f)~W~Pic. (g)~R~Scl.
    (h)~\(\theta\)~Aps.}
  \label{fig:herschel-arc-fits-poor}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/mipsgal-Rc-R90-vs-Herschel}
  \vspace*{-\baselineskip}
  \caption[]{Comparison of bow shock shapes between Herschel RSG/AGB
    arcs (large red symbols and hollow red contours) and MIPSGAL OB
    star arcs (small purple symbols and filled purple contours).
    Other details of the plots are as in
    Figs.~\ref{fig:mipsgal-shapes}, \ref{fig:mipsgal-uncorrelated},
    and \ref{fig:mipsgal-correlated}.}
  \label{fig:herschel-compare-mipsgal}
\end{figure}

We obtain a second sample of bow shocks from a far-infrared survey
\citep{Cox:2012a} of circumstellar shells around known asymptotic
giant branch (AGB) stars and cool supergiants, obtained as part of the
\textit{Herschel} MESS (Mass-loss of Evolved StarS) program
\citep{Groenewegen:2011a}.  The survey sources were divided into four
classes by \citeauthor{Cox:2012a}, according to the overall shape of
the extended emission at \SI{70}{\um} (see their Table~4):
``fermata~\textfermata'', or arcs (Class~I); ``eyes~\faEye''
(Class~II); ``rings~{\Large \(\odot\)}'' (Class~III), and
``irregular~\staveXXV'' (Class~IV).  Of these, only the Class~I
sources clearly correspond to bow shocks, which represent 22 out of 50
total sources detected with extended emission.  For our sample, we
select the 16 Class~I sources where the shape parameters can be
reliably measured, and which are shown in
Figures~\ref{fig:herschel-arc-fits}
and~\ref{fig:herschel-arc-fits-poor}. The remaining six Class~I
sources (Fig.~1 of \citealp{Cox:2012a}) are too asymmetrical or
irregular to reliably determine \(R_{c}\) and \(R_{90}\), and are
therefore excluded from our sample.

In most sources, the bow shock shell is most clearly visible in the
\SI{70}{\um} band, with the exception being CW~Leo where the
\SI{180}{\um} band is used instead.  In several sources, the shell is
split into multiple filaments, with the clearest example being
\(\alpha\)~Ori (Betelgeuse), shown in
Figure~\ref{fig:herschel-arc-fits}(a).  For such sources, we take the
outer envelope of the filaments as the bow shock arc.  Given this
complication, and the fact that the angular resolution relative to the
bow shock size is much better than in the MIPSGAL sources, we judge
that the arc tracing is best performed by eye and this is carried out
using the SAOImage~DS9 FITS viewer \citep{Joye:2003a}, with results
shown as small red open circles in Figures~\ref{fig:herschel-arc-fits}
and~\ref{fig:herschel-arc-fits-poor} (more details of this technique
are given in \S\S~6 and 7 of Paper~0).  Subsequently, the bow shock
parameters \(R_0\), \(\Pi\), and \(\Lambda\) are determined by circle fits, as
in steps 5--9 of \S~\ref{sec:autom-trac-fitt}, with results shown in
magenta on the figures.

Figure~\ref{fig:herschel-compare-mipsgal} compares the distributions
of bow shock shapes between the Herschel (RSG/AGB) and MIGSGAL (OB
star) samples.  The Kuiper test gives a highly significant
difference between the \(\Lambda\) distributions of the two samples, but
only a marginally significant difference between the \(\Pi\)
distributions (detailed results are shown in Table~\ref{tab:big-p}).
The Herschel sources show considerably smaller alatude (median
\(\Lambda \approx 1.4\)), implying bow shock wings that are more closed than in
the MIPSGAL sources (median \(\Lambda \approx 1.7\)).  For the planitude, there is
only a slight difference in average values: median \(\Pi \approx 1.5\) for
Herschel versus \(\approx 1.6\) for MIPSGAL, which is not statistically
significant (rank biserial \(p = 0.29\)).  On the other hand, the
dispersion in \(\Pi\) is four times smaller for the Herschel sample,
which is marginally significant (Brown--Forsythe \(p = 0.015\)).  In
particular, the MIPSGAL sample shows a substantial minority of very
flat-nosed shapes (\(\Pi > 2\)), but these are absent in the Herschel
sample.  The same can be seen directly by comparing the RSG/AGB bow
shock shapes in Figures~\ref{fig:herschel-arc-fits}
and~\ref{fig:herschel-arc-fits-poor} with the representative OB bow
shocks of Figure~\ref{fig:mipsgal-shapes}b.  The MIPSGAL sources K489
and K123 have shapes that are similar to specific Herschel sources
(UU~Aur and R~Hya, respectively), while sources such as K447 (extreme
flat head) or K517 (extreme open wings) have no analog among the
Herschel sources.

\section{Stationary emission line arcs in M42}
\label{sec:stat-emiss-line}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/annotated-ll-arcs}
  \vspace*{-\baselineskip}
  \caption[]{Stationary bow shock arcs in the Orion Nebula.  Images
    are of H\(\alpha\) plus [\ion{N}{ii}] emission through the HST ACS f658n
    filter. The angular scale of each image is different, with the
    yellow horizontal lines indicating \(5''\), corresponding to a
    physical scale of \(\approx\SI{0.01}{pc}\).  Orange arrows indicate the
    direction to the principal ionizing star, \thC. }
  \label{fig:ll-arcs}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/mipsgal-Rc-R90-vs-Orion-R0-D}
  \vspace*{-\baselineskip}
  \caption[]{Shape comparison between Orion Nebula arcs (green) and OB
    star bow shocks (purple).  Other details of the plots are as in
    Figs.~\ref{fig:mipsgal-shapes} and~\ref{fig:mipsgal-uncorrelated},
    except that a larger KDE smoothing bandwidth of
    \(0.06 \times 0.045\) is used for the Orion Nebula sample.  The colors
    of the individual points for the Orion Nebula sample show the
    size-to-distance ratio, \(R_0/D\), on a logarithmic scale from
    \(R_0/D = 0.002\) (dark green) to \(R_0/D = 0.018\) (light
    green).}
  \label{fig:ll-compare-mipsgal}
\end{figure}

We obtain a third sample of bow shocks from a catalog of stationary
emission line arcs in the Orion Nebula \citep{Bally:2000a}, which are
detected via their H\(\alpha\) emission in HST~ACS surveys of the nebula
\citep{Bally:2006a, Robberto:2013a}.  Out of a total of 73 such
objects \citep{Gutierrez-Soto:2015a} we have selected 18 where the
observations are of sufficient quality to reliably measure the
planitude and alatude.  Images in the ACS f658n filter of these
sources are shown in Figure~\ref{fig:ll-arcs}.  Some of the sources
were identified as LL~Orionis-type objects by \citet{Bally:2001a} and
are identified as LL~1--6. The remaining sources are labeled by their
coordinate-based designation according to the system of
\citet{ODell:1994a}.  The determination of the bow shock shape for
each source was carried out as described in \S~7 of Paper~0, using the
``ridge'' method for manually tracing the arc (see Fig.~28 of
Paper~0).  In some sources, such as LL~1, LL~6, 109--246, and 4468--605,
a high-velocity jet flow is also seen to issue from the star
\citep{Bally:2006a, Henney:2013a}.  We took care when tracing the arcs
of such sources to avoid regions where jet knot emission is projected
on the bow shock shell.

The planitude and alatude are found by circle fitting using the
algorithm described in Paper~0's Appendix~E, as implemented in the
python program
\texttt{circle-fit.py}.\footnote{\url{https://github.com/div-B-equals-0/circle-fit}}
The only adjustable parameter of the algorithm is \(\Delta\theta\), the maximum
angle from the bow shock axis of points that are included in the
circle fit.  We tested all values of \(\Delta\theta\) in \ang{5} increments from
\ang{45} to \ang{80} and found that \(\Delta\theta = \ang{55}\) to \ang{70}
gives stable results for all sources.  The accuracy of the fitted
planitudes can be estimated from the dispersion in values of \(\Pi\) for
the same source with different \(\Delta\theta\), which we find to be
\(\approx 10\%\).  The fitted alatudes have a much smaller dispersion
\(\approx 1 \%\), but in this case the accuracy is limited by the asymmetry
of the wings, as characterized by \(\Delta\Lambda/\Lambda\) with an RMS value of
\(\approx 15\%\).

The resultant shape distribution for the Orion Nebula arcs is shown in
Figure~\ref{fig:ll-compare-mipsgal}, where it is compared with the OB
star shapes from \S~\ref{sec:ob-shapes}.  It can be seen that both the
planitude and alatude distributions are shifted towards higher values,
with median values of \(\Pi = 2.70\) and \(\Lambda = 2.55\).  The difference
from the MIPSGAL sources (\(\Pi = 1.57\) and \(\Lambda = 1.69\)) is highly
statistically significant, as indicated by extremely low \(p\)-values
for the Kuiper tests.  It is also at least 4 times larger
than the systematic fitting uncertainties discussed in the previous
paragraph.  On closer inspection, it appears that the shape
distribution is bimodal.  Four of the sources (005--514, 4468--605,
116--3101, 308--3036) are similar to typical OB stars, with
\(\Lambda \approx \Pi \approx 1.5\), whereas the remainder are concentrated in the range
\(\Pi = 2\) to \(4\), with an extended tail towards higher planitude.
The alatudes mainly follow the line for parabolas,
\(\Lambda = (2\Pi)^{1/2}\) (see App.~C of Paper~0), and the deviations from
this line can be readily understood in terms of peculiar morphologies
of the source (Fig.~\ref{fig:ll-arcs}).  For instance, LL~4 and
266--558 both show large alatudes for their respective planitudes and
these are both sources where the bow shock wings seem to bend outwards
instead of following the curve from the apex region.  In contrast,
030--524 has a small alatude for its planitude and this is a very
asymmetric source with an apparent corner to the shell on one side.

Uniquely among the three bow shock datasets, for the Orion Nebula arcs
we know the source of the external flow.  This is not the stellar wind
from the high-mass stars at the center of the cluster, which has
insufficient momentum at the position of the arcs \citep{Bally:2000a},
nor is it due to stars' own motions, since the space velocities of all
these sources are very low, conforming to the velocity dispersion
\(\sigma_{\text{1D}} \approx \SI{2}{km.s^{-1}}\) of the Orion Nebular Cluster
\citep{Dzib:2017a, Kim:2018b}.  Instead it is due to the transonic
champagne flow of photoionized gas away from the core of the nebula
\citep{Zuckerman:1973a, Henney:2005a}, which has its density peak in
the Orion~S region \citep{Weilbacher:2015a}, located roughly \(30''\)
SW of the dominant Trapezium star \thC.  The bow shock sources are
located at much larger distances of \(D = 90''\) to \(700''\) from
\thC, so that the ratio \(R_0/D\) can be used as a proxy for the local
divergence of the champagne flow on the scale of the bow shock (notice
that the symmetry axes of the bow shocks in Fig.~\ref{fig:ll-arcs} are
always roughly parallel to the direction to \thC{}).  This is
indicated by the color of the points in
Figure~\ref{fig:ll-compare-mipsgal}, with lighter colors corresponding
to a larger \(R_0/D\) and hence a more divergent external flow.  It
can be seen that there is some correlation between the bow shock
shapes and \(R_0/D\): the source with the largest \(\Pi\), 109--246, also
has the largest \(R_0/D\), while the four sources with low \(\Pi\) and
\(\Lambda\) all have low \(R_0/D\).  However, the correlation is far from
perfect, with a Pearson correlation coefficient of only \(r = 0.47\).

%Mention future JWST observations with \(0.85\arcsec\) resolution at \SI{22}{\um}.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "obs-bowshocks.tex"
%%% End: